### 2009-07-18 inst/doc/intro2fitdistrplus.Rnw
add an header for R files.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%   Copyright (c) 2009 Marie Laure Delignette-Muller, Regis Pouillot, Jean-Baptiste Denis,
%   Christophe Dutang                        
%                                                                                                                                                       
%    This program is free software; you can redistribute it and/or modify                               
%    it under the terms of the GNU General Public License as published by                        
%    the Free Software Foundation; either version 2 of the License, or                                 
%    (at your option) any later version.                                                                                        
%                                                                                                                                                      
%    This program is distributed in the hope that it will be useful,                                          
%    but WITHOUT ANY WARRANTY; without even the implied warranty of                        
%    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the                   
%    GNU General Public License for more details.                                                                    
%                                                                                                                                                         
%    You should have received a copy of the GNU General Public License                           
%    along with this program; if not, write to the                                                                            
%    Free Software Foundation, Inc.,                                                                                             
%    59 Temple Place, Suite 330, Boston, MA 02111-1307, USA                                            
%                                                                                                                                                         
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% fitdistrplus
%%%
%%%			Sweave vignette file
%%%

\documentclass[a4paper]{article}

% sweave commands for vignette
%\VignetteIndexEntry{Fit parametric distributions on non-censored or censored data}
%\VignettePackage{fitdistrplus}
%\VignetteKeyword{distribution}

%geometry
\usepackage[a4paper,textwidth=18cm,textheight=27cm]{geometry}

%sweave
\usepackage{Sweave}

%footnote
\usepackage[stable]{footmisc}
\usepackage{perpage}
\MakePerPage[1]{footnote}

%citation
\usepackage{natbib}

%reference hypertext
\usepackage[pagebackref=true, hyperfootnotes=false]{hyperref}

%title
\title{Use of the library \texttt{fitdistrplus} to specify a distribution from 
non-censored or censored data}

%authors
\author{Marie Laure Delignette-Muller, R\'egis Pouillot, Jean-Baptiste Denis and Christophe Dutang}

% macros
\newcommand\HRule{\noindent\rule{\linewidth}{1pt}}
\newcommand{\II}{\mbox{\large 1\hskip -0,353em 1}}
\newcommand{\abs}[1]{\lvert#1\rvert}
\newcommand{\norm}[1]{\lVert#1\rVert}
\newcommand{\ligne}{\rule[2mm]{.3\textwidth}{0,5mm}\\}
\newcommand{\pkg}{\textbf}
\newcommand{\sigle}{\textsc}
\newcommand{\code}{\texttt}
\newcommand{\soft}{\textsf}
\newcommand{\myskip}{\vspace{\parskip}}
\newcommand{\txtm}[1]{\textrm{~~#1~~}}
\newcommand{\expo}{\textsuperscript}
\newcommand{\ind}{1\!\!1}
\newcommand{\mat}[1]{\mathbf{#1}}
\newcommand{\blank}{ \clearpage{\pagestyle{empty}\cleardoublepage} }
\newcommand{\card}{\textrm{Card}}

\begin{document}

\maketitle

Here you will find some easy examples of use of the functions of the library \pkg{fitdistrplus}. 
The aim is to show you by examples how to use these functions to help you to specify a parametric distribution 
from data corresponding to a random sample drawn from a theoretical distribution that you want to describe. 
For details, see the documentation of each function, using the R help command (ex.: \code{?fitdist}). 
Do not forget to load the library using the function \code{library} before testing following examples.
<<>>=
library(fitdistrplus)
@

\tableofcontents

\pagebreak

\section{Specification of a distribution from non-censored continuous data}
%si on enlève l'étoile on a des numéros de section

\subsection{Graphical display of the observed distribution}

First of all, the observed distribution may be plotted using the function \code{plotdist}.
%%% R code
<<fig=TRUE, eps=FALSE>>=
x1<-c(6.4,13.3,4.1,1.3,14.1,10.6,
9.9,9.6,15.3,22.1,13.4,13.2,
8.4,6.3,8.9,5.2,10.9,14.4)
plotdist(x1)
@
%%%

\subsection{Characterization of the observed distribution}

Descriptive parameters of the empirical distribution may be computed using the function \code{descdist}.
This function will also provide by default a skewness-kurtosis plot which may help you to select 
which distribution(s) to fit among the potential candidates. 
%%% R code
<<fig=TRUE>>=
descdist(x1)
@
%%%

In order to take into account the uncertainty of the estimated values of kurtosis and skewness, 
the data set may be boostrapped by fixing the argument \code{boot} to an integer above 10 in \code{descdist}. 
\code{boot} values of 
skewness and kurtosis corresponding to the boot nonparametric bootstrap samples are then computed and reported in blue color 
on the skewness-kurtosis plot.
%%% R code
<<fig=TRUE, eps=FALSE>>=
descdist(x1,boot=1000)
@
%%%

\subsection{Fitting of a distribution}

One or more parametric distributions may then be fitted to the data set, one at a time, using the
function \code{fitdist}. This function uses the maximum likelihood
method if the argument \code{method="mle"} (or if it is omitted) 
or the matching moments method if the argument \code{method="mom"}.
When fitting continuous\footnote{The reference book on continuous distribution
is \cite{kotz}.} distributions, 
Kolmogorov-Smirnov and Anderson-Darling statistics are computed and corresponding tests are performed when possible. 
Even if less appropriate for continuous distributions, the Chi-squared statistic is
also computed when possible. For this calculation, cells are defined by the argument \code{chisqbreaks}
or automatically defined from the data set and from the argument \code{meancount} (the approximate mean count per cell)
which is  fixed to $(4n)^{2/5}$ if omitted (with n the length of the data set).  
For more details, see the help of the function \code{fitdist}.
Four goodness of fit plots are also provided.

Below is the result of a fit of a gamma distribution by maximum likelihood.
%%% R code
<<fig=TRUE, eps=FALSE>>=
f1g<-fitdist(x1,"gamma")
plot(f1g)
summary(f1g)
@
%%%

Below is the result of another fit of the same distribution by matching moments.
%%% R code
<<>>=
f1gbis<-fitdist(x1,"gamma",method="mom")
summary(f1gbis)
@ 
%%%

As can be seen in this returned summary, the automatic definition of the cells required to calculate the Chi-squared
statistic does not give theoretical counts large enough to validate the use of the test in this example.
It is often the case for small
data sets. The observed and theoretical counts may be printed as below :
%%% R code
<<>>=
f1g$chisqtable
@
%%%

Below is the fit of a lognormal distribution.
%%% R code
<<fig=TRUE, eps=FALSE>>=
f1l<-fitdist(x1,"lnorm")
plot(f1l)
summary(f1l)
@
%%%

Below is the fit of a normal distribution.
%%% R code
<<fig=TRUE, eps=FALSE>>=
f1n<-fitdist(x1,"norm")
plot(f1n)
summary(f1n)
@
%%%

Below is the fit of a Weibull distribution.
%%% R code
<<fig=TRUE, eps=FALSE>>=
f1w<-fitdist(x1,"weibull")
plot(f1w)
summary(f1w)
@
%%%

The values of the Anderson-Darling statistic (or another result of the fit: see the help of \code{fitdist} for details)
for the different fittings may be extracted and compared to help the selection of a distribution :
%%% R code
<<>>=
anderson<-list(lnorm=f1l$ad,gamma=f1g$ad,norm=f1n$ad,weibull=f1w$ad)
anderson
@
%%%

For some distributions (see the help of \code{fitdist} for details), it is necessary to specify initial 
values for the distribution parameters in the argument \code{start} when using the maximum likelihood method.
\code{start} must be a named list
of parameters initial values. The names of
the parameters in \code{start} must correspond exactly to their definition in R or to their 
definition in a previous R code.
The function \code{plotdist} may help to find correct initial values for 
the distribution parameters in non trivial cases, by an manual iterative use if necessary.

For example, below is the definition of the Gumbel distribution (also named extreme value distribution) and 
a first plot of the data set with the Gumbel distribution with arbitrary values for parameters.
%%% R code
<<fig=TRUE, eps=FALSE>>=
dgumbel<-function(x,a,b) 1/b*exp((a-x)/b)*exp(-exp((a-x)/b))
pgumbel<-function(q,a,b) exp(-exp((a-q)/b))
qgumbel<-function(p,a,b) a-b*log(-log(p))
plotdist(x1,"gumbel",para=list(a=3,b=2))
@
%%%

The same data set may be plotted with a Gumbel distribution with modified values for parameters.
%%% R code
<<fig=TRUE, eps=FALSE>>=
plotdist(x1,"gumbel",para=list(a=10,b=5))
@
%%%

And a Gumbel distribution may be fitted to data with these values for initial parameter values.
%%% R code
<<fig=TRUE, eps=FALSE>>=
fgu<-fitdist(x1,"gumbel",start=list(a=10,b=5))
plot(fgu)
summary(fgu)
@
%%%

\subsection{Simulation of the uncertainty by boostrap}

The uncertainty in the parameters of the fitted distribution may be simulated by parametric or nonparametric
boostrap using the function \code{boodist}. This function returns the boostrapped values of parameters which
may be plotted to visualize the bootstrap region. It also calculates the 95 percent confidence intervals for each 
parameter from the 2.5 and 97.5 percentiles
of the boostrap values of each parameter (see the help of the function \code{bootdist} for details).

Below is an example of the use of this function with the previous fit of the gamma distribution. 
%%% R code
<<fig=TRUE, eps=FALSE>>=
b1g<-bootdist(f1g)
plot(b1g)
summary(b1g)
@
%%%

\section{Specification of a distribution from non-censored discrete data}

A discrete data set may be considered as a continuous one for example for a large data set
from a binomial distribution converging to a normal one. A discrete plot of the distribution 
may also be provided, fixing the argument discrete of the function \code{plotdist} to \code{TRUE}.

%%% R code
<<fig=TRUE, eps=FALSE>>=
x2<-rnbinom(n=100,size=2,prob=0.3)
plotdist(x2,discrete=TRUE)
@
%%%

As for continuous distributions, descriptive parameters of the empirical distribution 
may be computed using the function \code{descdist} which also provides a skewness-kurtosis plot 
which may help you to choose which distribution(s) to fit.

%%% R code
<<fig=TRUE, eps=FALSE>>=
descdist(x2,discrete=T)
@
%%%

As for continuous distributions, one or more parametric distributions may then be fitted 
to the data set by maximum likelihood or matching moments.

Below is the result of the fit of a Poisson distribution with the bootstrap simulations.
%%% R code
<<fig=TRUE, eps=FALSE>>=
f2p<-fitdist(x2,"pois")
plot(f2p)
summary(f2p)
b2p<-bootdist(f2p)
summary(b2p)
@
%%%

Below is the result of the fit of a negative binomial distribution with the boostrap simulations.
%%% R code
<<fig=TRUE, eps=FALSE>>=
f2n<-fitdist(x2,"nbinom")
plot(f2n)
summary(f2n)
b2n<-bootdist(f2n)
summary(b2n)
@
%%%

\section{Specification of a distribution from censored data}
%si on enlève l'étoile on a des numéros de section
Censored data may contain left censored, right censored and interval censored values, 
with several lower and upper bounds. Data must be coded into a dataframe with two columns,
respectively named \code{left} 
and \code{right}, describing each observed value as an interval.
The \code{left} column contains either \code{NA} for left censored observations,
the left bound of the interval for interval censored observations,
or the observed value for non-censored observations.
The \code{right} column contains either \code{NA} for right censored observations,
the right bound of the interval for interval censored observations,
or the observed value for non-censored observations.

\subsection{Graphical display of the observed distribution}

First of all, the observed distribution may be plotted using the function \code{plotdistcens}.
Data are reported directly as segments for interval, left and right censored data, 
and as points for non-censored data. For more details, see the help of the function \code{plotdistcens}.

%%% R code
<<fig=TRUE, eps=FALSE>>=
d1<-data.frame(
left=c(1.73,1.51,0.77,1.96,1.96,-1.4,-1.4,NA,-0.11,0.55,0.41,
   2.56,NA,-0.53,0.63,-1.4,-1.4,-1.4,NA,0.13),
right=c(1.73,1.51,0.77,1.96,1.96,0,-0.7,-1.4,-0.11,0.55,0.41,
   2.56,-1.4,-0.53,0.63,0,-0.7,NA,-1.4,0.13))
plotdistcens(d1)
@
%%%

When left or right NA-values correspond to finite value (for example 0 for
left NA-values of positive data), the arguments \code{leftNA} (or \code{rightNA})
must be affected to this finite value to ensure a correct plot of left (or right) censored
observations, as in the example below.

%%% R code
<<fig=TRUE, eps=FALSE>>=
d2<-data.frame(left=10^(d1$left),right=10^(d1$right))
plotdistcens(d2,leftNA=0)
@
%%%

It is also possible to fix \code{rightNA} or \code{leftNA} to a realistic 
extreme value, even if not exactly known, to obtain a reasonable global ranking of 
observations, as in the example below for the first dataset.
%%% R code
<<fig=TRUE, eps=FALSE>>=
plotdistcens(d1,rightNA=3)
@
%%%

\subsection{Fitting of a distribution}

One or more parametric distributions may then be fitted to the censored data set, one at a time, using the
function \code{fitdistcens}. This function always uses the maximum likelihood
method.  For more details, see the help of the function \code{fitdistcens}.
Only one goodness of fit plot is provided for censored data, in cumulative frequencies.
The uncertainty in the parameters of the fitted distribution may be simulated by nonparametric
boostrap only, using the function \code{boodistcens}.

Below is the result of a fit of a Weibull distribution by maximum likelihood and the results of the 
corresponding boostrap simulations.
%%% R code
<<fig=TRUE, eps=FALSE>>=
f2w<-fitdistcens(d2,"weibull")
summary(f2w)
plot(f2w,leftNA=0)
@
%%%

%%% R code
<<fig=TRUE, eps=FALSE>>=
b2w<-bootdistcens(f2w)
summary(b2w)
plot(b2w)
@
%%%

Goodness of fit statistics are not computed for fit on censored data, so the quality of fit 
may only be estimated from the loglikelihood and the goodness of fit plot.

Below is the fit of a lognormal distribution to the same censored data set.
%%% R code
<<fig=TRUE, eps=FALSE>>=
f2l<-fitdistcens(d2,"lnorm")
summary(f2l)
plot(f2l,leftNA=0)
@
%%%

Below is the fit of an exponential distribution.
%%% R code
<<fig=TRUE, eps=FALSE>>=
f2e<-fitdistcens(d2,"exp")
summary(f2e)
plot(f2e,leftNA=0)
@
%%%

As with \code{fitdist}, for some distributions (see the help of \code{fitdistcens} for details), 
it is necessary to specify initial 
values for the distribution parameters in the argument \texttt{start}.
\code{start} must be a named list
of parameters initial values. The names of
the parameters in \code{start} must correspond exactly to their definition in R or to their 
definition in a previous R code.
The function \code{plotdistcens} may help to find correct initial values for 
the distribution parameters in non trivial cases, by an manual iterative use if necessary, as explained 
previously for non-censored continuous data.

%%%%%%%%%%%
%  Bibiliography.
%%%%%%%%%%%
\newpage
\addcontentsline{toc}{section}{Bibliography}
\bibliographystyle{agsm}  
\bibliography{intro2fitdistrplus}

\end{document}

### 2009-07-18 R/bootdistcens.R

@@ -1,3 +1,27 @@
+#############################################################################
+#   Copyright (c) 2009 Marie Laure Delignette-Muller, Regis Pouillot, Jean-Baptiste Denis                                                                                                  
+#                                                                                                                                                                        
+#   This program is free software; you can redistribute it and/or modify                                               
+#   it under the terms of the GNU General Public License as published by                                         
+#   the Free Software Foundation; either version 2 of the License, or                                                   
+#   (at your option) any later version.                                                                                                            
+#                                                                                                                                                                         
+#   This program is distributed in the hope that it will be useful,                                                             
+#   but WITHOUT ANY WARRANTY; without even the implied warranty of                                          
+#   MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the                                 
+#   GNU General Public License for more details.                                                                                    
+#                                                                                                                                                                         
+#   You should have received a copy of the GNU General Public License                                           
+#   along with this program; if not, write to the                                                                                           
+#   Free Software Foundation, Inc.,                                                                                                              
+#   59 Temple Place, Suite 330, Boston, MA 02111-1307, USA                                                             
+#                                                                                                                                                                         
+#############################################################################
+### bootstrap in fitdistrplus with censored data
+###
+###			R functions
+### 
+
 bootdistcens<-function (f, niter=999)
 { 
     if (niter<10) 

### 2009-07-18 R/descdist.R

@@ -1,3 +1,27 @@
+#############################################################################
+#   Copyright (c) 2009 Marie Laure Delignette-Muller, Regis Pouillot, Jean-Baptiste Denis                                                                                                  
+#                                                                                                                                                                        
+#   This program is free software; you can redistribute it and/or modify                                               
+#   it under the terms of the GNU General Public License as published by                                         
+#   the Free Software Foundation; either version 2 of the License, or                                                   
+#   (at your option) any later version.                                                                                                            
+#                                                                                                                                                                         
+#   This program is distributed in the hope that it will be useful,                                                             
+#   but WITHOUT ANY WARRANTY; without even the implied warranty of                                          
+#   MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the                                 
+#   GNU General Public License for more details.                                                                                    
+#                                                                                                                                                                         
+#   You should have received a copy of the GNU General Public License                                           
+#   along with this program; if not, write to the                                                                                           
+#   Free Software Foundation, Inc.,                                                                                                              
+#   59 Temple Place, Suite 330, Boston, MA 02111-1307, USA                                                             
+#                                                                                                                                                                         
+#############################################################################
+### Description of an empirical distribution
+###
+###			R functions
+### 
+
 descdist <- function(data,discrete=FALSE,boot=NULL,graph=TRUE)
 {
     #if(is.mcnode(data)) data <- as.vector(data)

### 2009-07-18 R/mledist.R

@@ -1,3 +1,27 @@
+#############################################################################
+#   Copyright (c) 2009 Marie Laure Delignette-Muller, Regis Pouillot, Jean-Baptiste Denis                                                                                                  
+#                                                                                                                                                                        
+#   This program is free software; you can redistribute it and/or modify                                               
+#   it under the terms of the GNU General Public License as published by                                         
+#   the Free Software Foundation; either version 2 of the License, or                                                   
+#   (at your option) any later version.                                                                                                            
+#                                                                                                                                                                         
+#   This program is distributed in the hope that it will be useful,                                                             
+#   but WITHOUT ANY WARRANTY; without even the implied warranty of                                          
+#   MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the                                 
+#   GNU General Public License for more details.                                                                                    
+#                                                                                                                                                                         
+#   You should have received a copy of the GNU General Public License                                           
+#   along with this program; if not, write to the                                                                                           
+#   Free Software Foundation, Inc.,                                                                                                              
+#   59 Temple Place, Suite 330, Boston, MA 02111-1307, USA                                                             
+#                                                                                                                                                                         
+#############################################################################
+### maximum likelihood estimation for non-censored data
+###
+###			R functions
+### 
+
 mledist<-function (data, distr, start=NULL) 
 {
     if (!is.character(distr)) distname<-substring(as.character(match.call()$distr),2)

### 2009-07-18 R/plotdist.R

@@ -1,3 +1,27 @@
+#############################################################################
+#   Copyright (c) 2009 Marie Laure Delignette-Muller, Regis Pouillot, Jean-Baptiste Denis                                                                                                  
+#                                                                                                                                                                        
+#   This program is free software; you can redistribute it and/or modify                                               
+#   it under the terms of the GNU General Public License as published by                                         
+#   the Free Software Foundation; either version 2 of the License, or                                                   
+#   (at your option) any later version.                                                                                                            
+#                                                                                                                                                                         
+#   This program is distributed in the hope that it will be useful,                                                             
+#   but WITHOUT ANY WARRANTY; without even the implied warranty of                                          
+#   MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the                                 
+#   GNU General Public License for more details.                                                                                    
+#                                                                                                                                                                         
+#   You should have received a copy of the GNU General Public License                                           
+#   along with this program; if not, write to the                                                                                           
+#   Free Software Foundation, Inc.,                                                                                                              
+#   59 Temple Place, Suite 330, Boston, MA 02111-1307, USA                                                             
+#                                                                                                                                                                         
+#############################################################################
+### plot functions for non-censored data
+###
+###			R functions
+### 
+
 plotdist <- function(data,distr,para,breaks="default",discrete=FALSE,...){
     def.par <- par(no.readonly = TRUE)
     if (missing(data) || !is.vector(data,mode="numeric"))

### 2009-07-18 R/plotdistcens.R

@@ -1,3 +1,27 @@
+#############################################################################
+#   Copyright (c) 2009 Marie Laure Delignette-Muller, Regis Pouillot, Jean-Baptiste Denis                                                                                                  
+#                                                                                                                                                                        
+#   This program is free software; you can redistribute it and/or modify                                               
+#   it under the terms of the GNU General Public License as published by                                         
+#   the Free Software Foundation; either version 2 of the License, or                                                   
+#   (at your option) any later version.                                                                                                            
+#                                                                                                                                                                         
+#   This program is distributed in the hope that it will be useful,                                                             
+#   but WITHOUT ANY WARRANTY; without even the implied warranty of                                          
+#   MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the                                 
+#   GNU General Public License for more details.                                                                                    
+#                                                                                                                                                                         
+#   You should have received a copy of the GNU General Public License                                           
+#   along with this program; if not, write to the                                                                                           
+#   Free Software Foundation, Inc.,                                                                                                              
+#   59 Temple Place, Suite 330, Boston, MA 02111-1307, USA                                                             
+#                                                                                                                                                                         
+#############################################################################
+### plot functions for censored data
+###
+###			R functions
+### 
+
 plotdistcens <- function(censdata,distr,para,leftNA=-Inf,rightNA=Inf,...){
     def.par <- par(no.readonly = TRUE)
     if (missing(censdata) ||

### 2009-08-21 inst/doc/intro2fitdistrplus.Rnw
add the vignette and test the makefile with citations.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% fitdistrplus
+%%% Fit parametric distributions on non-censored or censored data with fitdistrplus
 %%%
@@
 \documentclass[a4paper]{article}
-
-% sweave commands for vignette
-%\VignetteIndexEntry{Fit parametric distributions on non-censored or censored data}
-%\VignettePackage{fitdistrplus}
-%\VignetteKeyword{distribution}
-
-%geometry
 \usepackage[a4paper,textwidth=18cm,textheight=27cm]{geometry}
 
-%sweave
-\usepackage{Sweave}
-
-%footnote
-\usepackage[stable]{footmisc}
-\usepackage{perpage}
-\MakePerPage[1]{footnote}
-
-%citation
-\usepackage{natbib}
-
-%reference hypertext
-\usepackage[pagebackref=true, hyperfootnotes=false]{hyperref}
-
-%title
 \title{Use of the library \texttt{fitdistrplus} to specify a distribution from 
 non-censored or censored data}
-
-%authors
-\author{Marie Laure Delignette-Muller, R\'egis Pouillot, Jean-Baptiste Denis and Christophe Dutang}
-
-% macros
-\newcommand\HRule{\noindent\rule{\linewidth}{1pt}}
-\newcommand{\II}{\mbox{\large 1\hskip -0,353em 1}}
-\newcommand{\abs}[1]{\lvert#1\rvert}
-\newcommand{\norm}[1]{\lVert#1\rVert}
-\newcommand{\ligne}{\rule[2mm]{.3\textwidth}{0,5mm}\\}
-\newcommand{\pkg}{\textbf}
-\newcommand{\sigle}{\textsc}
-\newcommand{\code}{\texttt}
-\newcommand{\soft}{\textsf}
-\newcommand{\myskip}{\vspace{\parskip}}
-\newcommand{\txtm}[1]{\textrm{~~#1~~}}
-\newcommand{\expo}{\textsuperscript}
-\newcommand{\ind}{1\!\!1}
-\newcommand{\mat}[1]{\mathbf{#1}}
-\newcommand{\blank}{ \clearpage{\pagestyle{empty}\cleardoublepage} }
-\newcommand{\card}{\textrm{Card}}
+\author{Marie Laure Delignette-Muller, Régis Pouillot , Jean-Baptiste Denis and Christophe Dutang}
 
@@
 
-Here you will find some easy examples of use of the functions of the library \pkg{fitdistrplus}. 
+Here you will find some easy examples of use of the functions of the library \texttt{fitdistrplus}. 
 The aim is to show you by examples how to use these functions to help you to specify a parametric distribution 
 from data corresponding to a random sample drawn from a theoretical distribution that you want to describe. 
-For details, see the documentation of each function, using the R help command (ex.: \code{?fitdist}). 
-Do not forget to load the library using the function \code{library} before testing following examples.
+For details, see the documentation of each function, using the R help command (ex.: \texttt{?fitdist}). 
+Do not forget to load the library using the function \texttt{library} before testing following examples.
 <<>>=
@@
 
-First of all, the observed distribution may be plotted using the function \code{plotdist}.
-%%% R code
-<<fig=TRUE, eps=FALSE>>=
+ First of all, the observed distribution may be plotted using the function \texttt{plotdist}.
+
+<<fig=TRUE>>=
 x1<-c(6.4,13.3,4.1,1.3,14.1,10.6,
@@
 @
-%%%
 
@@
 
-Descriptive parameters of the empirical distribution may be computed using the function \code{descdist}.
+Descriptive parameters of the empirical distribution may be computed using the function \texttt{descdist}.
 This function will also provide by default a skewness-kurtosis plot which may help you to select 
 which distribution(s) to fit among the potential candidates.
-%%% R code
+
 <<fig=TRUE>>=
@@
 @
-%%%
 
 In order to take into account the uncertainty of the estimated values of kurtosis and skewness, 
-the data set may be boostrapped by fixing the argument \code{boot} to an integer above 10 in \code{descdist}. 
-\code{boot} values of 
+the data set may be boostrapped by fixing the argument \texttt{boot} to an integer above 10 in \texttt{descdist}. 
+\texttt{boot} values of 
 skewness and kurtosis corresponding to the boot nonparametric bootstrap samples are then computed and reported in blue color 
 on the skewness-kurtosis plot.
-%%% R code
-<<fig=TRUE, eps=FALSE>>=
+
+<<fig=TRUE>>=
 descdist(x1,boot=1000)
 @
-%%%
 
@@
 One or more parametric distributions may then be fitted to the data set, one at a time, using the
-function \code{fitdist}. This function uses the maximum likelihood
-method if the argument \code{method="mle"} (or if it is omitted) 
-or the matching moments method if the argument \code{method="mom"}.
-When fitting continuous\footnote{The reference book on continuous distribution
-is \cite{kotz}.} distributions, 
+fonction \texttt{fitdist}. This function uses the maximum likelihood
+method if the argument \texttt{method="mle"} (or if it is omitted) 
+or the matching moments method if the argument \texttt{method="mom"}.
+When fitting continuous distributions, 
 Kolmogorov-Smirnov and Anderson-Darling statistics are computed and corresponding tests are performed when possible. 
 Even if less appropriate for continuous distributions, the Chi-squared statistic is
-also computed when possible. For this calculation, cells are defined by the argument \code{chisqbreaks}
-or automatically defined from the data set and from the argument \code{meancount} (the approximate mean count per cell)
+also computed when possible. For this calculation, cells are defined by the argument \texttt{chisqbreaks}
+or automatically defined from the data set and from the argument \texttt{meancount} (the approximate mean count per cell)
 which is  fixed to $(4n)^{2/5}$ if omitted (with n the length of the data set).  
-For more details, see the help of the function \code{fitdist}.
+For more details, see the help of the function \texttt{fitdist}.
 Four goodness of fit plots are also provided.
@@
 Below is the result of a fit of a gamma distribution by maximum likelihood.
-%%% R code
-<<fig=TRUE, eps=FALSE>>=
+<<fig=TRUE>>=
 f1g<-fitdist(x1,"gamma")
@@
 @
-%%%
 
 Below is the result of another fit of the same distribution by matching moments.
-%%% R code
 <<>>=
@@
 data sets. The observed and theoretical counts may be printed as below :
-%%% R code
 <<>>=
@@
 @
-%%%
 
 Below is the fit of a lognormal distribution.
-%%% R code
-<<fig=TRUE, eps=FALSE>>=
+<<fig=TRUE>>=
 f1l<-fitdist(x1,"lnorm")
@@
 @
-%%%
 
 Below is the fit of a normal distribution.
-%%% R code
-<<fig=TRUE, eps=FALSE>>=
+<<fig=TRUE>>=
 f1n<-fitdist(x1,"norm")
@@
 @
-%%%
 
 Below is the fit of a Weibull distribution.
-%%% R code
-<<fig=TRUE, eps=FALSE>>=
+<<fig=TRUE>>=
 f1w<-fitdist(x1,"weibull")
@@
 @
-%%%
 
-The values of the Anderson-Darling statistic (or another result of the fit: see the help of \code{fitdist} for details)
+The values of the Anderson-Darling statistic (or another result of the fit: see the help of \texttt{fitdist} for details)
  for the different fittings may be extracted and compared to help the selection of a distribution :
-%%% R code
+
 <<>>=
@@
 @
-%%%
 
-For some distributions (see the help of \code{fitdist} for details), it is necessary to specify initial 
-values for the distribution parameters in the argument \code{start} when using the maximum likelihood method.
-\code{start} must be a named list
+For some distributions (see the help of \texttt{fitdist} for details), it is necessary to specify initial 
+values for the distribution parameters in the argument \texttt{start} when using the maximum likelihood method.
+\texttt{start} must be a named list
 of parameters initial values. The names of
-the parameters in \code{start} must correspond exactly to their definition in R or to their 
+the parameters in \texttt{start} must correspond exactly to their definition in R or to their 
 definition in a previous R code.
-The function \code{plotdist} may help to find correct initial values for 
+ The function \texttt{plotdist} may help to find correct initial values for 
 the distribution parameters in non trivial cases, by an manual iterative use if necessary.
@@
 a first plot of the data set with the Gumbel distribution with arbitrary values for parameters.
-%%% R code
-<<fig=TRUE, eps=FALSE>>=
+<<fig=TRUE>>=
 dgumbel<-function(x,a,b) 1/b*exp((a-x)/b)*exp(-exp((a-x)/b))
@@
 @
-%%%
 
 The same data set may be plotted with a Gumbel distribution with modified values for parameters.
-%%% R code
-<<fig=TRUE, eps=FALSE>>=
+<<fig=TRUE>>=
 plotdist(x1,"gumbel",para=list(a=10,b=5))
 @
-%%%
 
 And a Gumbel distribution may be fitted to data with these values for initial parameter values.
-%%% R code
-<<fig=TRUE, eps=FALSE>>=
+<<fig=TRUE>>=
 fgu<-fitdist(x1,"gumbel",start=list(a=10,b=5))
@@
 The uncertainty in the parameters of the fitted distribution may be simulated by parametric or nonparametric
-boostrap using the function \code{boodist}. This function returns the boostrapped values of parameters which
+boostrap using the function \texttt{boodist}. This function returns the boostrapped values of parameters which
 may be plotted to visualize the bootstrap region. It also calculates the 95 percent confidence intervals for each 
 parameter from the 2.5 and 97.5 percentiles
-of the boostrap values of each parameter (see the help of the function \code{bootdist} for details).
+of the boostrap values of each parameter (see the help of the function \texttt{bootdist} for details).
 
 Below is an example of the use of this function with the previous fit of the gamma distribution. 
-%%% R code
-<<fig=TRUE, eps=FALSE>>=
+<<fig=TRUE>>=
 b1g<-bootdist(f1g)
@@
  from a binomial distribution converging to a normal one. A discrete plot of the distribution 
-may also be provided, fixing the argument discrete of the function \code{plotdist} to \code{TRUE}.
+ may also be provided, fixing the argument discrete of the function \texttt{plotdist} to \texttt{TRUE}.
 
-%%% R code
-<<fig=TRUE, eps=FALSE>>=
+<<fig=TRUE>>=
 x2<-rnbinom(n=100,size=2,prob=0.3)
@@
 @
-%%%
 
 As for continuous distributions, descriptive parameters of the empirical distribution 
-may be computed using the function \code{descdist} which also provides a skewness-kurtosis plot 
+may be computed using the function \texttt{descdist} which also provides a skewness-kurtosis plot 
 which may help you to choose which distribution(s) to fit.
 
-%%% R code
-<<fig=TRUE, eps=FALSE>>=
+<<fig=TRUE>>=
 descdist(x2,discrete=T)
 @
-%%%
 
@@
 Below is the result of the fit of a Poisson distribution with the bootstrap simulations.
-%%% R code
-<<fig=TRUE, eps=FALSE>>=
+<<fig=TRUE>>=
 f2p<-fitdist(x2,"pois")
@@
 @
-%%%
 
 Below is the result of the fit of a negative binomial distribution with the boostrap simulations.
-%%% R code
-<<fig=TRUE, eps=FALSE>>=
+<<fig=TRUE>>=
 f2n<-fitdist(x2,"nbinom")
@@
 with several lower and upper bounds. Data must be coded into a dataframe with two columns,
-respectively named \code{left} 
-and \code{right}, describing each observed value as an interval.
-The \code{left} column contains either \code{NA} for left censored observations,
+ respectively named \texttt{left} 
+ and \texttt{right}, describing each observed value as an interval.
+ The \texttt{left} column contains either \texttt{NA} for left censored observations,
  the left bound of the interval for interval censored observations,
  or the observed value for non-censored observations.
-The \code{right} column contains either \code{NA} for right censored observations,
+ The \texttt{right} column contains either \texttt{NA} for right censored observations,
  the right bound of the interval for interval censored observations,
@@
 
-First of all, the observed distribution may be plotted using the function \code{plotdistcens}.
+First of all, the observed distribution may be plotted using the function \texttt{plotdistcens}.
 Data are reported directly as segments for interval, left and right censored data, 
-and as points for non-censored data. For more details, see the help of the function \code{plotdistcens}.
+and as points for non-censored data. For more details, see the help of the function \texttt{plotdistcens}.
 
-%%% R code
-<<fig=TRUE, eps=FALSE>>=
+<<fig=TRUE>>=
 d1<-data.frame(
@@
 @
-%%%
 
 When left or right NA-values correspond to finite value (for example 0 for
-left NA-values of positive data), the arguments \code{leftNA} (or \code{rightNA})
+left NA-values of positive data), the arguments \texttt{leftNA} (or \texttt{rightNA})
 must be affected to this finite value to ensure a correct plot of left (or right) censored
@@
 
-%%% R code
-<<fig=TRUE, eps=FALSE>>=
+<<fig=TRUE>>=
 d2<-data.frame(left=10^(d1$left),right=10^(d1$right))
@@
 @
-%%%
 
-It is also possible to fix \code{rightNA} or \code{leftNA} to a realistic 
+It is also possible to fix \texttt{rightNA} or \texttt{leftNA} to a realistic 
 extreme value, even if not exactly known, to obtain a reasonable global ranking of 
 observations, as in the example below for the first dataset.
-%%% R code
-<<fig=TRUE, eps=FALSE>>=
+<<fig=TRUE>>=
 plotdistcens(d1,rightNA=3)
 @
-%%%
 
@@
 One or more parametric distributions may then be fitted to the censored data set, one at a time, using the
-function \code{fitdistcens}. This function always uses the maximum likelihood
-method.  For more details, see the help of the function \code{fitdistcens}.
+fonction \texttt{fitdistcens}. This function always uses the maximum likelihood
+method.  For more details, see the help of the function \texttt{fitdistcens}.
 Only one goodness of fit plot is provided for censored data, in cumulative frequencies.
 The uncertainty in the parameters of the fitted distribution may be simulated by nonparametric
-boostrap only, using the function \code{boodistcens}.
+boostrap only, using the function \texttt{boodistcens}.
 
@@
 corresponding boostrap simulations.
-%%% R code
-<<fig=TRUE, eps=FALSE>>=
+<<fig=TRUE>>=
 f2w<-fitdistcens(d2,"weibull")
@@
 @
-%%%
 
-%%% R code
-<<fig=TRUE, eps=FALSE>>=
+<<fig=TRUE>>=
 b2w<-bootdistcens(f2w)
@@
 Below is the fit of a lognormal distribution to the same censored data set.
-%%% R code
-<<fig=TRUE, eps=FALSE>>=
+<<fig=TRUE>>=
 f2l<-fitdistcens(d2,"lnorm")
@@
 @
-%%%
 
 Below is the fit of an exponential distribution.
-%%% R code
-<<fig=TRUE, eps=FALSE>>=
+<<fig=TRUE>>=
 f2e<-fitdistcens(d2,"exp")
@@
 @
-%%%
 
-As with \code{fitdist}, for some distributions (see the help of \code{fitdistcens} for details), 
+As with \texttt{fitdist}, for some distributions (see the help of \texttt{fitdistcens} for details), 
 it is necessary to specify initial 
 values for the distribution parameters in the argument \texttt{start}.
-\code{start} must be a named list
+\texttt{start} must be a named list
 of parameters initial values. The names of
-the parameters in \code{start} must correspond exactly to their definition in R or to their 
+the parameters in \texttt{start} must correspond exactly to their definition in R or to their 
 definition in a previous R code.
-The function \code{plotdistcens} may help to find correct initial values for 
+The function \texttt{plotdistcens} may help to find correct initial values for 
 the distribution parameters in non trivial cases, by an manual iterative use if necessary, as explained 
@@
 
-%%%%%%%%%%%
-%  Bibiliography.
-%%%%%%%%%%%
-\newpage
-\addcontentsline{toc}{section}{Bibliography}
-\bibliographystyle{agsm}  
-\bibliography{intro2fitdistrplus}
-
 \end{document}

### 2009-08-21 R/bootdist.R

@@ -19,11 +19,11 @@
 #############################################################################
 ### bootstrap in fitdistrplus
 ###
-###			R functions
+###         R functions
 ### 
 
 
-bootdist<-function (f, bootmethod="param", niter=999)
+bootdist<-function (f, bootmethod="param", niter=1001)
 { 
     if (niter<10) 
         stop("niter must be an integer above 10")
@@ -119,15 +119,15 @@ plot.bootdist <- function(x,...){
         stop("Use only with 'bootdist' objects")
     if (dim(x$estim)[2]==1) {
         stripchart(x$estim,method="jitter",
-            xlab="Scatterplot of boostrapped values of the parameter",...)
+            xlab="Boostrapped values of the parameter",...)
     }
     else {
         if (dim(x$estim)[2]==2)
             plot(x$estim,
-            main="Scatterplot of boostrapped values of parameters",...)
+            main="Boostrapped values of parameters",...)
         else 
             plot(x$estim,
-            main="Scatterplots of boostrapped values of parameters",...)
+            main="Boostrapped values of parameters",...)
     }
 }
 
@@ -137,9 +137,9 @@ summary.bootdist <- function(object,...){
     op<-options()
     options(digits=3)
     if (object$method=="param") 
-        cat("Parametric bootstrap medians and 95% CI \n")
+        cat("Parametric bootstrap medians and 95% percentile CI \n")
     else
-       cat("Nonparametric bootstrap medians and 95% CI \n")
+       cat("Nonparametric bootstrap medians and 95% percentile CI \n")
     print(object$CI)
     
      if (!is.null(object$converg)) { 

### 2009-08-21 R/fitdistcens.R

@@ -19,10 +19,10 @@
 #############################################################################
 ### fit parametric distributions for censored data
 ###
-###			R functions
+###         R functions
 ### 
 
-fitdistcens<-function (censdata, distr, start) 
+fitdistcens<-function (censdata, distr, start,...) 
 {
     if (!is.character(distr)) distname<-substring(as.character(match.call()$distr),2)
     else distname<-distr
@@ -34,9 +34,9 @@ fitdistcens<-function (censdata, distr, start)
         stop(paste("The ",pdistname," function must be defined"))
     # MLE fit with mledistcens 
     if (missing(start))
-        mle<-mledistcens(censdata,distname) 
+        mle<-mledistcens(censdata,distname,...) 
     else 
-    mle<-mledistcens(censdata,distname,start)
+    mle<-mledistcens(censdata,distname,start,...)
     if (mle$convergence>0) 
         stop("the function mle failed to estimate the parameters, 
         with the error code ",mle$convergence) 
@@ -45,8 +45,12 @@ fitdistcens<-function (censdata, distr, start)
     sd<-sqrt(diag(varcovar))
     correl<-cov2cor(varcovar)
     loglik<-mle$loglik
+    n<-nrow(censdata)
+    npar<-length(estimate)
+    aic<--2*loglik+2*npar
+    bic<--2*loglik+log(n)*npar
          
-    return(structure(list(estimate = estimate, sd = sd, cor = correl, loglik = loglik, 
+    return(structure(list(estimate = estimate, sd = sd, cor = correl, loglik = loglik,aic=aic,bic=bic, 
         censdata=censdata, distname=distname), class = "fitdistcens"))
         
 }
@@ -82,7 +86,9 @@ summary.fitdistcens <- function(object,...){
     "' BY MAXIMUM LIKELIHOOD ON CENSORED DATA \n")
     cat("PARAMETERS\n")
     print(cbind.data.frame("estimate" = object$estimate, "Std. Error" = object$sd))
-    cat("Loglikelihood: ",object$loglik,"\n")
+    cat("Loglikelihood: ",object$loglik,"  ")
+    cat("AIC: ",object$aic,"  ")
+    cat("BIC: ",object$bic,"\n")
     if (length(object$estimate) > 1) {
         cat("Correlation matrix:\n")
         print(object$cor)

### 2009-08-21 R/plotdist.R

@@ -19,7 +19,7 @@
 #############################################################################
 ### plot functions for non-censored data
 ###
-###			R functions
+###         R functions
 ### 
 
 plotdist <- function(data,distr,para,breaks="default",discrete=FALSE,...){
@@ -41,7 +41,7 @@ plotdist <- function(data,distr,para,breaks="default",discrete=FALSE,...){
                 h<-hist(data,freq=FALSE,xlab="data",main=paste("Histogram"),...)
             else 
                 h<-hist(data,freq=FALSE,xlab="data",main=paste("Histogram"),breaks=breaks,...)
-            plot(s,obsp,main=paste("Cumulative distribution plot"),xlab="data",
+            plot(s,obsp,main=paste("Cumulative distribution"),xlab="data",
             xlim=c(h$breaks[1],h$breaks[length(h$breaks)]),ylab="CDF",pch=4,...)
         }
         else {

### 2009-08-21 R/plotdistcens.R

@@ -19,7 +19,7 @@
 #############################################################################
 ### plot functions for censored data
 ###
-###			R functions
+###         R functions
 ### 
 
 plotdistcens <- function(censdata,distr,para,leftNA=-Inf,rightNA=Inf,...){
@@ -63,7 +63,7 @@ plotdistcens <- function(censdata,distr,para,leftNA=-Inf,rightNA=Inf,...){
     xmax<-xmax+0.3*xrange
     xlim<-c(xmin,xmax)
     plot(c(0,0),c(0,0),type="n",xlim=xlim,ylim=c(0,1),xlab="censored data",
-    ylab="CDF",main="Cumulative distribution plot")
+    ylab="CDF",main="Cumulative distribution")
     # functions to plot one interval or point for each observation for 
     # observation ordered i out of n
     plotlcens<-function(i) {

### 2009-10-04 inst/doc/intro2fitdistrplus.Rnw
fitdistrplus 0.1-2 with some new things (see file NEWS)
correction of minor erros on .Rd files and try to correct the vignette
change mom method to mme method. TODO man page.

@@
 \documentclass[a4paper]{article}
+% sweave commands for vignette
+%\VignetteIndexEntry{Fit parametric distributions on non-censored or censored data}
+%\VignettePackage{fitdistrplus}
+%\VignetteKeyword{distribution}
+
 \usepackage[a4paper,textwidth=18cm,textheight=27cm]{geometry}
@@
 method if the argument \texttt{method="mle"} (or if it is omitted) 
-or the matching moments method if the argument \texttt{method="mom"}.
+or the matching moments estimation if the argument \texttt{method="mme"}.
 When fitting continuous distributions, 
@@
 <<>>=
-f1gbis<-fitdist(x1,"gamma",method="mom")
+f1gbis<-fitdist(x1,"gamma",method="mme")
 summary(f1gbis)

### 2009-10-04 R/fitdist.R

@@ -22,25 +22,31 @@
 ###         R functions
 ### 
 
-fitdist<-function (data, distr, method="mle", start, chisqbreaks, meancount, ...) 
+fitdist<-function (data, distr, method=c("mle", "mme"), start, chisqbreaks, meancount, ...) 
 {
-    if (!is.character(distr)) distname<-substring(as.character(match.call()$distr),2)
-    else distname<-distr
+    if (!is.character(distr)) 
+		distname<-substring(as.character(match.call()$distr),2)
+    else 
+		distname<-distr
     ddistname<-paste("d",distname,sep="")
+	
     if (!exists(ddistname,mode="function"))
         stop(paste("The ",ddistname," function must be defined"))
     pdistname<-paste("p",distname,sep="")
     if (!exists(pdistname,mode="function"))
         stop(paste("The ",pdistname," function must be defined"))
-    if (!is.element(method,c("mle","mom")))
-        stop("method must be affected to 'mle' or 'mom'") 
-    if (!missing(start) & method=="mom")
+    
+	method <- match.arg(method)
+#	if (!is.element(method,c("mle","mom")))
+#        stop("method must be affected to 'mle' or 'mom'") 
+    
+	if (!missing(start) & method=="mme")
         warnings("Starting values for parameters will not be used with matching moments")  
     if (!(is.vector(data) & is.numeric(data) & length(data)>1))
         stop("data must be a numeric vector of length greater than 1")
     n<-length(data)
     # MLE fit with mledist or matching moments fit with momdist
-    if (method=="mom")
+    if (method=="mme")
     {
         estimate <- momdist(data,distname)
         sd <- NULL
@@ -215,7 +221,7 @@ fitdist<-function (data, distr, method="mle", start, chisqbreaks, meancount, ...
 print.fitdist <- function(x,...){
     if (!inherits(x, "fitdist"))
         stop("Use only with 'fitdist' objects")
-    if (x$method=="mom") 
+    if (x$method=="mme") 
         cat("Fitting of the distribution '",x$distname,"' by matching moments \n")
     else
        cat("Fitting of the distribution '",x$distname,"' by maximum likelihood \n")
@@ -242,7 +248,7 @@ summary.fitdist <- function(object,...){
     
     op<-options()
     options(digits=3)
-    if (object$method=="mom") 
+    if (object$method=="mme") 
[…]

### 2009-12-10 R/bootdist.R
Adding of examples to fitdistcens.Rd and various minor changes

@@ -101,8 +101,8 @@ print.bootdist <- function(x,...){
         cat("Parameter values obtained with parametric bootstrap \n")
     else
        cat("Parameter values obtained with nonparametric bootstrap \n")
-    op<-options()
-    options(digits=3)
+    #op<-options()
+    #options(digits=3)
     print(x$estim,...)    
     if (!is.null(x$converg)) { 
         nconverg<-length(x$converg[x$converg==0])
@@ -110,7 +110,7 @@ print.bootdist <- function(x,...){
         cat("Maximum likelihood method converged for ",nconverg," among ",
             length(x$converg)," iterations \n")
     }
-    options(op)
+    #options(op)
 
 }
 
@@ -134,8 +134,8 @@ plot.bootdist <- function(x,...){
 summary.bootdist <- function(object,...){
     if (!inherits(object, "bootdist"))
         stop("Use only with 'bootdist' objects")
-    op<-options()
-    options(digits=3)
+    #op<-options()
+    #options(digits=3)
     if (object$method=="param") 
         cat("Parametric bootstrap medians and 95% percentile CI \n")
     else
@@ -148,5 +148,5 @@ summary.bootdist <- function(object,...){
         cat("Maximum likelihood method converged for ",nconverg," among ",
             length(object$converg)," iterations \n")
     }
-   options(op)
+   #options(op)
 }

### 2009-12-10 R/descdist.R

@@ -19,7 +19,7 @@
 #############################################################################
 ### Description of an empirical distribution
 ###
-###			R functions
+###         R functions
 ### 
 
 descdist <- function(data,discrete=FALSE,boot=NULL,graph=TRUE)
@@ -44,8 +44,8 @@ descdist <- function(data,discrete=FALSE,boot=NULL,graph=TRUE)
     mean=mean(data),sd=sqrt(moment(data,2)),
     skewness=skewness(data),kurtosis=kurtosis(data))
     
-    op<-options()
-    options(digits=3)
+    #op<-options()
+    #options(digits=3)
     cat("summary statistics\n")
     cat("------\n")
     cat("min: ",res$min,"  max: ",res$max,"\n")
@@ -54,7 +54,7 @@ descdist <- function(data,discrete=FALSE,boot=NULL,graph=TRUE)
     cat("sample sd: ",res$sd,"\n")
     cat("sample skewness: ",res$skewness,"\n")
     cat("sample kurtosis: ",res$kurtosis,"\n")
-    options(op)
+    #options(op)
     
     skewdata<-res$skewness
     kurtdata<-res$kurtosis

### 2009-12-10 R/mledist.R

@@ -25,7 +25,7 @@
 ### the mle function of the stat package.
 
 mledist <- function (data, distr, start=NULL, optim.method="default",
-    lower=-Inf, upper=Inf, custom.optim=NULL, ..., echo=FALSE)
+    lower=-Inf, upper=Inf, custom.optim=NULL, ...)
     # data may correspond to a vector for non censored data or to
     # a dataframe of two columns named left and right for censored data 
 {
@@ -205,19 +205,7 @@ mledist <- function (data, distr, start=NULL, optim.method="default",
         else meth <- "BFGS"
     else
         meth=optim.method
-    
-    if(echo & !cens){
-    
-    print("fn objective")
-    print(ddistname)
-#   cat("--\n")
-#   print(par)
-    cat("init\n")
-    print(fnobj(start, obs=data, ddistnam=ddistname))
-    cat("--\n") 
-    print(fnobj(start, data, ddistname))
-    }   
-    
+        
     # Try to minimize the minus (log-)likelihood using the base R optim function
     if(is.null(custom.optim))
     {
@@ -236,10 +224,6 @@ mledist <- function (data, distr, start=NULL, optim.method="default",
                         hessian = NA))
         }
         
-        if (echo){
-            cat("optimisation with optim\n")
-            print(opt)
-        }
         if (opt$convergence>0) {
             warnings("The function optim failed to converge, with the error code ",
                      opt$convergence)
@@ -267,11 +251,6 @@ mledist <- function (data, distr, start=NULL, optim.method="default",
                         hessian = NA))
         }
         
-        if (echo){  
-            cat("result with custom optimisation \n")
-            print(opt)          
-        }
-
         if (opt$convergence>0) {
             warnings("The customized optimization function failed to converge, with the error code ",
                      opt$convergence)
@@ -285,4 +264,10 @@ mledist <- function (data, distr, start=NULL, optim.method="default",
     }   
         
      
+}
+
+## old function with previous name for censored data
+mledistcens<-function (censdata, distr, start=NULL,optim.method="default",lower=-Inf,upper=Inf)
+{
+    stop("The function \"mledistcens\" is no more used. Now the same function \"mledist\" must be used for censored and non censored data.")
 }

### 2009-12-10 R/mmedist.R

@@ -19,7 +19,7 @@
 #############################################################################
 ### Matching moment estimation for non-censored data
 ###
-###			R functions
+###         R functions
 ### 
 
 mmedist<-function (data, distr) 
@@ -124,4 +124,10 @@ mmedist<-function (data, distr)
         return(estimate)        
    }
  
+}
+
+## old function with previous name 
+momdist<-function (data, distr) 
+{
+    stop("the name \"momdist\" for matching moments function is NO MORE used and is replaced by \"mmedist\".")
 }

### 2009-12-17 inst/doc/intro2fitdistrplus.Rnw
Adding of examples to fitdistcens.Rd and various minor changes
last changes before release for CRAN I hope (minor corrections in .Rd, a new paragraph in the vignette and correction of plotdist.R not to cut density plot).

@@
 
-\title{Use of the library \texttt{fitdistrplus} to specify a distribution from 
+\title{Use of the package \texttt{fitdistrplus} to specify a distribution from 
 non-censored or censored data}
@@
 \begin{document}
+\setkeys{Gin}{width=0.6\textwidth}
 
@@
 
-Here you will find some easy examples of use of the functions of the library \texttt{fitdistrplus}. 
+<<echo=FALSE,results=hide>>=
+options(digits=3)
+@
+
+Here you will find some easy examples of use of the functions of the package \texttt{fitdistrplus}. 
 The aim is to show you by examples how to use these functions to help you to specify a parametric distribution 
@@
 For details, see the documentation of each function, using the R help command (ex.: \texttt{?fitdist}). 
-Do not forget to load the library using the function \texttt{library} before testing following examples.
+Do not forget to load the package using the function \texttt{library} or \texttt{require}
+before testing following examples.
 <<>>=
@@
 
-<<fig=TRUE>>=
+<<plotdistx1,fig=TRUE>>=
 x1<-c(6.4,13.3,4.1,1.3,14.1,10.6,
@@
 
-<<fig=TRUE>>=
+<<descdistx1,fig=TRUE>>=
 descdist(x1)
@@
 
-<<fig=TRUE>>=
+<<descdistx1boot,fig=TRUE>>=
 descdist(x1,boot=1000)
@@
 Below is the result of a fit of a gamma distribution by maximum likelihood.
-<<fig=TRUE>>=
+<<fitdistx1g,fig=TRUE>>=
 f1g<-fitdist(x1,"gamma")
@@
 Below is the fit of a lognormal distribution.
-<<fig=TRUE>>=
+<<fitdistx1ln,fig=TRUE>>=
 f1l<-fitdist(x1,"lnorm")
@@
 Below is the fit of a normal distribution.
-<<fig=TRUE>>=
+<<fitdistx1n,fig=TRUE>>=
 f1n<-fitdist(x1,"norm")
@@
 Below is the fit of a Weibull distribution.
-<<fig=TRUE>>=
+<<fitdistx1w,fig=TRUE>>=
 f1w<-fitdist(x1,"weibull")
@@
 a first plot of the data set with the Gumbel distribution with arbitrary values for parameters.
-<<fig=TRUE>>=
+<<plotdistx1gu1,fig=TRUE>>=
 dgumbel<-function(x,a,b) 1/b*exp((a-x)/b)*exp(-exp((a-x)/b))
@@
 The same data set may be plotted with a Gumbel distribution with modified values for parameters.
-<<fig=TRUE>>=
+<<plotdistx1gu2,fig=TRUE>>=
 plotdist(x1,"gumbel",para=list(a=10,b=5))
@@
 And a Gumbel distribution may be fitted to data with these values for initial parameter values.
-<<fig=TRUE>>=
+<<fitdistx1gu,fig=TRUE>>=
 fgu<-fitdist(x1,"gumbel",start=list(a=10,b=5))
@@
 Below is an example of the use of this function with the previous fit of the gamma distribution. 
-<<fig=TRUE>>=
+<<bootdistf1g,fig=TRUE>>=
 b1g<-bootdist(f1g)
@@
 
-<<fig=TRUE>>=
+<<plotdistx2,fig=TRUE>>=
 x2<-rnbinom(n=100,size=2,prob=0.3)
@@
 
-<<fig=TRUE>>=
+<<descdistx2,fig=TRUE>>=
 descdist(x2,discrete=T)
@@
 Below is the result of the fit of a Poisson distribution with the bootstrap simulations.
-<<fig=TRUE>>=
+<<fitdistx2p,fig=TRUE>>=
 f2p<-fitdist(x2,"pois")
@@
 Below is the result of the fit of a negative binomial distribution with the boostrap simulations.
-<<fig=TRUE>>=
+<<fitdistx2nb,fig=TRUE>>=
 f2n<-fitdist(x2,"nbinom")
@@
 
-<<fig=TRUE>>=
+<<plotdistcensd1,fig=TRUE>>=
 d1<-data.frame(
@@
 
-<<fig=TRUE>>=
+<<plotdistcensd2,fig=TRUE>>=
 d2<-data.frame(left=10^(d1$left),right=10^(d1$right))
@@
 observations, as in the example below for the first dataset.
-<<fig=TRUE>>=
+<<plotdistcensd1bis,fig=TRUE>>=
 plotdistcens(d1,rightNA=3)
@@
 corresponding boostrap simulations.
-<<fig=TRUE>>=
+<<fitdistcensd2w,fig=TRUE>>=
 f2w<-fitdistcens(d2,"weibull")
@@
 
-<<fig=TRUE>>=
+<<bootdistcensf2w,fig=TRUE>>=
 b2w<-bootdistcens(f2w)
@@
 Below is the fit of a lognormal distribution to the same censored data set.
-<<fig=TRUE>>=
+<<fitdistcensd2ln,fig=TRUE>>=
 f2l<-fitdistcens(d2,"lnorm")
@@
 Below is the fit of an exponential distribution.
-<<fig=TRUE>>=
+<<fitdistcensd2e,fig=TRUE>>=
 f2e<-fitdistcens(d2,"exp")
@@
 
+\section{Changing the optimization algorithm used to maximize the likelihood}
+
+Sometimes the default algorithm used to maximize the likelihood fails to converge. It may then
+be interesting to change some options of the function \texttt{optim} or to use another optimization
+function than \texttt{optim} to maximize the likelihood.
+
+\subsection{Changing the arguments passed to \texttt{optim}}
+
+The argument optim.method may be used in the call to \texttt{fitdist} or
+\texttt{fitdistcens}. It will internally be passed to \texttt{mledist} and to \texttt{optim}.
+This argument may be fixed to \texttt{"Nelder-Mead"} (the robust Nelder and Mead method), 
+\texttt{"BFGS"} (the BFGS quasi-Newton method), \texttt{"CG"} (a conjugate gradients method), 
+\texttt{"SANN"} (a variant of simulated annealing) or \texttt{"L-BFGS-B"} (a modification of the BFGS 
+quasi-Newton method which enables box constraints optimization). For the use of the last method the 
+arguments \texttt{lower} and/or \texttt{upper} also have to be passed. More details on these optimization
+functions may be found in the help page of \texttt{optim} from the package \texttt{stats}.
+
+Below are examples of fits of a gamma distribution to non censored data with various options
+of \texttt{optim}.
+
+<<>>=
+fitdist(x1,"gamma",optim.method="Nelder-Mead")
+fitdist(x1,"gamma",optim.method="BFGS") 
+fitdist(x1,"gamma",optim.method="L-BFGS-B",lower=c(0,0))
+fitdist(x1,"gamma",optim.method="SANN")
+@
+
+\subsection{Supplying another optimization function}
+
+You may also want to use another function than \texttt{optim} to maximize the likelihood.
+This optimization function has to be specified by the argument \texttt{custom.optim} 
+in the call to \texttt{fitdist} or
+\texttt{fitdistcens}. But before that, it is necessary to customize this optimization function :
+\texttt{custom.optim} function must have (at least) the following arguments,
+\texttt{fn} for the function to be optimized, \texttt{par} for the initialized parameters. 
+It is assumed that \texttt{custom.optim} should carry out a MINIMIZATION. Finally, it should return
+at least the following components: \texttt{par} for the estimate, \texttt{convergence} for the convergence
+code, \texttt{value} for \texttt{fn(par)} and \texttt{hessian}.
+
+Below is an example of code written to customize \texttt{genoud} function from \texttt{rgenoud} package.
+\begin{verbatim}
+mygenoud <- function(fn, par, ...) 
+{
+   require(rgenoud)
+   res <- genoud(fn, starting.values=par, ...)        
+   standardres <- c(res, convergence=0)
+   return(standardres)
+}
+\end{verbatim}
+
+The customized optimization function may then be passed as the argument \texttt{custom.optim} 
+in the call to \texttt{fitdist}
+or \texttt{fitdistcens}. The following code may for example be used to fit a gamma distribution to the
+non censored data \texttt{x1}. Note that in this example various arguments are also passed from \texttt{fitdist} 
+to \texttt{genoud} :
+\texttt{nvars}, \texttt{Domains}, \texttt{boundary.enforcement}, 
+\texttt{print.level} and \texttt{hessian}.
+
+\begin{verbatim}
+fitdist(x1, "gamma", custom.optim=mygenoud, nvars=2,    
+   Domains=cbind(c(0,0), c(10, 10)), boundary.enforcement=1, 
+   print.level=1, hessian=TRUE)
+\end{verbatim}
+
 \end{document}

### 2010-04-23 inst/doc/intro2fitdistrplus.Rnw
Some minor and major modifications on fitdistrplus (see NEWS)

@@
 
-In order to take into account the uncertainty of the estimated values of kurtosis and skewness, 
+Skewness and kurtosis are known not to be robust. In order to try to
+take into account the uncertainty on the estimated values of kurtosis and skewness, 
 the data set may be boostrapped by fixing the argument \texttt{boot} to an integer above 10 in \texttt{descdist}. 
 \texttt{boot} values of 
-skewness and kurtosis corresponding to the boot nonparametric bootstrap samples are then computed and reported in blue color 
+skewness and kurtosis corresponding to the boot nonparametric bootstrap samples are then computed and reported
 on the skewness-kurtosis plot.
@@
 When fitting continuous distributions, 
-Kolmogorov-Smirnov and Anderson-Darling statistics are computed and corresponding tests are performed when possible. 
-Even if less appropriate for continuous distributions, the Chi-squared statistic is
-also computed when possible. For this calculation, cells are defined by the argument \texttt{chisqbreaks}
-or automatically defined from the data set and from the argument \texttt{meancount} (the approximate mean count per cell)
-which is  fixed to $(4n)^{2/5}$ if omitted (with n the length of the data set).  
-For more details, see the help of the function \texttt{fitdist}.
+Kolmogorov-Smirnov and Anderson-Darling statistics may be computed using the function \texttt{gofstat}. 
 Four goodness of fit plots are also provided.
 
-Below is the result of a fit of a gamma distribution by maximum likelihood.
-<<fitdistx1g,fig=TRUE>>=
-f1g<-fitdist(x1,"gamma")
-plot(f1g)
-summary(f1g)
+Below is the result of the fit of a logistic distribution by maximum likelihood to the previous dataset.
+<<fitdistx1l,fig=TRUE>>=
+f1l<-fitdist(x1,"logis")
+summary(f1l)
+gofstat(f1l)
+plot(f1l)
 @
 
-Below is the result of another fit of the same distribution by matching moments.
+In that case the Anderson-Darling test may be used to test the adequation of the distribution to data. 
 <<>>=
-f1gbis<-fitdist(x1,"gamma",method="mme")
-summary(f1gbis)
+gofstat(f1l,print.test=TRUE)
 @
 
-As can be seen in this returned summary, the automatic definition of the cells required to calculate the Chi-squared
-statistic does not give theoretical counts large enough to validate the use of the test in this example.
- It is often the case for small
-data sets. The observed and theoretical counts may be printed as below :
+Generally goodness-of-fit tests may be used carefully.
+In somme cases, especially on very big datasets, even if the null hypothesis is rejected a fitted distribution may be 
+chosen as the best one among simple distributions to describe an empirical distribution, if the goodness-of-fit
+plots do not show strong differences between empirical and theoretical distributions.
+
+In that case, the fit seems correct, but it is easy to compare goodness-of-fit statistics
+for other distributions fitted on the same dataset in order to check that other distributions could not give
+a better fit. Below are computed the goodness-of-fit
+statistics for logistic, lognormal, gamma, normal and weibull distributions. 
 <<>>=
-f1g$chisqtable
+gofstat(fitdist(x1,"logis"))
+gofstat(fitdist(x1,"lnorm"))
+gofstat(fitdist(x1,"gamma"))
+gofstat(fitdist(x1,"norm"))
+gofstat(fitdist(x1,"weibull"))
 @
 
-Below is the fit of a lognormal distribution.
-<<fitdistx1ln,fig=TRUE>>=
-f1l<-fitdist(x1,"lnorm")
-plot(f1l)
-summary(f1l)
-@
+Regarding the Kolmogorov-Smirnov statistic, the fit of a normal distribution seems better while regarding the 
+Anderson-Darling statistic the fit of a logistic distribution seems better.
+It is recommended to base the comparison of fits on the Anderson-Darling statistic when the modelling of the tails of 
+a distribution is important, as it is often the case in risk assessment.
 
-Below is the fit of a normal distribution.
+In order to choose between both distributions, the goodness-of-fit plot and the summary
+of the fit of a normal chould be compared to the 
+ones already obtained for the logistic distribution, but in that case the differences seem very small.
 <<fitdistx1n,fig=TRUE>>=
-f1n<-fitdist(x1,"norm")
+f1n <- fitdist(x1,"norm")
 plot(f1n)
 summary(f1n)
-@
-
-Below is the fit of a Weibull distribution.
-<<fitdistx1w,fig=TRUE>>=
-f1w<-fitdist(x1,"weibull")
-plot(f1w)
-summary(f1w)
-@
-
-The values of the Anderson-Darling statistic (or another result of the fit: see the help of \texttt{fitdist} for details)
- for the different fittings may be extracted and compared to help the selection of a distribution :
-
-<<>>=
-anderson<-list(lnorm=f1l$ad,gamma=f1g$ad,norm=f1n$ad,weibull=f1w$ad)
-anderson
+plot(f1n)
 @
@@
 summary(fgu)
+gofstat(fgu)
 @
@@
 
-Below is an example of the use of this function with the previous fit of the gamma distribution. 
-<<bootdistf1g,fig=TRUE>>=
-b1g<-bootdist(f1g)
-plot(b1g)
-summary(b1g)
+Below is an example of the use of this function with the previous fit of the logistic distribution. 
+<<bootdistf1l,fig=TRUE>>=
+b1l<-bootdist(f1l)
+plot(b1l)
+summary(b1l)
 @
@@
 <<descdistx2,fig=TRUE>>=
-descdist(x2,discrete=T)
+descdist(x2,discrete=T,boot=1000)
 @
@@
 summary(f2p)
+gofstat(f2p,print.test=TRUE)
 b2p<-bootdist(f2p)
@@
 summary(f2n)
+gofstat(f2n,print.test=TRUE)
 b2n<-bootdist(f2n)
@@
 @
+
+From goodness-of-fit graphs, Chi-squared statistics, AIC and BIC values,
+it seems better to choose the fit of a negative binomial distribution
+for this dataset even it has one more parameter than the Poisson one. This was not obvious while
+looking at the skewness-kurtosis graph. This graph must be used cautiously especially for continuous 
+distributions far from the normal distribution or for discrete distributions. It is only indicative.
 

### 2010-04-23 R/descdist.R

@@ -22,26 +22,62 @@
 ###         R functions
 ### 
 
-descdist <- function(data,discrete=FALSE,boot=NULL,graph=TRUE)
+descdist <- function(data,discrete=FALSE,boot=NULL,method="unbiased",graph=TRUE,
+obs.col="red",boot.col="pink")
 {
     #if(is.mcnode(data)) data <- as.vector(data)
     if (missing(data) || !is.vector(data,mode="numeric"))
         stop("data must be a numeric vector")
+    if (length(data) < 4)
+        stop("data must be a numeric vector containing at least four values")
     moment <- function(data,k){
-        m1<-mean(data)
+        m1 <- mean(data)
         return(sum((data-m1)^k)/length(data))
     }
-    skewness <- function(data){
-        sd<-sqrt(moment(data,2))
-        return(moment(data,3)/sd^3)
+    if (method=="unbiased")
+    {
+        skewness <- function(data){
+        # unbiased estimation (Fisher 1930)
+            sd <- sqrt(moment(data,2))
+            n <- length(data)
+            gamma1 <- moment(data,3)/sd^3
+            unbiased.skewness <- sqrt(n*(n-1)) * gamma1 / (n-2) 
+            return(unbiased.skewness)
+        }
+        kurtosis <- function(data){
+        # unbiased estimation (Fisher 1930)
+            n <- length(data)
+            var<-moment(data,2)
+            gamma2 <- moment(data,4)/var^2 - 3
+            unbiased.kurtosis <- (n-1)/ ((n-2)*(n-3)) * ((n+1)*gamma2 + 6 ) + 3
+            return(unbiased.kurtosis)
+        }
+        standdev <- function(data){
+            sd(data)
+        }
     }
-    kurtosis <- function(data){
-        var<-moment(data,2)
-        return(moment(data,4)/var^2)
-    }
+    else
+        if (method=="sample")
+        {
+            skewness <- function(data) 
+            {
+                sd <- sqrt(moment(data, 2))
+                return(moment(data, 3)/sd^3)
+            }
+            kurtosis <- function(data) 
+            {
+                var <- moment(data, 2)
+                return(moment(data, 4)/var^2)
+            }
+            standdev <- function(data){
[…]

### 2010-04-23 R/fitdist.R

@@ -22,7 +22,7 @@
 ###         R functions
 ### 
 
-fitdist <- function (data, distr, method=c("mle", "mme"), start, chisqbreaks, meancount, ...) 
+fitdist <- function (data, distr, method=c("mle", "mme"), start, ...) 
 {
     if (!is.character(distr)) 
         distname <- substring(as.character(match.call()$distr),2)
@@ -86,137 +86,11 @@ fitdist <- function (data, distr, method=c("mle", "mme"), start, chisqbreaks, me
         npar <- length(estimate)
         aic <- -2*loglik+2*npar
         bic <- -2*loglik+log(n)*npar
-    } 
-    # Goodness of fit statistics
-    if (is.element(distname,c("binom","nbinom","geom","hyper","pois"))) 
-        discrete <- TRUE
-    else 
-        discrete <- FALSE
-    
-    # chi-squared statistic
-    if (missing(chisqbreaks)) { 
-        if (missing(meancount))
-            meancount <- round(n/((4*n)^(2/5)))
-        sdata <- sort(data)
-        if (length(sdata)>ceiling(1.5*meancount)) {
-            limit <- sdata[meancount]
-            sdata <- sdata[sdata>limit]
-            chisqbreaks <- limit
-        }
-        else {
-            warnings("The sample is too small to automatically define chisqbreaks")
-            chisq <- NULL
-            chisqbreaks <- NULL
-            chisqpvalue <- NULL
-            chisqtable <- NULL
-            chisqdf <- NULL
-            
-        }
-        while (length(sdata)>ceiling(1.5*meancount)) {
-            limit <- sdata[meancount]
-            sdata <- sdata[sdata>limit]
-            chisqbreaks <- c(chisqbreaks,limit)
-        } 
-    }
-                
-    if (!is.null(chisqbreaks)) {
-        if(!is.numeric(chisqbreaks))
-            stop("chisqbreaks must be a numeric vector defining the cell boundaries")
-        nbreaks <- length(chisqbreaks)  
-        pbreaks <- do.call(pdistname,c(list(q=chisqbreaks),as.list(estimate)))
-        Fobsbreaks <- ecdf(data)(chisqbreaks)
-        Fobsunder <- c(0,Fobsbreaks[1:nbreaks-1]) 
-        punder <- c(0,pbreaks[1:nbreaks-1])
-        if (pbreaks[nbreaks]==1 & Fobsbreaks[nbreaks]==1) {
-            p <- pbreaks-punder
[…]

### 2010-04-23 R/plotdist.R

@@ -42,7 +42,7 @@ plotdist <- function(data,distr,para,breaks="default",discrete=FALSE,...){
             else 
                 h<-hist(data,freq=FALSE,xlab="data",main=paste("Histogram"),breaks=breaks,...)
             plot(s,obsp,main=paste("Cumulative distribution"),xlab="data",
-            xlim=c(h$breaks[1],h$breaks[length(h$breaks)]),ylab="CDF",pch=4,...)
+            xlim=c(h$breaks[1],h$breaks[length(h$breaks)]),ylab="CDF",...)
         }
         else {
             # plot for discrete data
@@ -54,11 +54,11 @@ plotdist <- function(data,distr,para,breaks="default",discrete=FALSE,...){
             xvalfin<-seq(min(xval),max(xval))
             ydobs<-as.vector(t)/n
             ydmax<-max(ydobs)
-            plot(xval,ydobs,type='h',lwd=5,xlim=xlim,ylim=c(0,ydmax),
+            plot(xval,ydobs,type='h',xlim=xlim,ylim=c(0,ydmax),
             main="Empirical distribution",xlab="data",ylab="Density",...)
             # plot of the cumulative probability distributions
             ycdfobs<-ecdf(data)(xvalfin)
-            plot(xvalfin,ycdfobs,type='h',lwd=5,xlim=xlim,ylim=c(0,1),
+            plot(xvalfin,ycdfobs,type='h',xlim=xlim,ylim=c(0,1),
             main="Empirical CDFs",xlab="data",ylab="CDF",...)
         }
     } #end of if (missing(distr))
@@ -121,7 +121,7 @@ plotdist <- function(data,distr,para,breaks="default",discrete=FALSE,...){
             xmin<-h$breaks[1]
             xmax<-h$breaks[length(h$breaks)]
             plot(s,obsp,main=paste("Empirical and theoretical CDFs"),xlab="data",
-            ylab="CDF",xlim=c(xmin,xmax),pch=4,...)
+            ylab="CDF",xlim=c(xmin,xmax),...)
             sfin<-seq(xmin,xmax,by=(xmax-xmin)/100)
             theopfin<-do.call(pdistname,c(list(q=sfin),as.list(para)))
             lines(sfin,theopfin,lty=1)
@@ -142,14 +142,15 @@ plotdist <- function(data,distr,para,breaks="default",discrete=FALSE,...){
             t<-table(data)
             xval<-as.numeric(names(t))
[…]

### 2011-03-11 inst/doc/intro2fitdistrplus.Rnw
Add of maximum goodness-of-fit estimation (mgedist) and supressionof the makefile

@@
 \section{Specification of a distribution from non-censored continuous data}
-%si on enlève l'étoile on a des numéros de section
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \subsection{Graphical display of the observed distribution}
+%%%%%%%%%%
 
@@
 \subsection{Characterization of the observed distribution}
+%%%%%%%%%%
 
@@
 \subsection{Fitting of a distribution}
+%%%%%%%%%%
 
@@
 fonction \texttt{fitdist}. This function uses the maximum likelihood
-method if the argument \texttt{method="mle"} (or if it is omitted) 
-or the matching moments estimation if the argument \texttt{method="mme"}.
+method by default or other estimation methods described in part 5 if specified in the argument \texttt{method}.
 When fitting continuous distributions, 
-Kolmogorov-Smirnov and Anderson-Darling statistics may be computed using the function \texttt{gofstat}. 
+Cramer-von Mises, Kolmogorov-Smirnov and Anderson-Darling statistics may be computed using the function \texttt{gofstat}. 
 Four goodness of fit plots are also provided.
@@
 
-In that case the Anderson-Darling test may be used to test the adequation of the distribution to data. 
+In that case Cramer-von Mises and Anderson-Darling tests may be used to test the adequation of the distribution to data. 
 <<>>=
@@
 Generally goodness-of-fit tests may be used carefully.
-In somme cases, especially on very big datasets, even if the null hypothesis is rejected a fitted distribution may be 
+In some cases, especially on very big datasets, even if the null hypothesis is rejected a fitted distribution may be 
 chosen as the best one among simple distributions to describe an empirical distribution, if the goodness-of-fit
@@
 
-Regarding the Kolmogorov-Smirnov statistic, the fit of a normal distribution seems better while regarding the 
+Regarding the Cramer-von Mises and Kolmogorov-Smirnov statistics, 
+the fit of a normal distribution seems better while regarding the 
 Anderson-Darling statistic the fit of a logistic distribution seems better.
@@
 \subsection{Simulation of the uncertainty by boostrap}
+%%%%%%%%%%
 
@@
 \section{Specification of a distribution from non-censored discrete data}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 As for continuous distributions, one or more parametric distributions may then be fitted 
-to the data set by maximum likelihood or matching moments.
+to the data set by maximum likelihood or alternative methods described in part 5.
 
@@
 \section{Specification of a distribution from censored data}
-%si on enlève l'étoile on a des numéros de section
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 Censored data may contain left censored, right censored and interval censored values, 
@@
 \subsection{Graphical display of the observed distribution}
+%%%%%%%%%%
 
@@
 \section{Changing the optimization algorithm used to maximize the likelihood}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \subsection{Changing the arguments passed to \texttt{optim}}
+%%%%%%%%%%
 
@@
 \subsection{Supplying another optimization function}
+%%%%%%%%%%
 
@@
 
+\section{Alternative methods to maximum likelihhod for non censored data}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+For non censored data, alternative methods to maximum likelihood are available.
+
+\subsection{Moment matching estimation}
+%%%%%%%%%%
+Moment matching may be performed fixing the argument \texttt{method}
+to \texttt{"mme"} in the call to fitdist (see the help file page of the function
+\texttt{mmedist} in \texttt{fitdistrplus} reference manual for details).
+Below is an example of fit of a uniform distribution to data randomly
+generated in a uniform distribution.
+<<>>=
+u <- runif(50,min=5,max=10)
+fitdist(u,"unif",method="mme")
+@
+
+\subsection{Quantile matching estimation}
+%%%%%%%%%%
+Quantile matching may also be performed fixing the argument \texttt{method}
+to \texttt{"qme"} in the call to fitdist and adding an argument \texttt{probs}
+defining the probabilities for which the quantile matching is done.
+The length of this vector must be equal
+to the number of parameters to estimate.(see the help file page of the function
+\texttt{qmedist} in \texttt{fitdistrplus} reference manual for details).
+Below is an example of fit of a uniform distribution to the same dataset by matching first
+and third quartiles.
+<<>>=
+fitdist(u,"unif",method="qme",probs=c(0.25,0.75))
+@
+
+\subsection{Maximum goodness-of-fit estimation}
+%%%%%%%%%%
+Parameters may also be estimated by maximizing a goodness-of-fit distance
+by fixing the argument \texttt{method}
+to \texttt{"mge"} in the call to fitdist and adding an argument \texttt{gof}
+coding for the goodness-of-fit distance chosen. One may use the classical 
+Cramer-von Mises distance (\texttt{"CvM"}), the classical
+Kolmogorov-Smirnov distance (\texttt{"KS"}), the classical Anderson-Darling distance (\texttt{"AD"})
+which gives more weight to the tails of the distribution,
+or one of the variants of this last distance (see the help file page of the function
+\texttt{mgedist} in \texttt{fitdistrplus} reference manual for details).
+Below is an example of fit of a uniform distribution to the same dataset by maximizing the 
+Kolmogorov-Smirnov distance.
+<<>>=
+fitdist(u,"unif",method="mge",gof="KS")
+@
+
 \end{document}

### 2011-03-14 inst/doc/intro2fitdistrplus.Rnw
minor corrections on .Rd files and vignette

@@
 %%%%%%%%%%
-Parameters may also be estimated by maximizing a goodness-of-fit distance
+Parameters may also be estimated by maximizing goodness-of-fit (or minimizing a goodness-of-fit distance)
 by fixing the argument \texttt{method}

### 2011-04-07 inst/doc/intro2fitdistrplus.Rnw
correction of fitdistrplus vignette following Brian Ripley advice

--- 2011-03-11 inst/doc/intro2fitdistrplus.Rnw
@@
 \texttt{fitdistcens}. It will internally be passed to \texttt{mledist} and to \texttt{optim}.
-This argument may be fixed to \texttt{"Nelder-Mead"} (the robust Nelder and Mead method), 
-\texttt{"BFGS"} (the BFGS quasi-Newton method), \texttt{"CG"} (a conjugate gradients method), 
-\texttt{"SANN"} (a variant of simulated annealing) or \texttt{"L-BFGS-B"} (a modification of the BFGS 
+This argument may be fixed to ``Nelder-Mead'' (the robust Nelder and Mead method), 
+``BFGS'' (the BFGS quasi-Newton method), ``CG'' (a conjugate gradients method), 
+``SANN'' (a variant of simulated annealing) or ``L-BFGS-B'' (a modification of the BFGS 
 quasi-Newton method which enables box constraints optimization). For the use of the last method the 
@@
 
-\section{Alternative methods to maximum likelihhod for non censored data}
+\section{Alternative methods to maximum likelihood for non censored data}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Moment matching may be performed fixing the argument \texttt{method}
-to \texttt{"mme"} in the call to fitdist (see the help file page of the function
+to ``mme'' in the call to fitdist (see the help file page of the function
 \texttt{mmedist} in \texttt{fitdistrplus} reference manual for details).
@@
 Quantile matching may also be performed fixing the argument \texttt{method}
-to \texttt{"qme"} in the call to fitdist and adding an argument \texttt{probs}
+to ``qme'' in the call to fitdist and adding an argument \texttt{probs}
 defining the probabilities for which the quantile matching is done.
@@
 by fixing the argument \texttt{method}
-to \texttt{"mge"} in the call to fitdist and adding an argument \texttt{gof}
+to ``mge'' in the call to fitdist and adding an argument \texttt{gof}
 coding for the goodness-of-fit distance chosen. One may use the classical 
-Cramer-von Mises distance (\texttt{"CvM"}), the classical
-Kolmogorov-Smirnov distance (\texttt{"KS"}), the classical Anderson-Darling distance (\texttt{"AD"})
+Cramer-von Mises distance (``CvM''), the classical
+Kolmogorov-Smirnov distance (``KS''), the classical Anderson-Darling distance (``AD'')
 which gives more weight to the tails of the distribution,

### 2011-07-29 inst/doc/intro2fitdistrplus.Rnw
last non ASCII character eliminated
add of a draft of the new vignette

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%   Copyright (c) 2009 Marie Laure Delignette-Muller, Regis Pouillot, Jean-Baptiste Denis,
-%   Christophe Dutang                        
+%   Copyright (c) 2009 Marie Laure Delignette-Muller, Christophe Dutang                        
 %                                                                                                                                                       
@@
 \usepackage[a4paper,textwidth=18cm,textheight=27cm]{geometry}
+\usepackage{url}
 
-\title{Use of the package \texttt{fitdistrplus} to specify a distribution from 
-non-censored or censored data}
-\author{Marie Laure Delignette-Muller, Régis Pouillot , Jean-Baptiste Denis and Christophe Dutang}
+\title{Fitting parametric univariate distributions  to non censored or
+censored data using the R \texttt{fitdistrplus} package}
+\author{Marie Laure Delignette-Muller and Christophe Dutang}
 
@@
 
-Here you will find some easy examples of use of the functions of the package \texttt{fitdistrplus}. 
-The aim is to show you by examples how to use these functions to help you to specify a parametric distribution 
-from data corresponding to a random sample drawn from a theoretical distribution that you want to describe. 
-For details, see the documentation of each function, using the R help command (ex.: \texttt{?fitdist}). 
-Do not forget to load the package using the function \texttt{library} or \texttt{require}
-before testing following examples.
-<<>>=
+% A enlever dans version JSS !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
+\tableofcontents
+
+%\pagebreak
+
+\section{Introduction}
+\label{Introduction}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Overview}
+\label{Overview}
+%%%%%%%%%%%
+
+Fitting distributions to data is a very common task in statistics. It consists of choosing
+a probability distribution that gives a good representation of a statistical
+variable. It requires judgment and expertise and generally needs an iterative process of
+distribution choice, parameter estimation, and quality of fit evaluation. 
+Function \texttt{fitdistr} in the R package \texttt{MASS} \cite{MASS} 
+is a well known general-purpose maximum-likelihood
+fitting routine for the parameter estimation step in R. Other steps of
+the process may be developed using R \cite{Ricci05}. Our first objective while developping
+package \texttt{fitdistrplus} \cite{fitdistrplus} 
+was to provide R users a set of functions dedicated to help the overall
+process of fitting a univariate parametric distribution to data.
+
+Function \texttt{fitdistr} estimates distribution parameters by maximizing the log-likelihood
+using function \texttt{optim}. In some cases, other estimation methods could be prefered, 
+such as maximum goodness-of-fit estimation also commonly called minimum distance estimation, and
+proposed in package \texttt{actuar} 
+%\cite{actuar}
+with three different goodness-of-fit distances. While developping
+package \texttt{fitdistrplus}, our second objective  was to extend function \texttt{fitdistr} 
+by providing various
+estimation methods to fit distributions in addition to maximum likelihood. Functions
+were developped to enable matching moment estimation, matching quantile estimation, and 
+maximum goodness-of-fit estimation (or minimum distance estimation) using eight different distances.
+Moreover, package \texttt{fitdistrplus} offers the possibility to specify a user-supplied function
+for optimization, 
+useful in cases where optimization techniques not included in function \texttt{optim} may be more adequate.
+
+In applied statistics, it is not uncommon to have to fit distributions to censored data. 
+Function \texttt{fitdistr} 
+does not enable maximum likelihood estimation from this type data. Some packages 
+deal with censored data, especially survival data \cite{survival},
+but those packages generally focused on specific models, enabling the fit of only one distribution or a 
+restricted family of distributions. Our third objective was thus to provide R users a function 
+to estimate univariate distribution parameters from censored data, whatever the type of censoring.
+
+% A changer en cas de mise à jour de fitdistrplus !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
+This paper reviews the various features of version 0.3-0 of \texttt{fitdistrplus}. The package
+is available from the Comprehensive R Archive Network at \url{http://cran.r-project.org/package=fitdistrplus}.
+The development version of the package is located at R-forge as one the packages of the 
+project ``Risk Assessment with R''
+(\url{http://r-forge.r-project.org/projects/riskassessment/})
+The following command will load the package.
+<<results=hide>>=
 library(fitdistrplus)
@@
 
-\tableofcontents
+\subsection{Running examples}
+\label{Examples}
+%%%%%%%%%%%
+For illustrating the use of various functions of package \texttt{fitdistrplus}, we will use
+four examples published in various biological areas, 
+corresponding to data sets included in the package.
+% voir si on en met 3 ou 4 au final !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
 
-\pagebreak
+The two first data sets correspond to the observation of a continuous variable on a random sample
+of a population of interest.
 
-\section{Specification of a distribution from non-censored continuous data}
+The ``ground beef'' data set contains values of
+serving sizes in grams, collected in a French survey, for ground beef patties consumed by 
+children under 5 years old. This data set was used in a quantitative risk assessment published
+in a food microbiology journal (\cite{Delignette08}).
+<<echo=TRUE>>=
+data(groundbeef)
+str(groundbeef)
+@
+
+The ``endosulfan'' data set contains
+acute toxicity values for the organochlorine pesticide endosulfan
+(geometric mean of LC50 ou EC50 values in $\mu g.L^{-1}$),
+tested on Australian and non-Australian laboratory-species 
+(arthropods, fish or nonarthropod invertebrates) (\cite{Hose04}).
+<<echo=TRUE>>=
+data(endosulfan)
+str(endosulfan)
+@
+
+The third data set corresponds to the observation of a discrete variable, 
+the number of \emph{Toxocara cati} parasites present in digestive tract,
+on a random sample of feral 
+cats living on Kerguelen island (\cite{Fromont01}).
+<<echo=TRUE>>=
+data(toxocara)
+str(toxocara)
+@
+
+The last data set corresponds to the observation of a continuous censored variable, 
+the \emph{Listeria monocytogenes} microbial concentration, on a random
+sample of smoked fish distributed on the Belgian market in the period
+2005 to 2007 (\cite{Busschaert10}).
+Censored data are coded within 2 columns named left and right, describing
+each observed value of \emph{Listeria monocytogenes} concentration 
+(in $CFU.g^{-1}$) as an interval. 
+The left column contains either \texttt{NA} for
+left censored observations, the left bound of the interval for interval censored
+observations, or the observed value for non-censored observations. The right
+column contains either \texttt{NA} for right censored observations, the right bound of
+the interval for interval censored observations, or the observed value for noncensored
+observations.
+<<echo=TRUE>>=
+data(smokedfish)
+str(smokedfish)
+@
+
+\section{Choice of candidate distributions}
+\label{Choice}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Before fitting one or more distributions to a data set, it is generally
+necessary to choose good candidates among a predefined family of distributions.
+To help the user in this preliminary task, we developed functions to plot and
+characterise empirical distributions.
 
 \subsection{Graphical display of the observed distribution}
+\label{Graphical}
 %%%%%%%%%%
 
- First of all, the observed distribution may be plotted using the function \texttt{plotdist}.
+First of all, an empirical distribution may be plotted using classical R function
+or using Function \texttt{plotdist} which provides plots
+in density and in cdf as done in (Figure~\ref{plotdistcont}) for a continuous variable:
 
-<<plotdistx1,fig=TRUE>>=
-x1<-c(6.4,13.3,4.1,1.3,14.1,10.6,
-9.9,9.6,15.3,22.1,13.4,13.2,
-8.4,6.3,8.9,5.2,10.9,14.4)
-plotdist(x1)
+<<echo=TRUE, fig=FALSE>>=
+plotdist(groundbeef$serving)
+@
+\begin{figure}[htb]
+  \centering
+<<echo=FALSE, fig=TRUE>>=
+plotdist(groundbeef$serving)
+@
+  \caption{Density and cdf plots of an empirical distribution for a continuous variable}
+  \label{plotdistcont}
+\end{figure}
+
+In some cases a discrete variable may be plotted as a continuous one, for example for a large data set
+from a binomial distribution converging to a normal one, but
+Function \texttt{plotdist} also proposes specific plots in density and in cdf
+for discrete variables (Figure~\ref{plotdistdisc}):
+
+<<echo=TRUE, fig=FALSE>>=
+plotdist(toxocara$number,discrete = TRUE)
+@
+\begin{figure}[htb]
+  \centering
+<<echo=FALSE, fig=TRUE>>=
+plotdist(toxocara$number,discrete = TRUE)
+@
+  \caption{Density and cdf plots of an empirical distribution for a discrete variable}
+  \label{plotdistcont}
+\end{figure}
+
+%\clearpage
+
+\subsection{Empirical basis for selecting candidate distributions}
+\label{Empirical}
+%%%%%%%%%%
+Function \texttt{descdist} 
+Descriptives statistics may help the choice of good candidates to describe an empirical distribution
+among a family of parametric distributions. Especially the skewness and kurtosis are useful for this
+purpose. The concept of skewness relates to deviations from symmetry of the distribution.
+The normal distribution has a skewness of zero. A positive (resp. negative) skewness indicates 
+that the right (resp. left) tail of the distribution is more extended than the left (resp. right) one.
+The concept of kurtosis relates to the tail weight. The normal distribution has a kurtosis of 3.
+Distributions with a higher kurtosis are said to be leptokurtic, with heavier tails, such as the 
+logistic distribution, while distributions with a smaller kurtosis are said platykurtic, with lighter 
+tails, such as the uniform distribution.
+
+Function \texttt{descdist} provides calculations of classical descriptive statistics
+(minimum, maximum, median, mean, sample sd) and by default 
+unbiased estimations of skewness and 
+Pearsons's kurtosis values. Nevertheless estimations of skewness and kurtosis
+are unbiased only for normal distributions and estimated values are thus only indicative.
+A skewness-kurtosis plot such as the one proposed by \cite{Cullen99} is also provided for the 
+empirical distribution (Figure~\ref{Cullenplotcont}). 
+On this plot, values for common distributions are displayed as a tools 
+to help the choice of distributions to fit to data. For some distributions (normal, uniform,
+logistic, exponential for example), there is only one possible value for the skewness and the kurtosis
+ and the distribution is thus represented by a point on the plot. For other distributions, 
+areas of possible values are represented, consisting in lines (as for gamma and lognormal distributions), 
+or larger areas (as for beta distribution).
+
+Skewness and kurtosis are known not to be robust. In order to take into account the uncertainty 
+of the estimated values of kurtosis and skewness from data, the data set may be boostraped by 
+fixing the argument \texttt{boot} to an integer above 10. Values of skewness and kurtosis 
+corresponding to bootstrap samples are then computed and reported on the 
+skewness-kurtosis plot.
+
+<<echo=TRUE, fig=FALSE>>=
+descdist(groundbeef$serving,boot=1000)
+@
+\begin{figure}[htb]
+  \centering
+<<echo=FALSE, results=hide, fig=TRUE>>=
+descdist(groundbeef$serving,boot=1000)
+@
+  \caption{Skewness-kurtosis plot for a continuous variable}
+  \label{Cullenplotcont}
+\end{figure}
+
+For discrete variable, skewness and kurtosis values or set of values of Poisson and negative binomial
+distributions are represented in the skewness-kurtosis plot (Figure~\ref{Cullenplotdisc}), 
+together with values for the
+normal distribution, to which discrete distributions may converge.
+
+<<echo=TRUE, fig=FALSE>>=
+descdist(toxocara$number,discrete = TRUE,boot=1000)
+@
+\begin{figure}[htb]
+  \centering
+<<echo=FALSE, results=hide, fig=TRUE>>=
+descdist(toxocara$number,discrete = TRUE,boot=1000)
+@
+  \caption{Skewness-kurtosis plot for a continuous variable}
+  \label{Cullenplotdisc}
+\end{figure}
+
+\clearpage
+
+\section{Fit of a distribution}
+\label{FIT}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Parameter estimation}
+\label{Estim}
+%%%%%%%%%%
+Once selected, one or more parametric distributions may be fitted to the data set, one at a time, using
+Function \texttt{fitdist}. By default, distribution parameters $\theta$ are estimated by maximizing the
+likelihood defined as:
+  \begin{equation}
+    \label{likelihood}
+    L(\theta)=\prod_{i=1}^n f(y_{i}|\theta)
+  \end{equation}
+with $y_{i}$ the $n$ observations of variable $y$ and $f$ the density function of the fitted 
+parametric distribution.
+
+The other proposed estimation methods are described in Section~\ref{Alternatives}.
+
+Function \texttt{fitdist} returns the results of the fit of any parametric distribution 
+to a data set as an S3 class object that may be easily printed, 
+summarized or plotted (see Figure~\ref{fitgamma} in Section~\ref{GOF plots}). 
+The parametric distribution must be a classically defined R distributions,
+with at least d, p and q functions respectively for the density cdf and quantile functions
+(for example \texttt{dnorm}, \texttt{pnorm} and \texttt{qnorm} for the normal distribution).
+The name of the fitted distribution is specified in the first argument by its classical 
+abbreviation used as the second part of d, p and q functions (for example ``norm''
+for the normal distribution. Numerical results returned by Function \texttt{fitdist} are 
+parameter estimates with estimated standard errors computed from the estimate
+of the Hessian matrix at the maximum likelihood solution, correlation matrix
+between parameter estimates, the loglikelihood, the Akaïke and the Schwarz information criteria 
+(so called AIC and BIC).
+<<echo = TRUE,fig = FALSE>>=
+fw <- fitdist(groundbeef$serving,"weibull")
+print(fw)
+summary(fw)
 @
 
-\subsection{Characterization of the observed distribution}
-%%%%%%%%%%
-
-Descriptive parameters of the empirical distribution may be computed using the function \texttt{descdist}.
-This function will also provide by default a skewness-kurtosis plot which may help you to select 
-which distribution(s) to fit among the potential candidates.
-
-<<descdistx1,fig=TRUE>>=
-descdist(x1)
-@
-
-Skewness and kurtosis are known not to be robust. In order to try to
-take into account the uncertainty on the estimated values of kurtosis and skewness, 
-the data set may be boostrapped by fixing the argument \texttt{boot} to an integer above 10 in \texttt{descdist}. 
-\texttt{boot} values of 
-skewness and kurtosis corresponding to the boot nonparametric bootstrap samples are then computed and reported
-on the skewness-kurtosis plot.
-
-<<descdistx1boot,fig=TRUE>>=
-descdist(x1,boot=1000)
-@
-
-\subsection{Fitting of a distribution}
-%%%%%%%%%%
-
-One or more parametric distributions may then be fitted to the data set, one at a time, using the
-fonction \texttt{fitdist}. This function uses the maximum likelihood
-method by default or other estimation methods described in part 5 if specified in the argument \texttt{method}.
-When fitting continuous distributions, 
-Cramer-von Mises, Kolmogorov-Smirnov and Anderson-Darling statistics may be computed using the function \texttt{gofstat}. 
-Four goodness of fit plots are also provided.
-
-Below is the result of the fit of a logistic distribution by maximum likelihood to the previous dataset.
-<<fitdistx1l,fig=TRUE>>=
-f1l<-fitdist(x1,"logis")
-summary(f1l)
-gofstat(f1l)
-plot(f1l)
-@
-
-In that case Cramer-von Mises and Anderson-Darling tests may be used to test the adequation of the distribution to data. 
-<<>>=
-gofstat(f1l,print.test=TRUE)
-@
-
-Generally goodness-of-fit tests may be used carefully.
-In some cases, especially on very big datasets, even if the null hypothesis is rejected a fitted distribution may be 
-chosen as the best one among simple distributions to describe an empirical distribution, if the goodness-of-fit
-plots do not show strong differences between empirical and theoretical distributions.
-
-In that case, the fit seems correct, but it is easy to compare goodness-of-fit statistics
-for other distributions fitted on the same dataset in order to check that other distributions could not give
-a better fit. Below are computed the goodness-of-fit
-statistics for logistic, lognormal, gamma, normal and weibull distributions. 
-<<>>=
-gofstat(fitdist(x1,"logis"))
-gofstat(fitdist(x1,"lnorm"))
-gofstat(fitdist(x1,"gamma"))
-gofstat(fitdist(x1,"norm"))
-gofstat(fitdist(x1,"weibull"))
-@
-
-Regarding the Cramer-von Mises and Kolmogorov-Smirnov statistics, 
-the fit of a normal distribution seems better while regarding the 
-Anderson-Darling statistic the fit of a logistic distribution seems better.
-It is recommended to base the comparison of fits on the Anderson-Darling statistic when the modelling of the tails of 
-a distribution is important, as it is often the case in risk assessment.
-
-In order to choose between both distributions, the goodness-of-fit plot and the summary
-of the fit of a normal chould be compared to the 
-ones already obtained for the logistic distribution, but in that case the differences seem very small.
-<<fitdistx1n,fig=TRUE>>=
-f1n <- fitdist(x1,"norm")
-plot(f1n)
-summary(f1n)
-plot(f1n)
+The same procedure is required to fit a discrete distribution. As an example, 
+using ``toxocara'' data set, Poisson and
+negative distributions may be easily fitted and AIC values compared, in this case
+giving the preference
+to the negative binomial distribution, with a much smaller AIC value.
+<<echo = TRUE,fig = FALSE>>=
+(fp <- fitdist(toxocara$number,"pois"))
+(fnb <- fitdist(toxocara$number,"nbinom"))
+fp$aic
+fnb$aic
 @
@@
 of parameters initial values. The names of
-the parameters in \texttt{start} must correspond exactly to their definition in R or to their 
-definition in a previous R code.
- The function \texttt{plotdist} may help to find correct initial values for 
-the distribution parameters in non trivial cases, by an manual iterative use if necessary.
+the parameters in \texttt{start} must correspond exactly to their definition in R or in a user-supplied R code.
+Function \texttt{plotdist} (see Section~\ref{GOF plots}), 
+which can plot any parametric distribution with specified parameter values
+in argument \texttt{para} may help to find correct initial values for 
+the distribution parameters in non trivial cases, by iterative calls if necessary (see~\cite{fitdistrplus}
+for examples).
 
-For example, below is the definition of the Gumbel distribution (also named extreme value distribution) and 
-a first plot of the data set with the Gumbel distribution with arbitrary values for parameters.
-<<plotdistx1gu1,fig=TRUE>>=
-dgumbel<-function(x,a,b) 1/b*exp((a-x)/b)*exp(-exp((a-x)/b))
-pgumbel<-function(q,a,b) exp(-exp((a-q)/b))
-qgumbel<-function(p,a,b) a-b*log(-log(p))
-plotdist(x1,"gumbel",para=list(a=3,b=2))
+% Faudra trouver un autre exemple ou une autre distribution ou enlever cette partie
+
+%As an example, here is the fit of a user-supplied distribution, the Gumbel distribution 
+%(also named extreme value distribution) to an arbitrary data set, preceded by 
+%a plot of this data set with the Gumbel distribution which parameters 
+%were fixed at initial values manually obtained after iterative plots.
+
+%<<echo=TRUE,fig=TRUE>>=
+%x1 <- c(6.4, 13.3, 4.1, 1.3, 14.1, 10.6, 9.9, 9.6, 15.3, 22.1,
+%13.4, 13.2, 8.4, 6.3, 8.9, 5.2, 10.9, 14.4)
+%dgumbel<-function(x,a,b) 1/b*exp((a-x)/b)*exp(-exp((a-x)/b))
+%pgumbel<-function(q,a,b) exp(-exp((a-q)/b))
+%qgumbel<-function(p,a,b) a-b*log(-log(p))
+%plotdist(x1,"gumbel",para=list(a=10,b=5))
+%fitdist(x1,"gumbel",start=list(a=10,b=5))
+%@
+%\begin{figure}[htb]
+%  \centering
+%<<echo=FALSE, results=hide, fig=TRUE>>=
+%plotdist(x1,"gumbel",para=list(a=10,b=5))
+%@
+%  \caption{Preliminar plot of a distribution against data to find initial values for parameters}
+%  \label{plotdist}
+%\end{figure}
+
+\subsection{Goodness-of-fit plots}
+\label{GOF plots}
+%%%%%%%%%%
+The plot of an object of class \texttt{fitdist} provides two types of results depending of the nature 
+of the distribution, continuous or discrete. For continuous distributions, 
+four goodness-of-fit plots are provided : a draw of pdf curve and histogram together, an cdf plot
+of both empirical and theoretical distributions, a Q-Q plot
+(plot of the quantiles of the theoretical fitted distribution (x-axis) against the empirical quantiles of the data)
+and a P-P plot (i.e. for each value of the data set, plot of the cumulative density function of the fitted distribution
+(x-axis) against the empirical cumulative density function (y-axis)) are also given (\cite{Cullen99}).
+As an exemple, let us look at the plot of the previous fit of a weibull distribution to ``groundbeef'' data set.
+
+<<echo=TRUE, fig=FALSE>>=
+plot(fw)
+@
+\begin{figure}[htb]
+  \centering
+<<echo=FALSE, results=hide, fig=TRUE>>=
+plot(fw)
+@
+  \caption{Plot of the fit of a continuous distribution}
+  \label{plotcontfit}
+\end{figure}
+
+For continuous distributions, Function \texttt{cdfcomp} enables the visual comparison
+of empirical and theoretical cumulative distributions for various distributions fitted
+on a same data set. Function \texttt{cdfcomp} must be called with a first argument 
+corresponding to a list of objects of class \texttt{fitdist}, and optionnaly
+further arguments to customize the plot, as in the following example comparing the fit of
+Weibull, lognormal and gamma distributions to ``groundbeef'' data set.
+<<echo=TRUE, fig=FALSE>>=
+    fg <- fitdist(groundbeef$serving,"gamma")
+    fln <- fitdist(groundbeef$serving,"lnorm")
+    cdfcomp(list(fw,fln,fg),legendtext=c("Weibull","lognormal","gamma"),
+    xlab="serving sizes (g)",lwd=2)
+@
+\begin{figure}[htb]
+  \centering
+<<echo=FALSE, results=hide, fig=TRUE>>=
+    cdfcomp(list(fw,fln,fg),legendtext=c("Weibull","lognormal","gamma"),
+    xlab="serving sizes (g)",lwd=2)
+@
+  \caption{Comparison of CDF plots of various distributions fitted on continuous data}
+  \label{compcdf}
+\end{figure}
+
+In such a plot, data may be represented in a log scale when required, by just fixing the argument
+\texttt{xlogscale} to \texttt{TRUE} in the call to \texttt{cdfcomp}.
+% Ajouter ex endosulfan avec la bibliothèque actuar !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
+% mais comment faire dans la vignette : indiquer que fitdistrplus nécessite actuar????????????????????????????
+
+For discrete distributions, the plot of an object of class \texttt{fitdist} provides
+two goodness-of-fit plots comparing empirical and theoretical distributions
+in pdf and in cdf.
+As an exemple, let us look at the plot of the previous fit of a negative binomial distribution 
+to ``toxocara'' data set.
+
+<<echo=TRUE, fig=FALSE>>=
+plot(fnb)
+@
+\begin{figure}[htb]
+  \centering
+<<echo=FALSE, results=hide, fig=TRUE>>=
+plot(fnb)
+@
+  \caption{Plot of the fit of a discrete distribution}
+  \label{plotdiscfit}
+\end{figure}
+
+\subsection{Measures of goodness-of-fit}
+\label{GOF measures}
+%%%%%%%%%%
+When fitting continuous distributions, 
+Cramer-von Mises, Kolmogorov-Smirnov and Anderson-Darling statistics may be computed using the function \texttt{gofstat}
+as defined by Stephens (\cite{Stephens86}).
+<<echo=TRUE,fig=FALSE>>=
+gofstat(fw)
 @
 
-The same data set may be plotted with a Gumbel distribution with modified values for parameters.
-<<plotdistx1gu2,fig=TRUE>>=
-plotdist(x1,"gumbel",para=list(a=10,b=5))
+% ADD the definition of these statistics !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
+EQUATIONS TO BE ADDED
+
+As giving  more weight to distribution tails, 
+Anderson-Darling statistics is of special interest where it is important to place equal 
+emphasis on fitting a distribution at the tails as well as the main body, 
+as it is often the case in risk assessment \cite{Cullen99,Vose10}.
+
+% ADD AN EXAMPLE with comparison !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
+Nevertheless, this statistics should 
+be used cautiously when comparing fits of various distributions, keeping in mind that the weighting of each
+cdf quadratic difference is dependent of the theoretical distribution.
+
+When fitting discrete distributions, the Chi-squared statistic is computed by 
+Function \texttt{gofstat} using cells defined 
+by the argument \texttt{chisqbreaks} or cells automatically defined from the data in order 
+to reach roughly the same number of observations per cell, roughly equal to the argument
+\texttt{meancount}, or sligthly more if there are some ties. 
+The choice to define cells from the empirical distribution (data) and not from the
+theoretical distribution was done to enable the comparison of Chi-squared values obtained
+with different distributions fitted on a same dataset.
+If arguments \texttt{chisqbreaks} and \texttt{meancount} 
+are both omitted, \texttt{meancount} is fixed in order to obtain roughly $(4n)^{2/5}$ cells, 
+with $n$ the length of the dataset~\cite{Vose10}. Using this default option with 
+the fit of a negative binomial distribution to ``toxocara'' data set gives following results :
+<<echo=TRUE,fig=FALSE>>=
+gofstat(fnb)
 @
 
-And a Gumbel distribution may be fitted to data with these values for initial parameter values.
-<<fitdistx1gu,fig=TRUE>>=
-fgu<-fitdist(x1,"gumbel",start=list(a=10,b=5))
-plot(fgu)
-summary(fgu)
-gofstat(fgu)
+Among its returned values, Function \texttt{gofstat} provides a table with observed and theoretical 
+counts used for the Chi-squared calculations:
+<<echo=TRUE,fig=FALSE>>=
+gofstat(fnb)$chisqtable
 @
 
-\subsection{Simulation of the uncertainty by boostrap}
+Even if specifically recommended for discrete distributions, the Chi-squares statistic may also be used for
+continuous distributions (see~\cite{fitdistrplus} for examples).
+
+\subsection{Goodness-of-fit tests}
+\label{GOF tests}
 %%%%%%%%%%
 
-The uncertainty in the parameters of the fitted distribution may be simulated by parametric or nonparametric
-boostrap using the function \texttt{boodist}. This function returns the boostrapped values of parameters which
-may be plotted to visualize the bootstrap region. It also calculates the 95 percent confidence intervals for each 
-parameter from the 2.5 and 97.5 percentiles
-of the boostrap values of each parameter (see the help of the function \texttt{bootdist} for details).
+For continuous distributions, an approximate Kolmogorov-Smirnov test is 
+performed by assuming the distribution parameters known. The critical value defined by Stephens~\cite{Stephens86} 
+for a completely specified distribution is used to reject or not the 
+distribution at the significance level 0.05. Because of this approximation, the result of the test
+(decision of rejection of the distribution or not) is returned only for datasets with more 
+than 30 observations. Note that this approximate test may be too conservative.
 
-Below is an example of the use of this function with the previous fit of the logistic distribution. 
-<<bootdistf1l,fig=TRUE>>=
-b1l<-bootdist(f1l)
-plot(b1l)
-summary(b1l)
+For datasets with more than 5 observations and for continuous distributions for 
+which the test is described by Stephens~\cite{Stephens86} (normal, lognormal,
+exponential, Cauchy, gamma, logistic and Weibull),
+the Cramer-von Mises and Anderson-darling tests are performed as described by Stephens~\cite{Stephens86}. 
+Those tests take into 
+account the fact that the parameters are not known but estimated from the data. The result is the 
+decision to reject or not the distribution at the significance level 0.05. Both tests are available
+only for maximum likelihood estimations.
+
+When the Chi-squared statistic is computed (for discrete or optionnaly continuous distributions), 
+and if the degree of freedom (nb of cells - nb of parameters - 1)  of the corresponding distribution 
+is strictly positive, the p-value of the Chi-squared test is returned.
+
+Goodness-of-fit tests may be used carefully. As for any null-hypothesis significance
+test, the non reject of the null hypothesis dose not imply its acceptation. However, this
+misinterpretation of p-values is very common and comes from the wrong assumption that
+absence of evidence is evidence of absence~\cite{Altman96}.
+On the contrary, in some cases, especially on very big datasets, 
+even if the null hypothesis is rejected a fitted distribution may be 
+chosen as the best one among simple distributions to describe an empirical distribution, if the goodness-of-fit
+plots do not show strong differences between empirical and theoretical distributions.
+
+% ADD an example to argument that !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
+
+Now let us look at the Chi-squared test results for the fit of a negative binomial distribution 
+to ``toxocara'' data set :
+<<echo=TRUE,fig=FALSE>>=
+gofstat(fnb,print.test = TRUE)
 @
 
-\section{Specification of a distribution from non-censored discrete data}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
- A discrete data set may be considered as a continuous one for example for a large data set
- from a binomial distribution converging to a normal one. A discrete plot of the distribution 
- may also be provided, fixing the argument discrete of the function \texttt{plotdist} to \texttt{TRUE}.
-
-<<plotdistx2,fig=TRUE>>=
-x2<-rnbinom(n=100,size=2,prob=0.3)
-plotdist(x2,discrete=TRUE)
-@
-
-As for continuous distributions, descriptive parameters of the empirical distribution 
-may be computed using the function \texttt{descdist} which also provides a skewness-kurtosis plot 
-which may help you to choose which distribution(s) to fit.
-
-<<descdistx2,fig=TRUE>>=
-descdist(x2,discrete=T,boot=1000)
-@
-
-As for continuous distributions, one or more parametric distributions may then be fitted 
-to the data set by maximum likelihood or alternative methods described in part 5.
-
-Below is the result of the fit of a Poisson distribution with the bootstrap simulations.
-<<fitdistx2p,fig=TRUE>>=
-f2p<-fitdist(x2,"pois")
-plot(f2p)
-summary(f2p)
-gofstat(f2p,print.test=TRUE)
-b2p<-bootdist(f2p)
-summary(b2p)
-@
-
-Below is the result of the fit of a negative binomial distribution with the boostrap simulations.
-<<fitdistx2nb,fig=TRUE>>=
-f2n<-fitdist(x2,"nbinom")
-plot(f2n)
-summary(f2n)
-gofstat(f2n,print.test=TRUE)
-b2n<-bootdist(f2n)
-summary(b2n)
+A warning message appears as one of the theoretical counts is under 5 using the default breaks 
+(see Section~\ref{GOF measures}). In order to solve this problem, one may specify breaks 
+more adapted for the realization of the test.
+<<echo=TRUE,fig=FALSE>>=
+gofstat(fnb,chisqbreaks=c(0,1,4,8,20),print.test=TRUE)$chisqtable
 @
@@
 
-\section{Specification of a distribution from censored data}
+%\clearpage
+
+\section{The special case of censored data}
+\label{censored}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \subsection{Graphical display of the observed distribution}
+\label{censored graph}
 %%%%%%%%%%
 
-First of all, the observed distribution may be plotted using the function \texttt{plotdistcens}.
+Using censored data such as those coded in the ``smokedfish'' data set,
+the empirical distribution may be plotted using the function \texttt{plotdistcens}.
 Data are reported directly as segments for interval, left and right censored data, 
-and as points for non-censored data. For more details, see the help of the function \texttt{plotdistcens}.
+and as points for non-censored data. Before plotting, observations are ordered and a rank r
+is associated to each of them. Left censored observations are ordered
+first, by their right bounds. Interval censored and non censored observations
+are then ordered by their mid-points and, at last, right censored observations are
+ordered by their left bounds. If argument \texttt{leftNA} (resp. \texttt{rightNA}) is finite,
+left censored (resp. right censored) observations are considered as interval censored
+observations and ordered by mid-points with non-censored and interval censored data.
+It is sometimes necessary to fix \texttt{leftNA} or \texttt{rightNA} to a realistic 
+extreme value, even if not exactly known, to obtain a reasonable global ranking of 
+observations. After ranking, each of the n observations is plotted as a point (one x-value) 
+or a segment (an interval of possible x-values),
+with an y-value equal to r/n, r being the rank of each observation in the global ordering
+previously described.
 
-<<plotdistcensd1,fig=TRUE>>=
-d1<-data.frame(
-left=c(1.73,1.51,0.77,1.96,1.96,-1.4,-1.4,NA,-0.11,0.55,0.41,
-    2.56,NA,-0.53,0.63,-1.4,-1.4,-1.4,NA,0.13),
-right=c(1.73,1.51,0.77,1.96,1.96,0,-0.7,-1.4,-0.11,0.55,0.41,
-    2.56,-1.4,-0.53,0.63,0,-0.7,NA,-1.4,0.13))
-plotdistcens(d1)
+Let us see the sulting plot for ``smokedfish'' data set after classical transformation
+of microbila counts in decimal logarithm.
+
+<<echo=TRUE, fig=FALSE>>=
+log10C <- data.frame(left=log10(smokedfish$left),right=log10(smokedfish$right))
+plotdistcens(log10C)
+@
+\begin{figure}[htb]
+  \centering
+<<echo=FALSE, results=hide, fig=TRUE>>=
+plotdistcens(log10C)
+@
+  \caption{CDF plot of a censored data set as}
+  \label{plotdistcens}
+\end{figure}
+
+\subsection{Maximum likelihood estimation}
+\label{censored MLE}
+%%%%%%%%%%
+
+As for non censored data, one or more parametric distributions may then 
+be fitted to the censored data set, one at a time, but using in this case 
+Function \texttt{fitdistcens}. This function estimates the distribution parameters
+by maximizing the likelihood given by following equation for censored data.
+As \texttt{fitdist}, it returns the results of the fit of any parametric distribution 
+to a data set as an S3 class object that may be easily printed, 
+summarized or plotted.
+% ADD the equation !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
+EQUATION TO ADD
+
+For ``smokedfish'' data set, a normal distribution may be fitted to log
+transformed data as commonly done for microbial count data. 
+<<echo=TRUE,fig=FALSE>>=
+flog10C <- fitdistcens(log10C,"norm")
+print(flog10C)
+summary(flog10C)
 @
 
-When left or right NA-values correspond to finite value (for example 0 for
-left NA-values of positive data), the arguments \texttt{leftNA} (or \texttt{rightNA})
-must be affected to this finite value to ensure a correct plot of left (or right) censored
-observations, as in the example below.
+As with \texttt{fitdist}, for some distributions (see \cite{fitdistrplus} for details), 
+it is necessary to specify initial 
+values for the distribution parameters in the argument \texttt{start}.
+The function \texttt{plotdistcens} may help to find correct initial values for 
+the distribution parameters in non trivial cases, by an manual iterative use if necessary.
 
-<<plotdistcensd2,fig=TRUE>>=
-d2<-data.frame(left=10^(d1$left),right=10^(d1$right))
-plotdistcens(d2,leftNA=0)
+\subsection{Goodness-of-fit plot}
+\label{censored GOF graph}
+%%%%%%%%%%
+Only one goodness-of-fit plot is provided for censored data, correponding to the theoretical 
+cumulative distribution function added to the plot of censored data presented
+in Section~\ref{censored graph}.
+
+<<echo=TRUE, fig=FALSE>>=
+plot(flog10C)
+@
+\begin{figure}[htb]
+  \centering
+<<echo=FALSE, results=hide, fig=TRUE>>=
+plot(flog10C)
+@
+  \caption{Goodness-of-fit CDF plot for a fit of a continuous distribution
+  on censored data}
+  \label{plotfitdistcens}
+\end{figure}
+
+Computations of goodness of fit statistics have not yet been 
+developed for fits using censored data, so the quality of fit 
+may only be estimated from the loglikelihood and the goodness-of-fit CDF plot.
+
+\section{Alternative methods for parameter estimation}
+\label{Alternatives}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Maximum goodness-of-fit estimation}
+\label{MGE}
+%%%%%%%%%%
+Maximum likelihood is only the default estimation method proposed by Function
+\texttt{fitdist}, but other methods may be used to estimate parameters for non-censored data.
+One of the alternative for continuous distributions
+is the maximum goodness-of-fit estimation method also called
+minimum distance estimation method. In this package this method is proposed with
+eight different distances the three classical distances defined in \cite{Stephens86}, 
+(Cramer-von Mises, Kolmogorov-Smirnov and Anderson-Darling 
+which gives more weight to the tails of the distribution),
+or one of the variants of this last distance proposed by \cite{Luceno06}. 
+The right-tail AD gives more weight only to the right tail, the left-tail AD 
+gives more weight only to the left tail. Either of the tails, or both of them, 
+can receive even larger
+weights by using second order Anderson-Darling Statistics.
+
+%Computational forms of these distances are given in Table~\ref{MGEdist}.
+% ADD TABLE with formulas !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
+TABLE with formulas TO BE ADDED
+
+To fit a distribution by maximum goodness-of-fit estimation, one
+needs to fix the argument \texttt{method}
+to ``mge'' in the call to \texttt{fitdist} and to specify the argument \texttt{gof}
+coding for the chosen goodness-of-fit distance. 
+% Character string coding for the eight 
+% proposed goodness-of-fit distances are given in Table~\ref{MGEdist}.
+This function is intended to be used only with continuous variables and distributions.
+It may be useful to fit distributions for which maximum likelihood does not provide
+good estimations, such as the uniform distribution (\cite{Luceno06}).
+
+<<>>=
+u <- runif(50)
+fitdist(u,"unif",method="mge",gof="KS")
 @
 
-It is also possible to fix \texttt{rightNA} or \texttt{leftNA} to a realistic 
-extreme value, even if not exactly known, to obtain a reasonable global ranking of 
-observations, as in the example below for the first dataset.
-<<plotdistcensd1bis,fig=TRUE>>=
-plotdistcens(d1,rightNA=3)
+Maximum goddness-of-fit estimation may also be useful to give
+more weight to data at one tail of the distribution. In ecotoxicology, species sensitivity distributions such 
+as those presented in \cite{Hose04} are often
+fitted by a lognormal distribution (or another parametric distribution)
+so as to estimate a low percentile, often 5$\%$ percentile,
+named the hazardous concentration 5$\%$ (HC5). This value is then interpreted as a value of the
+contaminant concentration protecting 95$\%$ of the species. In this context, one may consider
+to fit the parametric distribution by giving more weight to the left tail of the empirical distribution
+such as in the following example using left tail Anderson-Darling distances of first or second order.
+
+<<echo=TRUE, fig=FALSE>>=
+data(endosulfan)
+ATV <-subset(endosulfan,group == "NonArthroInvert")$ATV
+flnMGEKS <- fitdist(ATV,"lnorm",method="mge",gof="KS")
+flnMGEAD <- fitdist(ATV,"lnorm",method="mge",gof="AD")
+flnMGEADL <- fitdist(ATV,"lnorm",method="mge",gof="ADL")
+flnMGEAD2L <- fitdist(ATV,"lnorm",method="mge",gof="AD2L")
+cdfcomp(list(flnMGEKS,flnMGEAD,flnMGEADL,flnMGEAD2L),
+xlogscale = TRUE,main="",
+legendtext = c("Kolmogorov-Smirnov (KS)","Anderson-Darling",
+"Left-tail Anderson-Darling","Left tailed Anderson-Darling of second order"),cex=0.7,
+xlegend = 500, ylegend = 0.15)
+@
+\begin{figure}[htb]
+  \centering
+<<echo=FALSE, results=hide, fig=TRUE>>=
+cdfcomp(list(flnMGEKS,flnMGEAD,flnMGEADL,flnMGEAD2L),
+xlogscale = TRUE,main="",
+legendtext = c("Kolmogorov-Smirnov (KS)","Anderson-Darling",
+"Left-tail Anderson-Darling","Left tailed Anderson-Darling of second order"),cex=0.7,
+xlegend = 500, ylegend = 0.15)
+@
+  \caption{Comparison of lognormal distributions fitted by maximum goodness-of-fit using
+  various goodness-of-fit distances}
+  \label{plotfitMGE}
+\end{figure}
+
+\subsection{Moment matching estimation}
+\label{MME}
+%%%%%%%%%%
+Moment matching estimation may also be performed fixing the argument \texttt{method}
+to ``mme'' in the call to fitdist.
+<<>>=
+fitdist(u,"unif",method="mme")
 @
 
-\subsection{Fitting of a distribution}
+The estimate is computed by a closed formula for following distributions:
+normal, lognormal, exponential, Poisson, gamma, logistic, negative binomial,
+geometric, beta and uniform distributions.
+For distributions characterized by one parameter (geometric, Poisson and exponential),
+this parameter is simply estimated by matching theoretical and observed means, and for distributions 
+characterized by two parameters, these parameters are estimated by matching theoretical and observed 
+means and variances (\cite{Vose10}).
+
+For other distributions, Function \texttt{fitdist} carries out 
+the matching numerically, by minimization of the
+sum of squared differences between observed and theoretical moments (see \cite{fitdistrplus}
+for technical details).
+
+\subsection{Quantile matching estimation}
+\label{QME}
 %%%%%%%%%%
-
-One or more parametric distributions may then be fitted to the censored data set, one at a time, using the
-fonction \texttt{fitdistcens}. This function always uses the maximum likelihood
-method.  For more details, see the help of the function \texttt{fitdistcens}.
-Only one goodness of fit plot is provided for censored data, in cumulative frequencies.
-The uncertainty in the parameters of the fitted distribution may be simulated by nonparametric
-boostrap only, using the function \texttt{boodistcens}.
-
-Below is the result of a fit of a Weibull distribution by maximum likelihood and the results of the 
-corresponding boostrap simulations.
-<<fitdistcensd2w,fig=TRUE>>=
-f2w<-fitdistcens(d2,"weibull")
-summary(f2w)
-plot(f2w,leftNA=0)
+Quantile matching may also be performed fixing the argument \texttt{method}
+to ``qme'' in the call to fitdist and adding an argument \texttt{probs}
+defining the probabilities for which the quantile matching is performed.
+The length of this vector must be equal
+to the number of parameters to estimate. The quantile matching is carried out numerically, 
+by minimizing the
+sum of squared differences between observed and theoretical quantiles.
+Here is an example of fit of a uniform distribution by matching first
+and third quartiles.
+<<>>=
+fitdist(u,"unif",method="qme",probs=c(0.25,0.75))
 @
 
-<<bootdistcensf2w,fig=TRUE>>=
-b2w<-bootdistcens(f2w)
-summary(b2w)
-plot(b2w)
-@
+\subsection{Customization of the optimization algorithm}
+\label{Customization}
+%%%%%%%%%%
+Each time a numerical minimization (or maximization) is carried out using \texttt{fitdist}, 
+Function \texttt{optim}of the package \texttt{stats} is used by default with the 
+``Nelder-Mead'' method for distributions characterized by more than one parameter and
+the ``BFGS'' method for distributions characterized by only one parameter
 
-Goodness of fit statistics are not computed for fit on censored data, so the quality of fit 
-may only be estimated from the loglikelihood and the goodness of fit plot.
-
-Below is the fit of a lognormal distribution to the same censored data set.
-<<fitdistcensd2ln,fig=TRUE>>=
-f2l<-fitdistcens(d2,"lnorm")
-summary(f2l)
-plot(f2l,leftNA=0)
-@
-
-Below is the fit of an exponential distribution.
-<<fitdistcensd2e,fig=TRUE>>=
-f2e<-fitdistcens(d2,"exp")
-summary(f2e)
-plot(f2e,leftNA=0)
-@
-
-As with \texttt{fitdist}, for some distributions (see the help of \texttt{fitdistcens} for details), 
-it is necessary to specify initial 
-values for the distribution parameters in the argument \texttt{start}.
-\texttt{start} must be a named list
-of parameters initial values. The names of
-the parameters in \texttt{start} must correspond exactly to their definition in R or to their 
-definition in a previous R code.
-The function \texttt{plotdistcens} may help to find correct initial values for 
-the distribution parameters in non trivial cases, by an manual iterative use if necessary, as explained 
-previously for non-censored continuous data.
-
-\section{Changing the optimization algorithm used to maximize the likelihood}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Sometimes the default algorithm used to maximize the likelihood fails to converge. It may then
+Sometimes the default algorithm fails to converge. It may then
 be interesting to change some options of the function \texttt{optim} or to use another optimization
@@
 
-\subsection{Changing the arguments passed to \texttt{optim}}
-%%%%%%%%%%
-
-The argument optim.method may be used in the call to \texttt{fitdist} or
+The argument \texttt{optim.method} may be used in the call to \texttt{fitdist} or
 \texttt{fitdistcens}. It will internally be passed to \texttt{mledist} and to \texttt{optim}.
@@
 
-Below are examples of fits of a gamma distribution to non censored data with various options
+Here are examples of fits of a gamma distribution to ``ground beeef'' data set with various options
 of \texttt{optim}.
@@
 <<>>=
-fitdist(x1,"gamma",optim.method="Nelder-Mead")
-fitdist(x1,"gamma",optim.method="BFGS") 
-fitdist(x1,"gamma",optim.method="L-BFGS-B",lower=c(0,0))
-fitdist(x1,"gamma",optim.method="SANN")
+fitdist(groundbeef$serving,"gamma",optim.method="Nelder-Mead")
+fitdist(groundbeef$serving,"gamma",optim.method="BFGS") 
+fitdist(groundbeef$serving,"gamma",optim.method="L-BFGS-B",lower=c(0,0))
+fitdist(groundbeef$serving,"gamma",optim.method="SANN")
 @
-
-\subsection{Supplying another optimization function}
-%%%%%%%%%%
 
@@
 or \texttt{fitdistcens}. The following code may for example be used to fit a gamma distribution to the
-non censored data \texttt{x1}. Note that in this example various arguments are also passed from \texttt{fitdist} 
+``ground beef'' data set. Note that in this example various arguments are also passed from \texttt{fitdist} 
 to \texttt{genoud} :
@@
 \begin{verbatim}
-fitdist(x1, "gamma", custom.optim=mygenoud, nvars=2,    
+fitdist(groundbeef$serving, "gamma", custom.optim=mygenoud, nvars=2,    
    Domains=cbind(c(0,0), c(10, 10)), boundary.enforcement=1, 
@@
 
-\section{Alternative methods to maximum likelihood for non censored data}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-For non censored data, alternative methods to maximum likelihood are available.
+\section{Uncertainty on parameter estimates}
+\label{Uncertainty}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Bootstrap procedures}
+\label{Bootstrap}
+%%%%%%%%%%
 
-\subsection{Moment matching estimation}
-%%%%%%%%%%
-Moment matching may be performed fixing the argument \texttt{method}
-to ``mme'' in the call to fitdist (see the help file page of the function
-\texttt{mmedist} in \texttt{fitdistrplus} reference manual for details).
-Below is an example of fit of a uniform distribution to data randomly
-generated in a uniform distribution.
-<<>>=
-u <- runif(50,min=5,max=10)
-fitdist(u,"unif",method="mme")
+The uncertainty in the parameters of the fitted distribution may be simulated by parametric or nonparametric
+boostrap using the function \texttt{boodist}. This function returns the boostrapped values of parameters in a 
+S3 class object which
+may be plotted to visualize the bootstrap region. 
+The medians and the 95 percent confidence intervals of parameters (2.5 and 97.5 percentiles) are printed 
+in the summary.
+If inferior to the whole number of iterations, the number of iterations for which the function converges 
+is also printed in the summary.
+
+The plot of an object of class \texttt{bootdist} consists in a scatterplot or a matrix of scatterplots
+of the bootstrapped values of parameters providing
+a representation of the joint uncertainty distribution of the fitted parameters.
+
+Below is an example of the use of this function with the previous of the Weibull distribution to 
+``groundbeef`` data set. 
+<<echo=TRUE, fig=FALSE>>=
+bw<-bootdist(fw,niter=1001)
+plot(bw)
+summary(bw)
+@
+\begin{figure}[htb]
+  \centering
+<<echo=FALSE, results=hide, fig=TRUE>>=
+plot(bw)
+@
+  \caption{Boostrappped values of parameters for a fit of a distribution
+characterized by two parameters on data}
+  \label{bootstrap}
+\end{figure}
+
+Function \texttt{boodistcens} provides the same type of results for fit on continuous distributions to
+censored data.
+<<echo=TRUE, fig=FALSE>>=
+blog10C<-bootdistcens(flog10C,niter=1001)
+summary(blog10C)
 @
 
-\subsection{Quantile matching estimation}
+ADD an example with the Burr distribution characterized by three parameters
+
+\subsection{Bootstrap confidence intervals and regions}
+\label{confidence}
 %%%%%%%%%%
-Quantile matching may also be performed fixing the argument \texttt{method}
-to ``qme'' in the call to fitdist and adding an argument \texttt{probs}
-defining the probabilities for which the quantile matching is done.
-The length of this vector must be equal
-to the number of parameters to estimate.(see the help file page of the function
-\texttt{qmedist} in \texttt{fitdistrplus} reference manual for details).
-Below is an example of fit of a uniform distribution to the same dataset by matching first
-and third quartiles.
-<<>>=
-fitdist(u,"unif",method="qme",probs=c(0.25,0.75))
-@
 
-\subsection{Maximum goodness-of-fit estimation}
+To articulate with previous part
+
+\subsection{Use of bootstrap samples in a second order Monte Carlo simulations}
+\label{MC2D}
 %%%%%%%%%%
-Parameters may also be estimated by maximizing a goodness-of-fit distance
-by fixing the argument \texttt{method}
-to ``mge'' in the call to fitdist and adding an argument \texttt{gof}
-coding for the goodness-of-fit distance chosen. One may use the classical 
-Cramer-von Mises distance (``CvM''), the classical
-Kolmogorov-Smirnov distance (``KS''), the classical Anderson-Darling distance (``AD'')
-which gives more weight to the tails of the distribution,
-or one of the variants of this last distance (see the help file page of the function
-\texttt{mgedist} in \texttt{fitdistrplus} reference manual for details).
-Below is an example of fit of a uniform distribution to the same dataset by maximizing the 
-Kolmogorov-Smirnov distance.
-<<>>=
-fitdist(u,"unif",method="mge",gof="KS")
-@
+TO BE DONE
+
+\clearpage
+
+\section{Conclusion}
+\label{ccl}
+%%%%%%%%%%
+TO BE DONE
+
+\bibliographystyle{plain}
+\bibliography{intro2fitdistrplus}
 

### 2012-01-21 inst/doc/intro2fitdistrplus.Rnw
add scaling problem example, specify encoding for the vignette, update news.

@@
 %\VignetteKeyword{distribution}
+
+% accents 8 bits dans le fichier
+%\usepackage[applemac]{inputenc} %MAC encoding
+%\usepackage[utf8]{inputenc} %UNIX encoding
+\usepackage[ansinew]{inputenc} %WINDOWS encoding
 
@@
 
-% A changer en cas de mise à jour de fitdistrplus !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
+% A changer en cas de mise ‡ jour de fitdistrplus !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
 This paper reviews the various features of version 0.3-0 of \texttt{fitdistrplus}. The package
@@
 of the Hessian matrix at the maximum likelihood solution, correlation matrix
-between parameter estimates, the loglikelihood, the Akaïke and the Schwarz information criteria 
+between parameter estimates, the loglikelihood, the AkaÔke and the Schwarz information criteria 
 (so called AIC and BIC).
@@
 \texttt{xlogscale} to \texttt{TRUE} in the call to \texttt{cdfcomp}.
-% Ajouter ex endosulfan avec la bibliothèque actuar !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
-% mais comment faire dans la vignette : indiquer que fitdistrplus nécessite actuar????????????????????????????
+% Ajouter ex endosulfan avec la bibliothËque actuar !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
+% mais comment faire dans la vignette : indiquer que fitdistrplus nÈcessite actuar????????????????????????????
 

### 2012-03-21 inst/doc/intro2fitdistrplus.Rnw
correction of the problem with optim using "L-BFGS-B" and new version of the vignette

@@
 
-Fitting distributions to data is a very common task in statistics. It consists of choosing
+Fitting distributions to data is a very common task in statistics. It consists in choosing
 a probability distribution that gives a good representation of a statistical
@@
 fitting routine for the parameter estimation step in R. Other steps of
-the process may be developed using R \cite{Ricci05}. Our first objective while developping
+the process may be developed using R \cite{Ricci05}. Our first objective by developping
 package \texttt{fitdistrplus} \cite{fitdistrplus} 
@@
 
-% A changer en cas de mise ‡ jour de fitdistrplus !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
-This paper reviews the various features of version 0.3-0 of \texttt{fitdistrplus}. The package
+% A changer en cas de mise a jour de fitdistrplus !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
+This manuscript reviews the various features of version 0.3-4 of \texttt{fitdistrplus}. The package
 is available from the Comprehensive R Archive Network at \url{http://cran.r-project.org/package=fitdistrplus}.
@@
 corresponding to data sets included in the package.
-% voir si on en met 3 ou 4 au final !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
 
@@
 
-The third data set corresponds to the observation of a discrete variable, 
+The ``Toxocara'' data set corresponds to the observation of a discrete variable, 
 the number of \emph{Toxocara cati} parasites present in digestive tract,
@@
 
-The last data set corresponds to the observation of a continuous censored variable, 
+The ``smoked fish'' data set corresponds to the observation of a continuous censored variable, 
 the \emph{Listeria monocytogenes} microbial concentration, on a random
@@
 @
-  \caption{Density and cdf plots of an empirical distribution for a continuous variable}
+  \caption{Density and cdf plots of an empirical distribution for a continuous variable 
+  (serving size from the ``ground beef'' data set)}
   \label{plotdistcont}
@@
 @
-  \caption{Density and cdf plots of an empirical distribution for a discrete variable}
-  \label{plotdistcont}
+  \caption{Density and cdf plots of an empirical distribution for a discrete variable
+  (number of \emph{Toxocara cati} parasites from the ``Toxocara'' data set)}
+  \label{plotdistdisc}
 \end{figure}
@@
 %%%%%%%%%% 
-Function \texttt{descdist} 
 Descriptives statistics may help the choice of good candidates to describe an empirical distribution
@@
 empirical distribution (Figure~\ref{Cullenplotcont}). 
-On this plot, values for common distributions are displayed as a tools 
+On this plot, values for common distributions are displayed as tools 
 to help the choice of distributions to fit to data. For some distributions (normal, uniform,
@@
 corresponding to bootstrap samples are then computed and reported on the 
-skewness-kurtosis plot.
+skewness-kurtosis plot. Below is a call to function \texttt{descdist} to describe the distribution 
+of the serving size from the ``ground beef'' data set and to draw the corresponding skewness-kurtosis
+plot (Figure~\ref{Cullenplotcont}). Looking at the results on this example
+with a positive skewness and a kurtosis not far from 3,
+the fit of three common right-skewed distributions could be considered, Weibull, gamma and 
+lognormal distributions.
 
@@
 @
-  \caption{Skewness-kurtosis plot for a continuous variable}
+  \caption{Skewness-kurtosis plot for a continuous variable 
+  (serving size from the ``ground beef'' data set)}
   \label{Cullenplotcont}
@@
 
-For discrete variable, skewness and kurtosis values or set of values of Poisson and negative binomial
+For discrete variables, such as the number of \emph{Toxocara cati} parasites from 
+the ``Toxacara'' data set,  skewness and kurtosis values or set of values of Poisson and negative binomial
 distributions are represented in the skewness-kurtosis plot (Figure~\ref{Cullenplotdisc}), 
@@
 normal distribution, to which discrete distributions may converge. 
+Looking at the skewness-kurtosis plot (Figure~\ref{Cullenplotdisc}) obtained
+for the number of \emph{Toxocara cati} parasites from 
+the ``Toxacara'' data set, one could try the fit of Poisson and negative-binomial distributions.
 
@@
 @
-  \caption{Skewness-kurtosis plot for a continuous variable}
+  \caption{Skewness-kurtosis plot for a discrete variable
+  (number of \emph{Toxocara cati} parasites from the ``Toxocara'' data set)}
   \label{Cullenplotdisc}
@@
 
-\section{Fit of a distribution}
+\section{Fit of a distribution by maximum likelihood estimation}
 \label{FIT}
@@
 
+% TO DO : change the formalism of this equation
+
 The other proposed estimation methods are described in Section~\ref{Alternatives}.
@@
 to a data set as an S3 class object that may be easily printed, 
-summarized or plotted (see Figure~\ref{fitgamma} in Section~\ref{GOF plots}). 
+summarized or plotted (see Figure~\ref{plotcontfit} in Section~\ref{GOF plots}). 
 The parametric distribution must be a classically defined R distributions,
@@
 of the Hessian matrix at the maximum likelihood solution, correlation matrix
-between parameter estimates, the loglikelihood, the AkaÔke and the Schwarz information criteria 
+between parameter estimates, the loglikelihood, the Akaike and the Schwarz information criteria 
 (so called AIC and BIC).
+Below is a call to function \texttt{fitdist} to fit a Weibull distribution 
+to the serving size in the ``ground beef'' data set.
 <<echo = TRUE,fig = FALSE>>=
@@
 (x-axis) against the empirical cumulative density function (y-axis)) are also given (\cite{Cullen99}).
-As an exemple, let us look at the plot of the previous fit of a weibull distribution to ``groundbeef'' data set.
+The Q-Q plot emphasizes the lack-of-fit at the distribution tails while the P-P plot emphasizes
+the lack-of-fit at the distribution center.
+As an example, let us look at the plot of the previous fit of a Weibull distribution to the``groundbeef'' data set
+(Figure~\ref{plotcontfit}). The fit is not perfect,
+especially in the center of the distribution, but seems correct while looking at the tails.
 
@@
 @
-  \caption{Plot of the fit of a continuous distribution}
+  \caption{Plot of the fit of a continuous distribution (a Weibull
+  distribution fitted to serving sizes from the ``ground beef'' data set)}
   \label{plotcontfit}
@@
 on a same data set. Function \texttt{cdfcomp} must be called with a first argument 
-corresponding to a list of objects of class \texttt{fitdist}, and optionnaly
+corresponding to a list of objects of class \texttt{fitdist}, and optionaly
 further arguments to customize the plot, as in the following example comparing the fit of
-Weibull, lognormal and gamma distributions to ``groundbeef'' data set.
+Weibull, lognormal and gamma distributions to ``groundbeef'' data set (Figure~\ref{compcdf}).
 <<echo=TRUE, fig=FALSE>>=
@@
 @
-  \caption{Comparison of CDF plots of various distributions fitted on continuous data}
+  \caption{Comparison of CDF plots of various distributions fitted on continuous data
+  (Weibull, gamma and lognormal
+  distributions fitted to serving sizes from the ``ground beef'' data set)}
   \label{compcdf}
@@
 @
-  \caption{Plot of the fit of a discrete distribution}
+  \caption{Plot of the fit of a discrete distribution (a negative binomial distribution
+  fitted to numbers of \emph{Toxocara cati} parasites from the ``Toxocara'' data set)}
   \label{plotdiscfit}
@@
 % ADD the definition of these statistics !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
-EQUATIONS TO BE ADDED
+%EQUATIONS TO BE ADDED
 
@@
 as it is often the case in risk assessment \cite{Cullen99,Vose10}. 
-
 % ADD AN EXAMPLE with comparison !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
@@
 
-Even if specifically recommended for discrete distributions, the Chi-squares statistic may also be used for
+Even if specifically recommended for discrete distributions, the Chi-squared statistic may also be used for
 continuous distributions (see~\cite{fitdistrplus} for examples).
@@
 misinterpretation of p-values is very common and comes from the wrong assumption that
-absence of evidence is evidence of absence~\cite{Altman96}.
+absence of evidence is evidence of absence ~\cite{Altman96}.
 On the contrary, in some cases, especially on very big datasets, 
-even if the null hypothesis is rejected a fitted distribution may be 
+even if the null hypothesis is rejected, a fitted distribution may be 
 chosen as the best one among simple distributions to describe an empirical distribution, if the goodness-of-fit
@@
 the empirical distribution may be plotted using the function \texttt{plotdistcens}.
-Data are reported directly as segments for interval, left and right censored data, 
-and as points for non-censored data. Before plotting, observations are ordered and a rank r
-is associated to each of them. Left censored observations are ordered
-first, by their right bounds. Interval censored and non censored observations
-are then ordered by their mid-points and, at last, right censored observations are
-ordered by their left bounds. If argument \texttt{leftNA} (resp. \texttt{rightNA}) is finite,
-left censored (resp. right censored) observations are considered as interval censored
-observations and ordered by mid-points with non-censored and interval censored data.
-It is sometimes necessary to fix \texttt{leftNA} or \texttt{rightNA} to a realistic 
-extreme value, even if not exactly known, to obtain a reasonable global ranking of 
-observations. After ranking, each of the n observations is plotted as a point (one x-value) 
-or a segment (an interval of possible x-values),
-with an y-value equal to r/n, r being the rank of each observation in the global ordering
-previously described.
+By default this function uses the EM approach of Turnbull \cite{Turnbull74} to compute the overall
+empirical cdf curve with confidence intervals, by calls to functions \texttt{survfit}
+and \texttt{plot.survfit} from the \texttt{survival} package.
+Let us see such a plot for ``smokedfish'' data set after classical transformation
+of microbial counts in decimal logarithm (Figure~\ref{plotdistcens}).
 
-Let us see the sulting plot for ``smokedfish'' data set after classical transformation
-of microbila counts in decimal logarithm.
+% ------------------ previous method -----------------------------
+%Data are reported directly as segments for interval, left and right censored data, 
+%and as points for non-censored data. Before plotting, observations are ordered and a rank r
+%is associated to each of them. Left censored observations are ordered
+%first, by their right bounds. Interval censored and non censored observations
+%are then ordered by their mid-points and, at last, right censored observations are
+%ordered by their left bounds. If argument \texttt{leftNA} (resp. \texttt{rightNA}) is finite,
+%left censored (resp. right censored) observations are considered as interval censored
+%observations and ordered by mid-points with non-censored and interval censored data.
+%It is sometimes necessary to fix \texttt{leftNA} or \texttt{rightNA} to a realistic 
+%extreme value, even if not exactly known, to obtain a reasonable global ranking of 
+%observations. After ranking, each of the n observations is plotted as a point (one x-value) 
+%or a segment (an interval of possible x-values),
+%with an y-value equal to r/n, r being the rank of each observation in the global ordering
+%previously described.
 
@@
 @
-  \caption{CDF plot of a censored data set as}
+  \caption{CDF plot of censored data (microbial counts from 
+   the ``smoked fish'' data set)}
   \label{plotdistcens}
@@
 % ADD the equation !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
-EQUATION TO ADD
 
@@
 %%%%%%%%%%
-Only one goodness-of-fit plot is provided for censored data, correponding to the theoretical 
+Only one goodness-of-fit plot is provided for censored data (Figure~\ref{plotfitdistcens}), corresponding to the theoretical 
 cumulative distribution function added to the plot of censored data presented
@@
   \caption{Goodness-of-fit CDF plot for a fit of a continuous distribution
-  on censored data}
+  on censored data (a lognormal distribution fitted to microbial counts from 
+   the ``smoked fish'' data set)}
   \label{plotfitdistcens}
@@
 % ADD TABLE with formulas !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
-TABLE with formulas TO BE ADDED
 
@@
 to fit the parametric distribution by giving more weight to the left tail of the empirical distribution
-such as in the following example using left tail Anderson-Darling distances of first or second order.
+such as in the following example using left tail Anderson-Darling distances of first or second order
+(Figure~\ref{plotfitMGE}).
 
@@
 @
-  \caption{Comparison of lognormal distributions fitted by maximum goodness-of-fit using
-  various goodness-of-fit distances}
+  \caption{Comparison of one distribution fitted by maximum goodness-of-fit using
+  various goodness-of-fit distances (a lognormal distribution fitted to acute toxicity values
+  from the ``endosulfan'' data set)}
   \label{plotfitMGE}
@@
 Each time a numerical minimization (or maximization) is carried out using \texttt{fitdist}, 
-Function \texttt{optim}of the package \texttt{stats} is used by default with the 
+Function \texttt{optim} of the package \texttt{stats} is used by default with the 
 ``Nelder-Mead'' method for distributions characterized by more than one parameter and
@@
 
-Here are examples of fits of a gamma distribution to ``ground beeef'' data set with various options
+Here are examples of fits of a gamma distribution to ``ground beef'' data set with various options
 of \texttt{optim}.
@@
 fitdist(groundbeef$serving,"gamma",optim.method="BFGS") 
-fitdist(groundbeef$serving,"gamma",optim.method="L-BFGS-B",lower=c(0,0))
 fitdist(groundbeef$serving,"gamma",optim.method="SANN")
@@
 
-\section{Uncertainty on parameter estimates}
+\section{Uncertainty in parameter estimates}
 \label{Uncertainty}
@@
 of the bootstrapped values of parameters providing
-a representation of the joint uncertainty distribution of the fitted parameters.
+a representation of the joint uncertainty distribution of the fitted parameters (Figure~\ref{bootstrap}).
 
@@
 @
-  \caption{Boostrappped values of parameters for a fit of a distribution
-characterized by two parameters on data}
+  \caption{Bootstrappped values of parameters for a fit of a distribution
+characterized by two parameters (example on the fit of a Weibull distribution to
+serving sizes from the ``ground beef'' data set)}
   \label{bootstrap}
@@
 
-ADD an example with the Burr distribution characterized by three parameters
+% ADD an example with the Burr distribution characterized by three parameters
 
-\subsection{Bootstrap confidence intervals and regions}
-\label{confidence}
+%\subsection{Bootstrap confidence intervals and regions}
+%\label{confidence}
 %%%%%%%%%%
 
-To articulate with previous part
+%To articulate with previous part
 
-\subsection{Use of bootstrap samples in a second order Monte Carlo simulations}
-\label{MC2D}
+\subsection{Use of bootstrap samples}
+%\label{MC2D}
 %%%%%%%%%%
-TO BE DONE
+%TO BE DONE
+Bootstrap samples of parameter estimates may be used to calculate confidence intervals on each parameter
+of the fitted distribution, but it is also interesting to look at the marginal distribution of the bootstrap
+values in a scatterplot (or a matrix of scatterplots if the number of parameters exceeds two), and especially
+to look at the potential structural correlation between parameters.
+
+Moreover, the use of the whole bootstrap sample is of great interest in the risk assessment field. Its use enables
+the characterization of uncertainty in distribution parameters. It can be directly used within a second order Monte Carlo
+simulation framework, especially within the package \texttt{mc2d} (\cite{mc2d}). One could refer to Pouillot \emph{et al.}
+(\cite{Pouillot10}) for an introduction to the use of \texttt{mc2d} 
+and \texttt{fitdistrplus} packages in the context of quantitative 
+risk assessment.
 
@@
 
-\section{Conclusion}
-\label{ccl}
+%\section{Conclusion}
+%\label{ccl}
 %%%%%%%%%%
-TO BE DONE
+%TO BE DONE
 

### 2012-07-26 R/denscomp.R
remove warnings, minor code/man update and vignettes update.

@@ -26,7 +26,7 @@
 
 
 denscomp <- function(ft, xlim, ylim, probability, main, xlab, ylab, datapch, datacol, fitlty, fitcol, 
-	addlegend = TRUE, legendtext, xlegend = "bottomright", ylegend = NULL, 
+	addlegend = TRUE, legendtext, xlegend = "topright", ylegend = NULL, 
 	demp = FALSE, dempcol = "grey", ...)
 {
 	if(inherits(ft, "fitdist"))
@@ -78,9 +78,7 @@ denscomp <- function(ft, xlim, ylim, probability, main, xlab, ylab, datapch, dat
     }
 	if (missing(ylim))
 	{
-		options(warn=-1)
-		reshist <- hist(mydata, plot=FALSE, probability=probability)
-		options(warn=getOption("warn"))
+		reshist <- hist(mydata, plot=FALSE)
 		if(!probability)
 			ylim <- c(0, max(reshist$counts))
 		else

### 2012-07-26 R/quantiles.R

@@ -25,72 +25,77 @@
 ### 
 
 
-quantile.fitdist <- function(x, probs = seq(0.1,0.9,0.1), bootstrap = TRUE, CI.type = "two.sided",
+quantile.fitdist <- function(x, probs = seq(0.1, 0.9, 0.1), bootstrap = TRUE, CI.type = "two.sided", 
                 bootstrap.arg = list(bootmethod="param", niter=1001), ...)
 {
     if (!inherits(x, "fitdist"))
         stop("Use only with 'fitdist' objects")
     cens <- FALSE
-    quantiles(f = x,probs = probs,bootstrap = bootstrap,CI.type = CI.type,bootstrap.arg = bootstrap.arg,cens)
+    myquantiles(f = x, probs = probs, bootstrap = bootstrap, CI.type = CI.type, 
+				bootstrap.arg = bootstrap.arg, cens)
 }
 
-quantile.fitdistcens <- function(x, probs = seq(0.1,0.9,0.1), bootstrap = TRUE, CI.type = "two.sided",
+quantile.fitdistcens <- function(x, probs = seq(0.1, 0.9, 0.1), bootstrap = TRUE, CI.type = "two.sided", 
                 bootstrap.arg = list(niter=1001), ...)
 {
     if (!inherits(x, "fitdistcens"))
         stop("Use only with 'fitdistcens' objects")
     cens <- TRUE
-    quantiles(f = x,probs = probs,bootstrap = bootstrap,CI.type = CI.type,bootstrap.arg = bootstrap.arg,cens)
+    myquantiles(f = x, probs = probs, bootstrap = bootstrap, CI.type = CI.type, 
+				bootstrap.arg = bootstrap.arg, cens)
 }
 
-quantiles <- function(f, probs, bootstrap, CI.type, bootstrap.arg, cens )
+myquantiles <- function(f, probs, bootstrap, CI.type, bootstrap.arg, cens )
 {    
-    CI.type <- match.arg(CI.type, c("two.sided","less","greater"))
+    CI.type <- match.arg(CI.type, c("two.sided", "less", "greater"))
     
-    qdistname<-paste("q",f$distname,sep="")
-    if (!exists(qdistname,mode="function"))
-        stop(paste("The ",qdistname," function must be defined")) 
+    qdistname<-paste("q", f$distname, sep="")
+    if (!exists(qdistname, mode="function"))
+        stop(paste("The ", qdistname, " function must be defined")) 
      
[…]

### 2012-07-26 tests/fitdist.R

@@ -204,7 +204,7 @@ gofstat(fuKS)
 # (14) scaling problem
 #
 
-x <- c(-0.00707717, -0.000947418, -0.00189753, 
+x2 <- c(-0.00707717, -0.000947418, -0.00189753, 
 -0.000474947, -0.00190205, -0.000476077, 0.00237812, 0.000949668, 
 0.000474496, 0.00284226, -0.000473149, -0.000473373, 0, 0, 0.00283688, 
 -0.0037843, -0.0047506, -0.00238379, -0.00286807, 0.000478583, 
@@ -224,5 +224,44 @@ x <- c(-0.00707717, -0.000947418, -0.00189753,
 -0.00241138, -0.00144963)
 
 for(i in 6:0)
-cat(i, try(fitdist(x*10^i, "cauchy", method="mle")$estimate, silent=TRUE), "\n")
+	cat(i, try(fitdist(x2*10^i, "cauchy", method="mle")$estimate, silent=TRUE), "\n")
+
+
+# (15) Fit of a lognormal distribution on acute toxicity values of endosulfan for
+# nonarthropod invertebrates, using maximum likelihood estimation
+# to estimate what is called a species sensitivity distribution 
+# (SSD) in ecotoxicology, followed by estimation of the 5 percent quantile value of 
+# the fitted distribution, what is called the 5 percent hazardous concentration (HC5)
+# in ecotoxicology, with its two-sided 95 percent confidence interval calculated by 
+# parametric bootstrap
+#
+data(endosulfan)
+ATV <- subset(endosulfan, group == "NonArthroInvert")$ATV
+log10ATV <- log10(subset(endosulfan, group == "NonArthroInvert")$ATV)
+fln <- fitdist(log10ATV, "norm")
+
+quantile(fln, probs = 0.05, bootstrap=TRUE, 
+	bootstrap.arg = list(bootmethod = "param", niter = 501))
+
+quantile(fln, probs = c(0.05,0.1,0.2), bootstrap = TRUE, CI.type = "greater",
+	bootstrap.arg = list(bootmethod = "param", niter = 501))
+
+quantile(fln, probs = 0.05, bootstrap=TRUE, 
+	bootstrap.arg = list(bootmethod = "nonparam", niter = 501))
+
+
+
+# (16) uniform distribution
+#
+
+x3 <- runif(1000, 1/2, 5/2)
+fitdist(x3, "unif", method="mle")
[…]

### 2012-11-07 tests/mledist.R
update vignette. first stable version.

@@ -127,19 +127,20 @@ if(any(installed.packages()[,"Package"] == "actuar"))
 
 
 
-# (10) custom optim
+# (10) custom optim for exponential distribution
 #
 if(any(installed.packages()[,"Package"] == "rgenoud"))
 {
 	
 
-mysample <- rexp(100, 5)
+mysample <- rexp(1000, 5)
 mystart <- 8
 
-mledist(mysample, "exp", optim.method="Nelder-Mead") 
-mledist(mysample, "exp", optim.method="BFGS") 
-mledist(mysample, "exp", optim.method="L-BFGS-B", lower=0) 
-mledist(mysample, "exp", optim.method="SANN") 
+fNM <- mledist(mysample, "exp", optim.method="Nelder-Mead") 
+fBFGS <- mledist(mysample, "exp", optim.method="BFGS") 
+fLBFGSB <- mledist(mysample, "exp", optim.method="L-BFGS-B", lower=0) 
+fSANN <- mledist(mysample, "exp", optim.method="SANN") 
+fCG <- mledist(mysample, "exp", optim.method="CG") 
 
 #the warning tell us to use optimise...
 
@@ -150,7 +151,7 @@ myoptimize <- function(fn, par, ...)
 	c(res, convergence=0, value=res$objective, par=res$minimum, hessian=NA)
 }
 
-mledist(mysample, "exp", start=mystart, custom.optim=myoptimize, interval=c(0, 100))
+foptimize <- mledist(mysample, "exp", start=mystart, custom.optim=myoptimize, interval=c(0, 100))
 
 
 library(rgenoud)
@@ -163,14 +164,69 @@ mygenoud <- function(fn, par, ...)
 }
 
 
-mledist(mysample, "exp", start=mystart, custom.optim= mygenoud, nvars=1,    
+fgenoud <- mledist(mysample, "exp", start=mystart, custom.optim= mygenoud, nvars=1,    
         Domains=cbind(0, 10), boundary.enforcement=1, 
         hessian=TRUE, print.level=0)
 
 try( mledist(mysample, "exp", start=mystart, custom.optim= mygenoud) )
 
+c(NM=fNM$estimate,
+BFGS=fBFGS$estimate,
+LBFGSB=fLBFGSB$estimate,
+SANN=fSANN$estimate,
+CG=fCG$estimate,
+optimize=foptimize$estimate,
[…]

### 2012-11-23 tests/mledist.R
improve the computation time of the vignette.

@@ -213,12 +213,14 @@ fNM <- mledist(groundbeef$serving, "gamma", optim.method="Nelder-Mead")
 fBFGS <- mledist(groundbeef$serving, "gamma", optim.method="BFGS") 
 fLBFGSB <- mledist(groundbeef$serving, "gamma", optim.method="L-BFGS-B", lower=0) 
 fSANN <- mledist(groundbeef$serving, "gamma", optim.method="SANN") 
-fCG <- mledist(groundbeef$serving, "gamma", optim.method="CG", control=list(maxit=10000)) 
+fCG <- try( mledist(groundbeef$serving, "gamma", optim.method="CG", control=list(maxit=10000)) )
+if(class(fCG) == "try-error")
+	fCG <- list(estimate=NA)
 
-fgenoud <- mledist(groundbeef$serving, "gamma", start=mystart, 
-		custom.optim= mygenoud, nvars=2,    
-        Domains=cbind(c(0,0), c(100,100)), boundary.enforcement=1, 
-        hessian=TRUE, print.level=0)
+fgenoud <- mledist(groundbeef$serving, "gamma", 
+		custom.optim= mygenoud, nvars=2, max.generations=10,
+		Domains=cbind(c(0,0), c(10,10)), boundary.enforcement=1, 
+        hessian=TRUE, print.level=0, P9=10)
 
 cbind(NM=fNM$estimate,
 BFGS=fBFGS$estimate,

### 2015-08-19 R/bootdist.R
update starting/fixing values in mledist.

@@ -47,7 +47,7 @@ bootdist <- function (f, bootmethod="param", niter=1001)
     
     #compute bootstrap estimates
     foncestim <- switch(f$method, "mle"=mledist, "qme"=qmedist, "mme"=mmedist, "mge"=mgedist)
-    start <- f$estimate
+    start <- as.list(f$estimate) #a named vector is no longer is accepted as starting values.
     if (is.null(f$dots))
         func <- function(iter) {
             res <- do.call(foncestim, list(data=rdata[, iter], distr=f$distname, start=start, fix.arg=f$fix.arg))

### 2015-08-19 tests/bootdist.R

@@ -100,7 +100,7 @@ if(any(installed.packages()[, "Package"] == "actuar"))
     ifelse(order == 1,  mean(x),  sum(x^order)/length(x))
     
     f4 <- fitdist(x4,  "pareto",  "mme",  order=1:2,  
-                  start=c(shape=10,  scale=10),  
+                  start=list(shape=10,  scale=10),  
                   lower=1,  memp="memp",  upper=50)
     
     b4 <- bootdist(f4,  niter=nbboot)
@@ -119,7 +119,7 @@ if(any(installed.packages()[, "Package"] == "actuar"))
     data(danishuni)
 
 fdan <- fitdist(danishuni$Loss, "burr", method="mle", 
-    start=c(shape1=5, shape2=5, rate=10), lower=0+1e-1, control=list(trace=0))
+    start=list(shape1=5, shape2=5, rate=10), lower=0+1e-1, control=list(trace=0))
 bdan <- bootdist(fdan,  bootmethod="param", niter=nbboot)
 summary(bdan)
 plot(bdan)

### 2015-08-19 tests/fitdistcens.R

@@ -76,7 +76,8 @@ fitdistcens(log10EC50,"logis",optim.method="SANN")
   }
   
   # call fitdistcens with a 'custom' optimization function
-  fit.with.genoud <- fitdistcens(log10EC50, "logis", custom.optim=mygenoud, nvars=2,    
+  fit.with.genoud <- fitdistcens(log10EC50, "logis", custom.optim=mygenoud, nvars=2, 
+                                 start=list(location=0, scale=1),    
                                  Domains=cbind(c(0,0), c(5, 5)), boundary.enforcement=1, 
                                  print.level=1, hessian=TRUE)
   
@@ -86,7 +87,7 @@ fitdistcens(log10EC50,"logis",optim.method="SANN")
 # (7) estimation of the mean of a normal distribution 
 # by maximum likelihood with the standard deviation fixed at 1 using the argument fix.arg
 #
-flnb <- fitdistcens(log10EC50, "norm", start = list(mean = 1),fix.arg = list(sd = 1))
+flnb <- fitdistcens(log10EC50, "norm", start = list(mean = 1), fix.arg = list(sd = 1))
 
 # (8) Fit of a lognormal distribution on acute toxicity values of fluazinam for
 # macroinvertebrates and zooplancton, using maximum likelihood estimation
@@ -111,5 +112,6 @@ x <- data.frame(left=x, right=x)
 f1 <- fitdistcens(x, "exp", keepdata=FALSE)
 f2 <- fitdistcens(x, "exp", keepdata=TRUE)
  
+
 plot(f1)
 plot(f2)  

### 2015-08-19 tests/mledist.R

@@ -94,24 +94,24 @@ cat(i, try(mledist(x*10^i, "cauchy")$estimate, silent=TRUE), "\n")
 
 #mixture of two normal distributions
 #density
-dnorm2 <- function(x, poid, m1, s1, m2, s2)
-	poid*dnorm(x, m1, s1) + (1-poid)*dnorm(x, m2, s2)
+dnorm2 <- function(x, w, m1, s1, m2, s2)
+	w*dnorm(x, m1, s1) + (1-w)*dnorm(x, m2, s2)
 #numerically-approximated quantile function
-qnorm2 <- function(p, poid, m1, s1, m2, s2)
+qnorm2 <- function(p, w, m1, s1, m2, s2)
 {
 	L2 <- function(x, prob)
-		(prob - pnorm2(x, poid, m1, s1, m2, s2))^2	
+		(prob - pnorm2(x, w, m1, s1, m2, s2))^2	
 	sapply(p, function(pr) optimize(L2, c(-20, 30), prob=pr)$minimum)
 }	
 #distribution function		
-pnorm2 <- function(q, poid, m1, s1, m2, s2)
-	poid*pnorm(q, m1, s1) + (1-poid)*pnorm(q, m2, s2)		
+pnorm2 <- function(q, w, m1, s1, m2, s2)
+	w*pnorm(q, m1, s1) + (1-w)*pnorm(q, m2, s2)		
 
 
 #basic normal distribution
 x <- c(rnorm(1000, 5),  rnorm(1000, 10))
 #MLE fit
-fit1 <- mledist(x, "norm2", start=list(poid=1/3, m1=4, s1=2, m2=8, s2=2), 
+fit1 <- mledist(x, "norm2", start=list(w=1/3, m1=4, s1=2, m2=8, s2=2), 
 	lower=c(0, 0, 0, 0, 0))
 
 
@@ -119,7 +119,7 @@ fit1 <- mledist(x, "norm2", start=list(poid=1/3, m1=4, s1=2, m2=8, s2=2),
 
 
 
-# (10) fit a Pareto distribution
+# (10) fit a Pareto and log-logistic distributions
 #
 
 if(any(installed.packages()[,"Package"] == "actuar"))
@@ -129,8 +129,15 @@ if(any(installed.packages()[,"Package"] == "actuar"))
     x4 <- rpareto(1000, 6, 2)
 	
 #fit
-    mledist(x4, "pareto", start=c(shape=10, scale=10), lower=1, upper=Inf)
[…]

### 2015-12-08 R/mledist.R
add counts component in ***dist functions, remove a space in vignettes title, set the seed in the test file.

@@ -220,7 +220,7 @@ mledist <- function (data, distr, start=NULL, fix.arg=NULL, optim.method="defaul
           names(opt$par) <- names(vstart)
         res <- list(estimate = opt$par, convergence = opt$convergence, loglik = -opt$value, 
                     hessian = opt$hessian, optim.function="optim", fix.arg = fix.arg, 
-                    optim.method=meth, fix.arg.fun = fix.arg.fun, weights = weights)
+                    optim.method=meth, fix.arg.fun = fix.arg.fun, weights = weights, counts=opt$counts)
     }
     else # Try to minimize the minus (log-)likelihood using a user-supplied optim function 
     {
@@ -256,7 +256,7 @@ mledist <- function (data, distr, start=NULL, fix.arg=NULL, optim.method="defaul
           names(opt$par) <- names(vstart)
         res <- list(estimate = opt$par, convergence = opt$convergence, loglik = -opt$value, 
                       hessian = opt$hessian, optim.function = custom.optim, fix.arg = fix.arg,
-                      method = method.cust, fix.arg.fun = fix.arg.fun, weights = weights)        
+                      method = method.cust, fix.arg.fun = fix.arg.fun, weights = weights, counts=opt$counts)        
     }   
         
     return(res) 

### 2015-12-08 R/mmedist.R

@@ -261,7 +261,8 @@ mmedist <- function (data, distr, order, memp, start=NULL, fix.arg=NULL,
             if(is.null(names(opt$par)))
               names(opt$par) <- names(vstart)
             res <- list(estimate = opt$par, convergence = opt$convergence, value = opt$value, 
-                        hessian = opt$hessian, optim.function="optim", order=order, memp=memp)  
+                        hessian = opt$hessian, optim.function="optim", order=order, 
+                        memp=memp, counts=opt$counts)  
             
         }else # Try to minimize the stat distance using a user-supplied optim function 
         {
@@ -289,7 +290,8 @@ mmedist <- function (data, distr, order, memp, start=NULL, fix.arg=NULL,
             }
             
             res <- list(estimate = opt$par, convergence = opt$convergence, value = opt$value, 
-                        hessian = opt$hessian, optim.function=custom.optim, order=order, memp=memp)  
+                        hessian = opt$hessian, optim.function=custom.optim, order=order, 
+                        memp=memp, counts=opt$counts)  
             
         }   
         

### 2015-12-08 tests/test-lnL-surf.R

@@ -1,6 +1,8 @@
 library(fitdistrplus)
 
 
+
+
 #(1) beta distribution
 #
 
@@ -25,3 +27,49 @@ llcurve(plot.min=0.1, plot.max=7, plot.arg="shape1", fix.arg=list(shape2=3/4), p
                       obs=x, distr="beta")
 llcurve(plot.min=0.1, plot.max=2, plot.arg="shape2", fix.arg=list(shape1=3), plot.np=50, 
                       obs=x, distr="beta")
+
+
+
+
+#test
+
+psi <- function(x) digamma(x)
+
+grbetalnl <- function(x, a, b)
+  c(log(x)-psi(a)+psi(a+b), log(1-x)-psi(b)+psi(a+b))
+
+grbetalnl(x, 3, 4)
+
+grlnL <- function(par, obs, ...) 
+  -rowSums(sapply(obs, function(x) grbetalnl(x, a=par[1], b=par[2])))
+
+rowSums(sapply(x, function(x) grbetalnl(x, 3, 4)))
+grlnL(c(3, 4), x)
+grlnL(c(3, 3/4), x)
+
+
+ctr <- list(trace=0, REPORT=1, maxit=1000)
+
+bfgs_gr <- mledist(x, dist="beta", optim.method="BFGS", gr=grlnL, control=ctr)
+bfgs <- mledist(x, dist="beta", optim.method="BFGS", control=ctr)
+
+cg_gr <- mledist(x, dist="beta", optim.method="CG", gr=grlnL, control=ctr)
+cg <- mledist(x, dist="beta", optim.method="CG", control=ctr)
+
+nm_gr <- mledist(x, dist="beta", optim.method="Nelder", gr=grlnL, control=ctr)
+nm <- mledist(x, dist="beta", optim.method="Nelder", control=ctr)
+
+getval <- function(x)
+  c(x$estimate, loglik=x$loglik, x$counts)
+
+cbind(NM=getval(nm), NMgrad=getval(nm_gr), CG=getval(cg), 
+      CGgrad=getval(cg_gr), BFGS=getval(bfgs), BFGSgrad=getval(bfgs_gr))
+
+
+
+llsurface(plot.min=c(0.1, 0.1), plot.max=c(7, 3), plot.arg=c("shape1", "shape2"),
+          plot.np=50, obs=x, distr="beta", plot.type="contour")
+points(bfgs$estimate[1], bfgs$estimate[2], pch="+", col="red")
+points(3, 3/4, pch="x", col="green")
+
+

### 2015-12-14 tests/test-cvg-algo.R
start convergence test file.

@@ -1,0 +1,109 @@
+library(fitdistrplus)
+
+
+
+
+#(1) beta distribution
+#
+
+n <- 100
+set.seed(12345)
+x <- rbeta(n, 3, 3/4)
+
+psi <- function(x) digamma(x)
+grbetalnl <- function(x, a, b)
+  c(log(x)-psi(a)+psi(a+b), log(1-x)-psi(b)+psi(a+b))
+#grbetalnl(x, 3, 4)
+grlnL <- function(par, obs, ...) 
+  -rowSums(sapply(obs, function(x) grbetalnl(x, a=par[1], b=par[2])))
+#rowSums(sapply(x, function(x) grbetalnl(x, 3, 4)))
+#grlnL(c(3, 4), x)
+#grlnL(c(3, 3/4), x)
+
+
+constrOptim2 <- function(par, fn, gr=NULL, ui, ci, ...)
+  constrOptim(theta=unlist(par), f=fn, grad=gr, ui=ui, ci=ci, ...)
+
+
+
+#control parameters
+ctr <- list(trace=3, REPORT=1, maxit=1000)
+ctr <- list(trace=0, REPORT=1, maxit=1000)
+
+bfgs_gr$time <- system.time(bfgs_gr <- mledist(x, dist="beta", optim.method="BFGS", gr=grlnL, control=ctr))[3]
+bfgs <- mledist(x, dist="beta", optim.method="BFGS", control=ctr)
+
+lbfgs_gr <- mledist(x, dist="beta", optim.method="L-BFGS-B", gr=grlnL, control=ctr, lower=c(0,0))
+lbfgs <- mledist(x, dist="beta", optim.method="L-BFGS-B", control=ctr, lower=c(0,0))
+
+cg_gr <- mledist(x, dist="beta", optim.method="CG", gr=grlnL, control=ctr)
+cg <- mledist(x, dist="beta", optim.method="CG", control=ctr)
+
+nm_gr <- mledist(x, dist="beta", optim.method="Nelder", gr=grlnL, control=ctr)
+nm <- mledist(x, dist="beta", optim.method="Nelder", control=ctr)
+
+constr_nm_gr <- mledist(x, dist="beta", custom.optim=constrOptim2,
+      ui = diag(2), ci = c(0, 0), optim.method="Nelder", gr=grlnL, control=ctr)
+constr_nm <- mledist(x, dist="beta", custom.optim=constrOptim2,
+      ui = diag(2), ci = c(0, 0), optim.method="Nelder", control=ctr)
+
[…]

### 2016-04-14 tests/test-fitbench.R
change name of the vignette.

@@ -1,6 +1,8 @@
 require(fitdistrplus)
 
-x <- rgamma(1e3, shape=1/2, rate= 1/2)
+x <- rgamma(1e3, shape=3/2, rate= 1/2)
+
+initval <- unlist(fitdistrplus:::start.arg.default(x, "gamma"))
 
 fitdistrplus:::fitbench(x, "gamma", "mle")
 

### 2016-04-14 tests/test-cvg-algo.R
modification of functions llsurface and llcurve

@@ -101,8 +101,8 @@ cbind(trueval=c(3, 3/4, lnL(c(3, 3/4), NULL, x, "dbeta"), NA, NA),
 
 
 
-llsurface(plot.min=c(0.1, 0.1), plot.max=c(7, 3), plot.arg=c("shape1", "shape2"),
-          plot.np=50, obs=x, distr="beta", plot.type="contour")
+llsurface(min.arg=c(0.1, 0.1), max.arg=c(7, 3), plot.arg=c("shape1", "shape2"),
+          lseq=50, data=x, distr="beta")
 points(bfgs$estimate[1], bfgs$estimate[2], pch="+", col="red")
 points(3, 3/4, pch="x", col="green")
 

### 2016-04-14 tests/test-lnL-surf.R

@@ -2,31 +2,23 @@ library(fitdistrplus)
 
 
 
-
 #(1) beta distribution
 #
 
 n <- 100
 x <- rbeta(n, 3, 3/4)
 
-
-
-par(mfrow=c(1, 3))
-llsurface(plot.min=c(0.1, 0.1), plot.max=c(7, 3), plot.arg=c("shape1", "shape2"),
-                         plot.np=50, obs=x, distr="beta", plot.type="persp", theta=60, phi=30)
-llsurface(plot.min=c(0.1, 0.1), plot.max=c(7, 3), plot.arg=c("shape1", "shape2"),
-                         plot.np=50, obs=x, distr="beta", plot.type="contour")
+llsurface(data = x, distr = "beta", plot.arg=c("shape1", "shape2"),
+          min.arg=c(0.1, 0.1), max.arg=c(7, 3))
+llsurface(data = x, distr = "beta", plot.arg=c("shape1", "shape2"),
+          min.arg=c(0.1, 0.1), max.arg=c(7, 3), col = FALSE )
 points(3, 3/4, pch="+", col="red")
-llsurface(plot.min=c(0.1, 0.1), plot.max=c(7, 3), plot.arg=c("shape1", "shape2"),
-                         plot.np=50, obs=x, distr="beta", plot.type="image")
 
 
-
-par(mfrow=c(1,2))
-llcurve(plot.min=0.1, plot.max=7, plot.arg="shape1", fix.arg=list(shape2=3/4), plot.np=50, 
-                      obs=x, distr="beta")
-llcurve(plot.min=0.1, plot.max=2, plot.arg="shape2", fix.arg=list(shape1=3), plot.np=50, 
-                      obs=x, distr="beta")
+llcurve(data = x, distr = "beta", plot.arg = "shape1", min.arg = 0.1, max.arg = 7, 
+        fix.arg = list(shape2 = 3/4), lseq=100, col = "blue")
+llcurve(data = x, distr = "beta", plot.arg = "shape2", min.arg = 0.1, max.arg = 7, 
+        fix.arg = list(shape1 = 3), lseq=100, col = "red")
 
 
 
@@ -67,8 +59,8 @@ cbind(NM=getval(nm), NMgrad=getval(nm_gr), CG=getval(cg),
 
 
 
-llsurface(plot.min=c(0.1, 0.1), plot.max=c(7, 3), plot.arg=c("shape1", "shape2"),
[…]

### 2016-04-14 R/fitbench.R
remove warning and note in R CMD check.

@@ -76,12 +76,12 @@ fitbench <- function(data, distr, method, grad=NULL, control=list(trace=0, REPOR
   {
     for(meth in c("BFGS", "Nelder", "CG")) #CG with FR update
     {
-      res1fit$time <- system.time(res1fit <- mledist(x, distr=distr, optim.method=meth, control=control, lower=lower, upper=upper, gradient=grad, ...))[3]
+      res1fit$time <- system.time(res1fit <- mledist(data, distr=distr, optim.method=meth, control=control, lower=lower, upper=upper, gradient=grad, ...))[3]
       reslist <- c(reslist, list(res1fit))
     }
     for(type in 2:3) #CG with PR or BS updates
     {
-      res1fit$time <- system.time(res1fit <- mledist(x, distr=distr, optim.method="CG", control=c(control, type=type), lower=lower, upper=upper, gradient=grad, ...))[3] 
+      res1fit$time <- system.time(res1fit <- mledist(data, distr=distr, optim.method="CG", control=c(control, type=type), lower=lower, upper=upper, gradient=grad, ...))[3] 
       reslist <- c(reslist, list(res1fit)) 
     }
     fullname <- c(fullname, paste0("G-", c("BFGS", "NM", paste0("CG", c("FR", "PR", "BS"))), "-B") )
@@ -100,21 +100,12 @@ fitbench <- function(data, distr, method, grad=NULL, control=list(trace=0, REPOR
   resmat <- sapply(reslist, getval)
   if(is.null(dim(resmat)))
     stop("wrong extract")
-    
-  print(resmat)
-  
-#   resNM <- sapply(reslist[grep("NM", names(reslist))], getval)
-#   resCGFR <- sapply(reslist[grep("CGFR", names(reslist))], getval)
-#   resCGPR <- sapply(reslist[grep("CGPR", names(reslist))], getval)
-#   resCGBS <- sapply(reslist[grep("CGBS", names(reslist))], getval)  
-#   resBFGS <- sapply(reslist[grep("BFGS", names(reslist))], getval)  
   
   allname <- c(paste("fitted", names(reslist[[1]]$estimate)), "fitted loglik", "func. eval. nb.", "grad. eval. nb.", "time (sec)")
   if(NROW(resmat) != length(allname))
     stop("wrong extract")
-    
[…]

### 2016-04-15 R/util-transform.R
start prefitmle() function and update the FAQ.

@@ -1,0 +1,56 @@
+#############################################################################
+#   Copyright (c) 2009 Marie Laure Delignette-Muller, Christophe Dutang                                                                                                  
+#                                                                                                                                                                        
+#   This program is free software; you can redistribute it and/or modify                                               
+#   it under the terms of the GNU General Public License as published by                                         
+#   the Free Software Foundation; either version 2 of the License, or                                                   
+#   (at your option) any later version.                                                                                                            
+#                                                                                                                                                                         
+#   This program is distributed in the hope that it will be useful,                                                             
+#   but WITHOUT ANY WARRANTY; without even the implied warranty of                                          
+#   MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the                                 
+#   GNU General Public License for more details.                                                                                    
+#                                                                                                                                                                         
+#   You should have received a copy of the GNU General Public License                                           
+#   along with this program; if not, write to the                                                                                           
+#   Free Software Foundation, Inc.,                                                                                                              
+#   59 Temple Place, Suite 330, Boston, MA 02111-1307, USA                                                             
+#                                                                                                                                                                         
+#############################################################################
+### Infinitely differentiable transformations of R into a bounded or half-bounded interval
+###
+###         R functions
+### 
+
+
+#inverse are useless?
+
+
+#Transformation from (-Inf, +Inf) to (-1, 0)
+Tm10 <- function(x)
+  -1/(1+exp(-x))
+#Inverse
+iTm10 <- function(x) 
+  log(-x/(1+x))
+
+#Transformation from (-Inf, +Inf) to (0, 1)
+T01 <- function(x)
+  1/(1+exp(-x))
+#Inverse
+iT01 <- function(x)
+  log(x/(1-x))
+
+#Transformation from (-Inf, +Inf) to (0, +Inf)
+T0Inf <- function(x)
+  exp(x)
+#Inverse
+iT0Inf <- function(x)
+  log(x)
+
+#Transformation from (-Inf, +Inf) to (1, +Inf)
+T1Inf <- function(x)
+  1+exp(x)
+#Inverse
+iT1Inf <- function(x)
+  log(x-1)
+

### 2016-04-15 tests/test-prefit.R

@@ -1,0 +1,10 @@
+require(fitdistrplus)
+
+x <- rgamma(1e3, 5/2, 7/2)
+
+
+fitdistrplus:::prefitmle(x, "dgamma", c(shape=3, scale=3), lower=-Inf, upper=Inf)
+
+fitdistrplus:::prefitmle(x, "dgamma", c(shape=log(3), scale=log(3)), lower=0, upper=Inf)
+
+fitdistrplus:::prefitmle(x, "dgamma", c(shape=log(3)), fix.arg=list(scale=7/2), lower=-Inf, upper=Inf)

### 2016-05-25 tests/fitdist.R
first update of the FAQ.

@@ -323,7 +323,7 @@ fKS <- fitdist(t,"triang",method="mge",start = list(min=4, mode=6,max=9),gof="KS
 cdfcomp(list(fCvM,fKS))
 }
 
-# (18) uniform distribution
+# (18) gumbel distribution
 #
 dgumbel <- function(x, a, b) 1/b*exp((a-x)/b)*exp(-exp((a-x)/b))
 pgumbel <- function(q, a, b) exp(-exp((a-q)/b))
@@ -399,6 +399,8 @@ x2 <- rzmgeom(1000, 1/2, 1/10)
 table(x2)
 #this is the MLE which converges almost surely and in distribution to the true value.
 initp1 <- function(x) list(p1=mean(x == 0))
+
+fitdist(x2, "zmgeom", start=list(p1=1/2, p2=1/2))
 
 f2 <- fitdist(x2, "zmgeom", fix.arg=initp1, start=list(p2=1/2))
 print(f2)

### 2016-06-16 R/util-startarg.R
add tests in start.arg.default,  minor modifications in the FAQ and todo list update

@@ -24,8 +24,12 @@ start.arg.default <- function(x, distr)
   }else if (distr == "pois") {
     start <- list(lambda=mean(x))
   }else if (distr == "exp") {
+    if (any(x < 0)) 
+      stop("values must be positive to fit an exponential  distribution")
     start <- list(rate=1/mean(x))
   }else if (distr == "gamma") {
+    if (any(x < 0)) 
+      stop("values must be positive to fit an gamma  distribution")
     n <- length(x)
     m <- mean(x)
     v <- (n - 1)/n*var(x)
@@ -51,6 +55,8 @@ start.arg.default <- function(x, distr)
     aux <- m*(1-m)/v - 1
     start <- list(shape1=m*aux, shape2=(1-m)*aux)
   }else if (distr == "weibull") {
+    if (any(x < 0)) 
+      stop("values must be positive to fit an exponential  distribution")
     m <- mean(log(x))
     v <- var(log(x))
     shape <- 1.2/sqrt(v)
@@ -67,6 +73,8 @@ start.arg.default <- function(x, distr)
     start <- list(min=0, max=1)
   }else if (distr == "invgamma")
   {
+    if (any(x < 0)) 
+      stop("values must be positive to fit an inverse gamma  distribution")
     #http://en.wikipedia.org/wiki/Inverse-gamma_distribution
     m1 <- mean(x)
     m2 <- mean(x^2)
@@ -75,6 +83,8 @@ start.arg.default <- function(x, distr)
     start <- list(shape=shape, scale=scale)
   }else if (distr == "llogis")
   {
+    if (any(x < 0)) 
+      stop("values must be positive to fit a log-logistic  distribution")
     p25 <- as.numeric(quantile(x, 0.25))
     p75 <- as.numeric(quantile(x, 0.75))
     shape <- 2*log(3)/(log(p75)-log(p25))
@@ -82,6 +92,8 @@ start.arg.default <- function(x, distr)
     start <- list(shape=shape, scale=scale)
   }else if (distr == "invweibull")
   {
+    if (any(x < 0)) 
+      stop("values must be positive to fit an inverse Weibull distribution")
     g <- log(log(4))/(log(log(4/3)))
     p25 <- as.numeric(quantile(x, 0.25))
     p75 <- as.numeric(quantile(x, 0.75))
@@ -90,6 +102,8 @@ start.arg.default <- function(x, distr)
[…]

### 2016-06-19 tests/t-cdf-CI-band.R
update cdfband() to remove double plotting, correct typo in FAQ.

@@ -14,13 +14,17 @@ b1 <- bootdist(f1, niter=nbboot, silent=TRUE)
 plot(b1)
 quantile(b1)
 
-par(mfrow=c(1,3))
+par(mfrow=c(1,2))
 cdfband(b1, CI.level=90/100)
+cdfband(b1, CI.level=90/100, CI.col="grey80", CI.fill=TRUE, datacol="blue")
+
+
+par(mfrow=c(1,2))
 cdfband(b1, CI.level=90/100, CI.type = "less")
+cdfband(b1, CI.level=90/100, CI.col="grey85", CI.type = "less", CI.fill=TRUE, verticals=TRUE, datacol="blue", do.points=FALSE)
+
+par(mfrow=c(1,2))
 cdfband(b1, CI.level=90/100, CI.type = "greater")
+cdfband(b1, CI.level=90/100, CI.col="grey90", CI.type = "greater", CI.fill=TRUE, datacol="blue", datapch=21)
 
-
-par(mfrow=c(1,3))
-cdfband(b1, CI.level=90/100, CI.col="grey90", CI.fill=TRUE, datacol="blue", verticals=TRUE)
-cdfband(b1, CI.level=90/100, CI.col="grey90", CI.type = "less", CI.fill=TRUE, datacol="blue", do.points=FALSE)
-cdfband(b1, CI.level=90/100, CI.col="grey90", CI.type = "greater", CI.fill=TRUE, datacol="blue", verticals=TRUE)
+#some ideas from http://edild.github.io/ssd/

### 2016-06-27 R/gradlogLik.R
first cleaning of the vignette directory.

@@ -19,7 +19,7 @@
 #############################################################################
 ### gradient of log-likelihood
 ###
-###         R functions
+###         R functions (not exported)
 ### 
 
 #Beta distribution

### 2016-06-30 inst/doc/FAQ.R
Two pooints added to the FAQ

@@ -295,3 +295,22 @@ bg$CI
 fg$estimate + cbind("estimate"= 0, "2.5%"= -1.96*fg$sd, "97.5%"= 1.96*fg$sd)
 llplot(fg)
 
+## ---- fig.height=4, fig.width=6, warning = FALSE-------------------------
+data(salinity)
+log10LC50 <-log10(salinity)
+fit <- fitdistcens(log10LC50, "norm")
+# Bootstrap 
+bootsample <- bootdistcens(fit, niter = 101)
+#### We used only 101 iterations in that example to limit the calculation time but
+#### in practice you should take at least 1001 bootstrap iterations
+# Calculation of the quantile of interest (here the 5 percent hazard concentration)
+(HC5 <- quantile(bootsample, probs = 0.05))
+# visualizing pointwise confidence intervals on other quantiles
+CIcdfplot(bootsample, CI.output = "quantile", CI.fill = "pink", xlim = c(0.5,2))
+
+## ------------------------------------------------------------------------
+exposure <- 1.2
+# Bootstrap sample of the PAF at this exposure
+PAF <- pnorm(exposure, mean = bootsample$estim$mean, sd = bootsample$estim$sd)
+# confidence interval from 2.5 and 97.5 percentiles
+quantile(PAF, probs = c(0.025, 0.975))

### 2016-07-01 inst/doc/Optimalgo.R
version submitted to win-builder.r-project with light html vignettes

@@ -1,7 +1,8 @@
-## ---- message=FALSE------------------------------------------------------
+## ----setup, echo=FALSE, message=FALSE, warning=FALSE---------------------
 require(fitdistrplus)
-require(actuar)
-require(knitr)
+require(knitr) #for kable() function
+set.seed(12345)
+options(digits = 3)
 
 ## ---- echo=TRUE, eval=FALSE----------------------------------------------
 #  fitbench <- function(data, distr, method, grad=NULL, control=list(trace=0, REPORT=1, maxit=1000), lower=-Inf, upper=+Inf, ...)
@@ -14,10 +15,9 @@ lnL <- function(par, fix.arg, obs, ddistnam)
   fitdistrplus:::loglikelihood(par, fix.arg, obs, ddistnam) 
 grlnlbeta <- fitdistrplus:::grlnlbeta
 
-## ---- fig.height=5, fig.width=5------------------------------------------
+## ---- fig.height=4, fig.width=4------------------------------------------
 #(1) beta distribution
 n <- 200
-set.seed(12345)
 x <- rbeta(n, 3, 3/4)
 grlnlbeta(c(3, 4), x) #test
 hist(x, prob=TRUE)
@@ -57,17 +57,17 @@ kable(expopt[, grep("G-", colnames(expopt), invert=TRUE)], digits=3)
 ## ---- results='asis', echo=FALSE-----------------------------------------
 kable(expopt[, grep("G-", colnames(expopt))], digits=3)
 
-## ---- fig.width=5, fig.height=5------------------------------------------
+## ---- fig.width=4, fig.height=4------------------------------------------
 llsurface(min.arg=c(0.1, 0.1), max.arg=c(7, 3), 
           plot.arg=c("shape1", "shape2"), nlev=25,
-          plot.np=50, data=x, distr="beta")
+          plot.np=50, data=x, distr="beta", back.col = FALSE)
 points(unconstropt[1,"BFGS"], unconstropt[2,"BFGS"], pch="+", col="red")
 points(3, 3/4, pch="x", col="green")
 
-## ---- fig.width=5, fig.height=5------------------------------------------
+## ---- fig.width=4, fig.height=4------------------------------------------
 b1 <- bootdist(fitdist(x, "beta", method="mle", optim.method="BFGS"), niter=100, parallel="snow", ncpus=2)
 summary(b1)
-plot(b1, enhance=TRUE, trueval=c(3, 3/4))
+plot(b1, trueval=c(3, 3/4))
 
[…]

### 2018-01-06 R/mgedist.R
first version testing d, p, q functions.

@@ -105,6 +105,7 @@ mgedist <- function (data, distr, gof = "CvM", start=NULL, fix.arg=NULL, optim.m
       vstart <- unlist(chfixstt$start.arg(data))
     else
       vstart <- unlist(chfixstt$start.arg)
+    #set fix.arg.fun
     if(is.function(fix.arg)) #function
     { 
       fix.arg.fun <- fix.arg

### 2018-01-06 R/mledist.R

@@ -104,6 +104,7 @@ mledist <- function (data, distr, start=NULL, fix.arg=NULL, optim.method="defaul
       vstart <- unlist(chfixstt$start.arg(data))
     else
       vstart <- unlist(chfixstt$start.arg)
+    #set fix.arg.fun
     if(is.function(fix.arg)) #function
     { 
       fix.arg.fun <- fix.arg

### 2018-01-06 R/qmedist.R

@@ -87,6 +87,7 @@ qmedist <- function (data, distr, probs, start=NULL, fix.arg=NULL,
       vstart <- unlist(chfixstt$start.arg(data))
     else
       vstart <- unlist(chfixstt$start.arg)
+    #set fix.arg.fun
     if(is.function(fix.arg)) #function
     { 
       fix.arg.fun <- fix.arg

### 2018-01-06 tests/t-util-testdensity.R

@@ -4,10 +4,10 @@ library(fitdistrplus)
 testdpqfun <- fitdistrplus:::testdpqfun
 
 ##### existence ##### 
-#a list of TRUE
-testdpqfun("exp")
-#a list of error messages
-testdpqfun("exp2")
+#a data.frame of TRUE and ""
+testdpqfun("exp", start=c(rate=1))
+#a data.frame with error messages
+testdpqfun("exp2", start=c(rate=1))
 
 ##### void vector ##### 
 dexp2 <- function(x, rate)
@@ -80,9 +80,25 @@ testdpqfun("norm2", "d", c(mean=1, sd=1))
 
 ##### inconsistent name ##### 
 dnorm2 <- function(x, mean=0, sd=1, ...)
-    dnorm(x,mean,sd)
+    dnorm(x, mean, sd)
+
+dnorm3 <- dnorm2
+pnorm3 <- pnorm
+qnorm3 <- qnorm
 
 #TRUE
 testdpqfun("norm", "d", c(mean=1, sd=1))
 #error message
 testdpqfun("norm2", "d", c(mean=1, sd=1))
+
+
+#a data.frame with error messages
+testdpqfun("norm", c("d", "p", "q"), c(mean=1, sd=1))
+testdpqfun("norm2", c("d", "p", "q"), c(mean=1, sd=1))
+testdpqfun("norm3", c("d", "p", "q"), c(mean=1, sd=1))
+
+x <- rnorm(100)
+fitdist(x, "norm") #ok
+fitdist(x, "norm2", start=list(mean=1, sd=1)) #pnorm2 not defined
+fitdist(x, "norm3", start=list(mean=1, sd=1)) #The dnorm3 function should return raise an error when names are incorrectly named
+

### 2018-02-05 R/denscomp.R
Modifications following previous commits by Christophe concerning chechk of dpq functions, new starting values and discrete distributions in denscomp.

@@ -146,6 +146,10 @@ denscomp <- function(ft, xlim, ylim, probability = TRUE, main, xlab, ylab,
     if(length(legendtext) != length(unique(legendtext)))
       legendtext <- paste(legendtext, 1:nft, sep="-")
   }
+  
+  # forces demp to TRUE if discrete is TRUE
+  demp <- TRUE
+  
   #add empirical density/fmp to legend vectors
   if(demp)
   {
@@ -163,29 +167,32 @@ denscomp <- function(ft, xlim, ylim, probability = TRUE, main, xlab, ylab,
     
     if(!discrete)
     {
-      #plot fitted densities (line)
-      if(fittype %in% c("l", "o"))
-        for(i in 1:nft)
-          lines(sfin, fitteddens[,i], lty=fitlty[i], col=fitcol[i], ...)
-      #plot fitted densities (point)
-      if(fittype %in% c("p", "o"))
-        for(i in 1:nft)
-          points(sfin, fitteddens[,i], col=fitcol[i])
+      #main plotting
+      reshist <- hist(mydata, main = main, xlab = xlab, ylab = ylab, xlim = xlim, ylim = ylim, col = datacol, probability = probability, ...)
       
+      #plot fitted densities (line)
+      for(i in 1:nft)
+          lines(sfin, fitteddens[,i], lty=fitlty[i], col=fitcol[i], ...)
+ 
       #plot empirical density
       if(demp)
         lines(density(mydata)$x, density(mydata)$y * scalefactor, col=dempcol)
       
       if (addlegend)
         legend(x=xlegend, y=ylegend, bty="n", legend=legendtext, lty=fitlty, col=fitcol, ...)
-    }else
+    }else # so if discrete
     {
+      #main plotting
+      # plotting of an empty histogramm
+      reshist <- hist(mydata, main = main, xlab = xlab, ylab = ylab, xlim = xlim, ylim = ylim, border = "white",
+                      probability = probability, ...)
+       
       eps <- diff(range(sfin))/200
       if(fittype %in% c("l", "o"))
       {
         #plot fitted mass probability functions (line)
         for(i in 1:nft)
-          lines(sfin+(i-1)*eps, fitteddens[,i], lty=fitlty[i], col=fitcol[i], type="h", ...)
+          lines(sfin+(i)*eps, fitteddens[,i], lty=fitlty[i], col=fitcol[i], type="h", ...)
         #plot empirical mass probabilty function
         if(demp)
         {
[…]

### 2018-02-05 R/fitdist.R

@@ -61,6 +61,7 @@ fitdist <- function (data, distr, method = c("mle", "mme", "qme", "mge"), start=
     #check argument data
     if (!(is.vector(data) & is.numeric(data) & length(data)>1))
         stop("data must be a numeric vector of length greater than 1")
+ 
     #encapsulate three dots arguments
     my3dots <- list(...)    
     if (length(my3dots) == 0) 
@@ -68,6 +69,13 @@ fitdist <- function (data, distr, method = c("mle", "mme", "qme", "mge"), start=
     n <- length(data)
     
     # definition of starting/fixed values values
+    ### ML comment : it is perturbating to see the following lines here
+    # as the starting values are really computed in mledist
+    # is it really necessary to do that
+    # il yes the real computing of starting values should be here and not in mledist
+    # but it will require hazardous changes, especially concerning fitdistcens for which the
+    # computing of starting values needs more lines of code
+    # I am not in favour of such a BIG CHANGE in the struture of the code
     argddistname <- names(formals(ddistname))
     chfixstt <- checkparam(start.arg=start, fix.arg=fix.arg, argdistname=argddistname, 
                            errtxt=NULL, data10=head(data, 10), distname=distname)

### 2018-02-05 R/util-testdensity.R

@@ -77,7 +77,7 @@ test1fun <- function(fn, start.arg, fix.arg)
   x <- c(0, 1, NA)
   res2 <- try(do.call(fn, c(list(x), as.list(start.arg), as.list(fix.arg))), silent=TRUE)
   t4 <- paste("The", fn, "function should return a vector of with NA values when input has missing values and not raise an error")
-  t5 <- paste("The", fn, "function should return a vector of with NA values when input has missing values and not remove missing value")
+  t5 <- paste("The", fn, "function should return a vector of with NA values when input has missing values and not remove missing values")
   if(class(res2) == "try-error")
   {
     res$txt <- t4
@@ -104,7 +104,7 @@ test1fun <- function(fn, start.arg, fix.arg)
   x <- 0:1
   names(start.arg) <- paste0(names(start.arg), "_")
   res4 <- try(do.call(fn, c(list(x), as.list(start.arg), as.list(fix.arg))), silent=TRUE)
-  t8 <- paste("The", fn, "function should return raise an error when names are incorrectly named")
+  t8 <- paste("The", fn, "function should raise an error when names are incorrectly named")
   if(class(res4) != "try-error")
   {
     res$txt <- t8

### 2018-08-10 inst/doc/FAQ.R
FAQ and paper2JSS updated

@@ -411,7 +411,7 @@ plotdistcens(dsmo, NPMLE = FALSE)
 ## ---- fig.height= 6, fig.width= 6-----------------------
 par(mfrow = c(2, 1),  mar = c(3, 4, 3, 0.5))
 # Turnbull plot of the ECDF
-plotdistcens(dsmo, NPMLE.method = "Turnbull")
+plotdistcens(dsmo, NPMLE.method = "Turnbull", xlim = c(-1.8, 2.4))
 # Wang plot of the ECD
 plotdistcens(dsmo, NPMLE.method = "Wang")
 
@@ -467,7 +467,7 @@ qqcompcens(list(fnorm, flogis))
 ppcompcens(list(fnorm, flogis))
 
 ## ------ fig.height= 4, fig.width= 8 ---------------------
-par(mfrow = c(1,2))
-qqcompcens(fnorm, fitcol = "red", fillrect = "red", lwd = 2)
-qqcompcens(flogis, fitcol = "green", fillrect = "green", lwd = 2)
+qqcompcens(list(fnorm, flogis), lwd = 2, plotstyle = "ggplot",
+           fitcol = c("red", "green"), fillrect = c("pink", "lightgreen"),
+           legendtext = c("normal distribution", "logistic distribution"))
 

### 2020-05-15 inst/doc/FAQ.R
Upgraded vignettes

@@ -106,6 +106,18 @@ curve(ptiexp(x, 1, .5, 3), add=TRUE, col="blue", lty=3)
 legend("bottomright", lty=1:3, col=c("red", "green", "blue", "black"), 
         leg=c("full MLE", "MLE fixed arg", "true CDF", "emp. CDF"))
 
+## ------------------------------------------------------------------------
+x <- rbeta(1000, 3, 3)
+dbeta2 <- function(x, shape, ...)
+	dbeta(x, shape, shape, ...)
+pbeta2 <- function(q, shape, ...)
+	pbeta(q, shape, shape, ...)	
+fitdist(x, "beta2", start=list(shape=1/2))
+
+## ------------------------------------------------------------------------
+x <- rbeta(1000, .3, .3)
+fitdist(x, "beta2", start=list(shape=1/2), optim.method="L-BFGS-B", lower=1e-2)	
+
 ## ---- fig.height=3, fig.width=6------------------------------------------
 set.seed(1234)
 x <- rgamma(n = 100, shape = 2, scale = 1)
@@ -283,6 +295,18 @@ rsexp <- function(n, rate, shift)
 x <- rsexp(1000, 1/4, 1)
 fitdist(x, "sexp", start=list(rate=1, shift=0), lower= c(0, -min(x)))
 
+## ---- message=FALSE------------------------------------------------------
+library(GeneralizedHyperbolic)
+myoptim <- function(fn, par, ui, ci, ...)
+{
+  res <- constrOptim(f=fn, theta=par, method="Nelder-Mead", ui=ui, ci=ci, ...)
+  c(res, convergence=res$convergence, value=res$objective, par=res$minimum, hessian=res$hessian)
+}
+x <- rnig(1000, 3, 1/2, 1/2, 1/4)
+ui <- rbind(c(0,1,0,0), c(0,0,1,0), c(0,0,1,-1), c(0,0,1,1))
+ci <- c(0,0,0,0)
+fitdist(x, "nig", custom.optim=myoptim, ui=ui, ci=ci, start=list(mu = 0, delta = 1, alpha = 1, beta = 0))
+
 ## ---- fig.height=3, fig.width=6------------------------------------------
 pgeom(0:3, prob=1/2)
 qgeom(c(0.3, 0.6, 0.9), prob=1/2)
@@ -378,44 +402,46 @@ dcomp <- denscomp(list(fitW, fitln, fitg), legendtext = c("Weibull", "lognormal"
[…]

### 2020-05-15 inst/doc/paper2JSS.R

@@ -1,5 +1,5 @@
 ### R code from vignette source 'paper2JSS.Rnw'
-### Encoding: UTF-8
+### Encoding: ISO8859-1
 
 ###################################################
 ### code chunk number 1: par4vignette
@@ -204,7 +204,7 @@ fdanish.P.MLE <- fitdist(danishuni$Loss, "pareto",
 memp <- function(x, order) sum(x^order)/length(x)
 fdanish.P.MME <- fitdist(danishuni$Loss, "pareto", 
     method="mme", order=1:2, 
-memp="memp", start=c(shape=10, scale=10), lower=c(2+1e-6,2+1e-6), 
+memp="memp", start=list(shape=10, scale=10), lower=c(2+1e-6,2+1e-6), 
 upper=c(Inf,Inf))
 par(mfrow=c(1, 2))
 cdfcomp(list(fdanish.ln.MLE, fdanish.ln.MME), 
@@ -327,13 +327,13 @@ summary(fsal.ll)
 ###################################################
 ### code chunk number 34: fitsalinity.cdfcomp.echo (eval = FALSE)
 ###################################################
-par(mfrow=c(2, 2))
-cdfcompcens(list(fsal.ln, fsal.ll),
-            legendtext=c("lognormal", "loglogistic "))
-qqcompcens(fsal.ln, legendtext = "lognormal")
-ppcompcens(fsal.ln, legendtext = "lognormal")
-qqcompcens(list(fsal.ln, fsal.ll), legendtext = c("lognormal", "loglogistic "),
-           main = "Q-Q plot with 2 dist.")
+## par(mfrow=c(2, 2))
+## cdfcompcens(list(fsal.ln, fsal.ll), 
+##   legendtext = c("lognormal", "loglogistic "))
+## qqcompcens(fsal.ln, legendtext = "lognormal")
+## ppcompcens(fsal.ln, legendtext = "lognormal")
+## qqcompcens(list(fsal.ln, fsal.ll), legendtext = c("lognormal", "loglogistic "),
+##            main = "Q-Q plot with 2 dist.")
 
 
 ###################################################
@@ -342,7 +342,7 @@ qqcompcens(list(fsal.ln, fsal.ll), legendtext = c("lognormal", "loglogistic "),
 getOption("SweaveHooks")[["fig"]]()
 par(mfrow=c(2, 2))
 cdfcompcens(list(fsal.ln, fsal.ll),
-            legendtext=c("lognormal", "loglogistic "))
+    legendtext=c("lognormal", "loglogistic "))
[…]

### 2020-05-15 share/prez/2013rencontresR/src/fitdistrplus280613.Rnw
add a share/ directory with technical doc and presentation files

\documentclass[9pt, hide notes]{beamer}

\mode<presentation>
{

    %---- themes
  	\usetheme{Antibes}
    %\usetheme{Boadilla}
  	%\usetheme{Singapore}
  	%\usetheme{Warsaw}

	%nice theme for math fonts
	\usefonttheme{professionalfonts}

	%color theme
	\usecolortheme{lily}

	  %for math characters
  %	\usefonttheme[onlylarge]{structuresmallcapsserif}
	%\usefonttheme[onlysmall]{structurebold}

	%remove navigation symbols
	\setbeamertemplate{navigation symbols}{}

	%dark blue for normal text
%	\definecolor{MyDarkBlue}{rgb}{0.0235,0.1569,0.4745}
%	\setbeamercolor{normal text}{fg = MyDarkBlue}

	%white for background text color
	\setbeamercolor{structure}{bg = white}

	%---- headline
	% black and white for headline and footline
	\setbeamercolor{section in head/foot}{bg = white, fg=black}
	\setbeamercolor{subsection in head/foot}{bg = white, fg=black}
	\setbeamercolor{title in head/foot}{bg = white, fg=black}

	%What do you want in your header? by default, there is the current subsection

	%if you want to insert a complete navigation	
		\setbeamertemplate{headline}
		{%--
			\begin{beamercolorbox}{section in head/foot}
				\vskip2pt\insertnavigation{\paperwidth}\vskip2pt
			\end{beamercolorbox}
		}%--

	%---- footline	
	%add frame number and logo
	\setbeamertemplate{footline}
	{%--
	   %you can adjust the left and right white space with leftskip and rightskip
	\begin{beamercolorbox}[ht=2.5ex,dp=1.125ex,
	    leftskip=.15cm,rightskip=.15cm plus1fil]{title in head/foot}
	     \insertframenumber/\inserttotalframenumber
	     \hspace{0.5cm}	%you can use different horizontal spacing!
	     \insertshortdate
	     \hspace{0.5cm}
	     Rencontres R 2013
	     \hfill
       M.L. Delignette-Muller et C. Dutang
	\end{beamercolorbox}
  	}%--

	%% No footline on title slide
	\defbeamertemplate{footline}{title slide}{%
	\relax
	}

	%---- misc
  %\setbeamercovered{transparent}
  % or whatever (possibly just delete it)

}

\usepackage[francais]{babel}
\usepackage[latin1]{inputenc}
%\usepackage{beamerthemesplit}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{pdfcolmk}
\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage{url}

% packages de Christophe en plus ?
\usepackage{wrapfig}
\usepackage{amsmath,amsfonts,amssymb,latexsym,qsymbols}
\usepackage{multirow}

\title[fitdistrplus : nouvelles fonctionnalit\'es]% 
{Nouvelles fonctionnalit\'es du package \texttt{fitdistrplus}}

\author[M.L. Delignette-Muller et C. Dutang]
{\textbf{Marie Laure Delignette-Muller}\\ Universit\'e de Lyon \\
UMR CNRS 5558, Laboratoire de Biom\'etrie et Biologie Evolutive\\
VetAgro Sup, Campus v\'et\'erinaire de Lyon\\et\\
           \textbf{Christophe Dutang} \\ Universit\'e de Strasbourg \\
           UMR CNRS 7501, Institut de Recherche Math\'ematique Avanc\'ee}

\date[26-27 juin 2013]{Deuxi\`emes rencontres R, Lyon, 27-28 juin 2013}

\begin{document}
\SweaveOpts{concordance=TRUE}
\setkeys{Gin}{width=0.5\textwidth}
% une valeur inf?rieure au 0.8 par d?faut r?duit la taille des figures

\begin{frame}[fragile]
\titlepage
\begin{center}
\includegraphics[width=1cm]{Rlogo_2.jpg}
\end{center}
\end{frame}

  %DIA 2
\begin{frame}[fragile]
\frametitle{Pr\'esentation du package \texttt{fitdistrplus}}
\begin{center}
\url{http://cran.r-project.org/package=fitdistrplus}\\
\url{http://riskassessment.r-forge.r-project.org/}
\end{center}
Package pour \textbf{l'ajustement de distributions
param\'etriques \`a des donn\'ees univari\'ees},\\ proposant diverses fonctions pour :
  \begin{itemize}
\item le choix de distributions candidates pour d\'ecrire les donn\'ees,
\item l'ajustement de chacune des distributions candidates aux donn\'ees,
\item la comparaison des ajustements en vue de choisir la distribution la plus adapt\'ee,
\item le calcul, par bootstrap, de l'incertitude sur les param\`etres estim\'es de la distribution choisie.
\end{itemize}
\end{frame}

% DIA 3
\begin{frame}
\frametitle{Quelques sp\'ecificit\'es du package}
\begin{itemize}
\item \textbf{Diff\'erentes m\'ethodes d'estimation des param\`etres} sont propos\'ees :
  \begin{itemize}
  \item m\'ethode du \textcolor{blue}{maximum de vraisemblance (MLE)}, 
  \item m\'ethode des \textcolor{blue}{moments (MME)}, 
  \item m\'ethode des \textcolor{blue}{quantiles (QME)}, 
  \item m\'ethode de \textcolor{blue}{minimisation d'une statistique d'ajustement (MGE)} 
  \\(8 distances propos\'ees).
  \end{itemize}
\item Prise en compte 
    \begin{itemize}
  \item des donn\'ees discr\`etes 
    \item et des \textbf{donn\'ees censur\'ees} quel que
soit le type de censures \\(\`a droite, \`a gauche ou par intervalle).
  \end{itemize}
\item Possibilit\'e de modifier l'algorithme d'optimisation \\(par d\'efaut \texttt{optim}).
\end{itemize}
\end{frame}

% DIA 4
\begin{frame}
\frametitle{Utilisation actuelle du package}
\textbf{Package d\'evelopp\'e initialement pour l'appr\'eciation quantitative du risque}\\ 
%(projet Rforge
%\url{http://riskassessment.r-forge.r-project.org/})\\
notamment pour permettre
la caract\'erisation hi\'erarchique de la variabilit\'e et de l'incertitude dans le cadre de simulations de 
Monte carlo \`a deux dimensions (package \texttt{mc2d}).\\
\textbf{Package g\'en\'eraliste simple \`a utiliser}, \\utilis\'e dans des domaines vari\'es :
\begin{itemize}
\item risque alimentaire, 
\item \'epid\'emiologie, 
\item biologie mol\'eculaire, 
%\item g\'enomique, 
\item bioinformatique, 
\item math\'ematiques financi\`eres et actuarielles, \dots
\end{itemize}
\textcolor{blue}{$\rightarrow$ les retours des utilisateurs nous ont incit\'e \`a d\'evelopper de 
\textbf{nouvelles fonctionnalit\'es}}.
\end{frame}

% DIA 5
\begin{frame}[fragile]
\frametitle{Comparaison visuelle de plusieurs ajustements (\texttt{cdfcomp})}
\begin{footnotesize}
<<echo=FALSE,results=hide>>=
rm(list=ls(all=TRUE))
set.seed(1234) 
library(fitdistrplus)
options(digits=3)
@
\setkeys{Gin}{width=0.9\textwidth}
%<<cdfgroundbeef, echo=TRUE, fig=TRUE, width=7, height=4, eps=FALSE>>=
%data(groundbeef)
%fw <- fitdist(groundbeef$serving, "weibull")
%fg <- fitdist(groundbeef$serving,"gamma")
%fln <- fitdist(groundbeef$serving,"lnorm")
%cdfcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
%@
<<cdfgroundbeefecho, echo=TRUE, fig=FALSE>>=
data(groundbeef)
fw <- fitdist(groundbeef$serving, "weibull")
fg <- fitdist(groundbeef$serving,"gamma")
fln <- fitdist(groundbeef$serving,"lnorm")
cdfcomp(list(fw,fln,fg), 
        legendtext=c("Weibull", "lognormal", "gamma"))
@
<<cdfgroundbeeffig, echo=FALSE, fig=TRUE, width=7, height=4, eps=FALSE>>=
cdfcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"),
        fitlty=1,lwd=2)
@
\end{footnotesize}
\end{frame}

% DIA 6
\begin{frame}[fragile]
\frametitle{Comparaison visuelle de plusieurs ajustements : \\repr\'esentation en densit\'e de probabilit\'e (\texttt{denscomp})}
\begin{footnotesize}
\setkeys{Gin}{width=0.9\textwidth}
<<densgroundbeefecho, echo=TRUE, fig=FALSE>>=
denscomp(list(fw,fln,fg), 
         legendtext=c("Weibull", "lognormal", "gamma"))
@
<<densgroundbeeffig, echo=FALSE, fig=TRUE, width=7, height=4.5, eps=FALSE>>=
denscomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"),
        fitlty=1,lwd=2)
@
\end{footnotesize}
\end{frame}

% DIA 7
%\begin{frame}[fragile]
%\frametitle{Comparaison visuelle de plusieurs ajustements : P-P plot (\texttt{ppcomp})
%ou en Q-Q plot (\texttt{qqcomp})}
%\begin{footnotesize}
%\setkeys{Gin}{width=0.9\textwidth}
%<<ppqqgroundbeefecho, echo=TRUE, fig=FALSE>>=
%ppcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
%qqcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"))
%@
%<<ppqqgroundbeeffig, echo=FALSE, fig=TRUE, width=7, height=4, eps=FALSE>>=
%par(mfrow=c(1,2))
%par(mar=c(4,4,1,1))
%ppcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"),
%       fitpch=16)
%qqcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"),
%       fitpch=16)
%@
%\end{footnotesize}
%\end{frame}

% DIA 7
\begin{frame}[fragile]
\frametitle{Comparaison visuelle de plusieurs ajustements :\\ P-P plot (\texttt{ppcomp})}
\begin{footnotesize}
\setkeys{Gin}{width=0.9\textwidth}
<<ppgroundbeefecho, echo=TRUE, fig=FALSE>>=
ppcomp(list(fw,fln,fg), 
       legendtext=c("Weibull", "lognormal", "gamma"))
@
<<ppgroundbeeffig, echo=FALSE, fig=TRUE, width=7, height=4.5, eps=FALSE>>=
ppcomp(list(fw,fln,fg), 
       legendtext=c("Weibull", "lognormal", "gamma"),
       fitpch=16)
@
\end{footnotesize}
\end{frame}

% DIA 8
\begin{frame}[fragile]
\frametitle{Comparaison visuelle de plusieurs ajustements : \\
Q-Q plot (\texttt{qqcomp})}
\begin{footnotesize}
\setkeys{Gin}{width=0.9\textwidth}
<<qqgroundbeefecho, echo=TRUE, fig=FALSE>>=
qqcomp(list(fw,fln,fg), 
       legendtext=c("Weibull", "lognormal", "gamma"))
@
<<qqgroundbeeffig, echo=FALSE, fig=TRUE, width=7, height=4.5, eps=FALSE>>=
qqcomp(list(fw,fln,fg), legendtext=c("Weibull", "lognormal", "gamma"),
       fitpch=16)
@
\end{footnotesize}
\end{frame}

% DIA 9
\begin{frame}[fragile]
\frametitle{Comparaison num\'erique de plusieurs ajustements}
\begin{itemize}
\item \textbf{Statistiques d'ajustement}, \\
Kolmogorov-Smirnov, Cram\'er-von Mises et Anderson-Darling,\\
\item et \textbf{crit\`eres d'information} , AIC et BIC.
\end{itemize}
\begin{footnotesize}
\setkeys{Gin}{width=0.9\textwidth}
<<ppqqgroundbeefecho, echo=TRUE, fig=FALSE>>=
gofstat(list(fw,fln,fg),
        fitnames = c("Weibull", "lognormal", "gamma"))
@
\end{footnotesize}
\end{frame}

% DIA 10 
\begin{frame}[fragile]
\frametitle{Calcul de quantiles \`a partir d'une loi ajust\'ee : \\ exemple sur donn\'ees censur\'ees en \'ecotoxicologie}
\begin{footnotesize}
<<salinityecho, echo=TRUE, fig=FALSE>>=
data(salinity)
f <- fitdistcens(salinity, "lnorm")
quantile(f,probs = c(0.05, 0.20))
@
\begin{center}
\setkeys{Gin}{width=0.7\textwidth}
<<salinityfig, echo=FALSE, fig=TRUE, width=6, height=3.5, eps=FALSE>>=
par(mar=c(4,4,1,1))
cdfcompcens(f,lwd=2,addlegend=FALSE)
probs <- c(0.05,0.2)
q <- quantile(f,probs = probs)$quantiles
abline(h=probs[1],col="red",lty=2,lwd=2)
abline(h=probs[2],col="red",lty=3,lwd=2)
abline(v=q[1],col="red",lty=2,lwd=2)
abline(v=q[2],col="red",lty=3,lwd=2)
points(q,probs,col="red",pch=16)
text(2,0.05,"5%",pos=3,col="red")
text(2,0.2,"20%",pos=3,col="red")
@
\end{center}
\end{footnotesize}
\end{frame}

% DIA 11 
\begin{frame}[fragile]
\frametitle{Bootstrap}
\begin{footnotesize}
<<salinitybootecho, echo=TRUE, fig=FALSE>>=
b <- bootdistcens(f,niter=1001) 
summary(b)
plot(b)
@
\begin{center}
\setkeys{Gin}{width=0.45\textwidth}
<<salinitybootfig, echo=FALSE, fig=TRUE, width=6, height=5, eps=FALSE>>=
par(mar=c(4,4,1,1))
plot(b,pch=16)
@
\end{center}\end{footnotesize}
\end{frame}

% DIA 12
\begin{frame}[fragile]
\frametitle{Intervalles de confiance bootstrap sur les quantiles}
\begin{footnotesize}
<<salinityICecho, echo=TRUE, fig=FALSE>>=
cdfcompcens(f)
q <- quantile(b, probs = seq(0,1,length=101))
points(q$quantCI[1,],q$probs,type="l",lty=2,col="red")
points(q$quantCI[2,],q$probs,type="l",lty=2,col="red")
@
\begin{center}
\setkeys{Gin}{width=0.9\textwidth}
<<salinityICfig, echo=FALSE, fig=TRUE, width=7, height=4, eps=FALSE>>=
par(mar=c(4,4,1,1))
cdfcompcens(f,lwd=2,,addlegend=FALSE)
probs = seq(0,1,length=101)
q <- quantile(b, probs = probs)
points(q$quantCI[1,],q$probs,type="l",lty=2,col="red",lwd=2)
points(q$quantCI[2,],q$probs,type="l",lty=2,col="red",lwd=2)
arrows(q$quantCI[1,70],probs[70],q$quantCI[2,70],probs[70],
       col="red",lwd=2,code=3,length=0.1)
text((q$quantCI[1,70]+q$quantCI[2,70])/2,probs[70],"IC 95%",col="red",pos=3)
@
\end{center}\end{footnotesize}
\end{frame}

% DIA 13 
\begin{frame}[fragile]
\frametitle{Perspectives}
Nous envisageons 
\begin{itemize}
\item d'\textbf{\'etendre aux donn\'ees censur\'ees certaines m\'ethodes} disponibles
actuellement sur les donn\'ees non censur\'ees,
\item d'\textbf{\'elargir le choix de distances} utilisables pour l'ajustement par minimisation de ces distances et pour l'\'evaluation de la qualit\'e de l'ajustement \\(distances bas\'ees sur les quantiles),
\item d'\textbf{\'elargir le choix des moments} pour l'ajustement par la m\'ethode des moments \\(``L-moment'' et ``TL-moment''),
\item aborder l'\textbf{ajustement de distributions multivari\'ees}.
\end{itemize}
\textcolor{blue}{N'h\'esitez pas \`a nous envoyer vos retours ou id\'ees en vue de l'am\'elioration de ce package.}
\end{frame}

\end{document}

### 2020-05-15 share/prez/2018recontresR/Delignette_slides4R2108-v7.Rmd
add a share/ directory with technical doc and presentation files

---
title: "New goodness-of-fit plots for censored data in the package **fitdistrplus**"
author: |
 | M.L. Delignette-Muller (LBBE, Lyon)
 | C. Dutang (CEREMADE, Paris) and 
 | A. Siberchicot (LBBE, Lyon)
date: "July 6th 2018"
output:
  beamer_presentation:
 # theme: "Antibes"
 #   toc: true
    fig_width: 8
    fig_height: 6
    fig_caption: false
---
```{r, echo = FALSE, results = "hide", message = FALSE, warning = FALSE}
options(digits = 3)
require(fitdistrplus)
set.seed(123)
addbounds <- function(d)
{
  xbounds <- c(d$left, d$right)
  xboundsnotNA <- xbounds[!is.na(xbounds)]
  abline(v = xboundsnotNA, col = "grey")
}
addLR <- function(d)
{
  Lbounds <- d$left[!is.na(d$left)]
  Rbounds <- d$right[!is.na(d$right)]
  range <- range(c(Lbounds,Rbounds)) 
  eps <- (range[2] - range[1]) * 0.01
  text(x = Lbounds-eps, y = 0.05, labels = "L", col = "red")
  text(x = Rbounds+eps, y = 0.05, labels = "R", col = "red")
}

addeq <- function(d)
{
  left <- d$left
  left[is.na(left)] <- -100
  right <- d$right
  right[is.na(right)] <- 100
  rect(left, -2, right, 2, density = 10)
}

addTurnbull <- function(d, conf.int = FALSE)
{
  (survdata <- Surv(time = d$left, time2 = d$right, type="interval2"))
  survfitted <- survfit(Surv(left, right, type="interval2") ~ 1, data = d)
  points(survfitted$time, 1 - survfitted$surv, col = "orange", type = "s", 
         lwd = 5, lty = 2)
  if (conf.int == TRUE)
  {
      points(survfitted$time, 1 - survfitted$lower, col = "orange", type = "s", 
             lwd = 2, lty = 3)
      points(survfitted$time, 1 - survfitted$upper, col = "orange", type = "s", 
             lwd = 2, lty = 3)
  }
}
```

## General presentation of the package fitdistrplus
A package to help the **fit of parametric distributions** to 
univariate discrete our continuous non censored or censored data.

* stable version 1.0-9 on **CRAN** (first release in 2009).
* version 1.0-10 in development on **Rforge** (soon on CRAN).
* Delignette-Muller, M. L., & Dutang, C. (2015). fitdistrplus: An R package for fitting distributions. **Journal of Statistical Software**, 64(4), 1-34. (311 citations in scholar google)
* A FAQ vignette continuously updated in each new version.

```{r, echo = FALSE, fig.height=3}
par(mfrow = c(1,3))
set.seed(124)
r <- rgamma(100, shape = 8, rate = 1)
f <- fitdist(r, "gamma")
b <- bootdist(f)
f.mme <- fitdist(r, "gamma", method = "mme")
f.mge <- fitdist(r, "gamma", method = "mge", gof="AD2R")
f.qme <- fitdist(r, "gamma", method = "qme", probs = c(0.25, 0.75))
denscomp(list(f, f.mme, f.qme, f.mge), legendtext = c("MLE", "MME", "QME", "MGE"),
         main = "Different fitting methods", xlim = c(2, 16), fitlty = 1)
CIcdfplot(b, CI.output = "quantile", CI.fill = "pink", main = "Uncertainty using bootstrap")

rd <- rpois(50, lambda = 5)
fp <- fitdist(rd, "pois")
denscomp(fp, demp = TRUE, main = "Fit of discrete distributions")
```

# Goodness-of-fit plots for non censored data

## An example with non censored data
```{r}
r <- rweibull(100, shape = 3, scale = 1)
fw <- fitdist(r, "weibull")
fg <- fitdist(r, "gamma")
fl <- fitdist(r, "lnorm")
gofstat(list(fw, fg, fl), 
        fitnames = c("Weibull", "gamma", "lnorm"))
```

## A goodness-of-fit plot in density plot
```{r}
denscomp(list(fw, fg, fl), demp = TRUE, fitlty = 1)
```

## A goodness-of-fit plot in CDF
```{r}
cdfcomp(list(fw, fg, fl), fitlty = 1)
```

## A Q-Q plot which emphasizes differences at tails
```{r}
qqcomp(list(fw, fg, fl))
```

## a P-P plot which emphasizes differences in the center
```{r}
ppcomp(list(fw, fg, fl))
```

# Representation of the ECDF for censored data

## How to represent an ECDF from censored data ?

**A first toy example with left, right and interval censored data**
```{r, echo = FALSE}
d0 <- data.frame(left = c(NA, 2, 4, 7, 10), right = c(1, 2.5, 6, 8, NA))
d <- d0
```
```{r}
d
```
```{r, echo = FALSE, fig.height=2}
par(mar = c(3, 4, 0.1, 0.1))
plotdistcens(d, NPMLE = FALSE, lwd = 5, col = "blue", main = "") 
```

## Non Parametric Maximum Likelihood Estimation (NPMLE) of the ECDF: the Turnbull plot (package survival) used in former versions of fitdistrplus.
```{r, echo = FALSE,fig.height=5}
par(mar = c(3, 4, 0.1, 0.1))
layout(matrix(c(1,2,2,2), nrow =4, byrow = TRUE))
par(mar = c(3, 4, 0.1, 0.1))
plotdistcens(d, NPMLE = FALSE, lwd = 5, col = "blue", main = "")
addbounds(d)
plotdistcens(d, NPMLE.method = "Turnbull", lwd = 5, col = "orange", main = "")
addbounds(d)
```

## A new algorithm and plot from the package npsurv (Wang)
```{r, echo = FALSE}
# par(mfrow = c(2, 1))
layout(matrix(c(1,2,2,2), nrow =4, byrow = TRUE))
par(mar = c(3, 4, 0.1, 0.1))
plotdistcens(d, NPMLE = FALSE, lwd = 5, col = "blue", main = "")
addbounds(d)
plotdistcens(d, NPMLE = TRUE, lwd = 5, main = "")
addTurnbull(d)
addbounds(d)
legend("right", lty = c(1,2), lwd = c(5,5), col = c("black", "orange"),
       legend = c("Wang", "Turnbull"), bty = "n", cex = 2)
```

## The two steps of an NPMLE algorithm

1. Identification of **equivalence classes** (also named **Turnbull intervals** or **maximal intersection intervals** or **innermost intervals** or **maximal cliques** of the data) 
= set of points/intervals under which the ECDF may change (each region between a left bound \textcolor{red}{L} immediately followed by a right bound \textcolor{red}{R}, even if of null length). The NPMLE is unique only up to these equivalence classes (**non uniqueness** represented by **grey rectangles**).

Equivalence classes on the first toy example

```{r, echo = FALSE, fig.height = 2, fig.width = 8}
par(mar = c(3, 4, 0.1, 0.1))
plotdistcens(d, NPMLE = FALSE, lwd = 5, col = "blue", main = "")
addbounds(d)
addLR(d)
addeq(d)
```

## The two steps of an NPMLE algorithm

1. Identification of **equivalence classes** (also named **Turnbull intervals** or **maximal intersection intervals** or **innermost intervals** or **maximal cliques** of the data) 
= set of points/intervals under which the ECDF may change (each region between a left bound \textcolor{red}{L} immediately followed by a right bound \textcolor{red}{R}, even if of null length). The NPMLE is unique only up to these equivalence classes (**non uniqueness** represented by **grey rectangles**).

Equivalence classes on a second toy example

```{r, echo = FALSE, fig.height = 2, fig.width = 8}
d0b <- data.frame(left = c(NA, 2, 4, 6, 10), right = c(1, 3, 7, 8, NA))
d <- d0b
deq <- data.frame(left = c(NA, 2, 6, 10), right = c(1, 3, 7, NA))
par(mar = c(3, 4, 0.1, 0.1))
plotdistcens(d, NPMLE = FALSE, lwd = 5, col = "blue", main = "")
addbounds(d)
addLR(d)
addeq(deq)
```

## The two steps of an NPMLE algorithm

1. Identification of **equivalence classes** (also named **Turnbull intervals** or **maximal intersection intervals** or **innermost intervals** or **maximal cliques** of the data) 
= set of points/intervals under which the ECDF may change (each region between a left bound \textcolor{red}{L} immediately followed by a right bound \textcolor{red}{R}, even if of null length). The NPMLE is unique only up to these equivalence classes (**non uniqueness** represented by **grey rectangles**).

2. Assign a **probability mass** to each equivalence class (may be 0).

The Wang plot on the second toy example

```{r, echo = FALSE, fig.height = 2, fig.width = 8}
par(mar = c(3, 4, 0.1, 0.1))
plotdistcens(d, NPMLE = TRUE, lwd = 5, main = "")
addbounds(d)
```

## The two steps of an NPMLE algorithm

1. Identification of **equivalence classes** (also named **Turnbull intervals** or **maximal intersection intervals** or **innermost intervals** or **maximal cliques** of the data) 
= set of points/intervals under which the ECDF may change (each region between a left bound \textcolor{red}{L} immediately followed by a right bound \textcolor{red}{R}, even if of null length). The NPMLE is unique only up to these equivalence classes (**non uniqueness** represented by **grey rectangles**).

2. Assign a **probability mass** to each equivalence class (may be 0).

Various algorithms implemented in the packages **Icens**, **interval** and **npsurv** (more or less performant and not all handling left censored data).

## A third toy example
```{r, echo = FALSE}
d3 <- data.frame(left = c(-1.4, -1.4, 2, -1.4, 0),
                 right = c(1, 2, NA, 0, 2))
d <- d3
```
```{r, echo = FALSE}
# par(mfrow = c(2, 1))
layout(matrix(c(1,2,2,2), nrow =4, byrow = TRUE))
par(mar = c(3, 4, 0.1, 0.1))
plotdistcens(d, NPMLE = FALSE, lwd = 5, col = "blue", main = "")
addbounds(d)
addLR(d)
plotdistcens(d, NPMLE = TRUE, lwd = 5, main = "")
addTurnbull(d)
addbounds(d)
legend("right", lty = c(1,2), lwd = c(5,5), col = c("black", "orange"),
       legend = c("Wang", "Turnbull"), bty = "n", cex = 2)
```

## The third toy example with the add of a non censored obs.
```{r, echo = FALSE}
d3b <- data.frame(left = c(1.18, -1.4, -1.4, 2, -1.4, 0),
                 right = c(1.18, 1, 2, NA, 0, 2))
d <- d3b
```
```{r, echo = FALSE}
# par(mfrow = c(2, 1))
layout(matrix(c(1,2,2,2), nrow =4, byrow = TRUE))
par(mar = c(3, 4, 0.1, 0.1))
plotdistcens(d, NPMLE = FALSE, lwd = 5, col = "blue", main = "")
addbounds(d)
addLR(d)
plotdistcens(d, NPMLE = TRUE, lwd = 5, main = "")
addTurnbull(d)
addbounds(d)
legend("right", lty = c(1,2), lwd = c(5,5), col = c("black", "orange"),
       legend = c("Wang", "Turnbull"), bty = "n", cex = 2)
```

## A realistic example: data salinity
```{r, echo = FALSE}
data(salinity)
d <-  log10(salinity)
```
```{r, echo = FALSE}
# par(mfrow = c(2, 1))
layout(matrix(c(1,2,2,2), nrow =4, byrow = TRUE))
par(mar = c(3, 4, 0.1, 0.1))
plotdistcens(d, NPMLE = FALSE, lwd = 0.2, col = "blue", main = "")
addbounds(d)
addLR(d)
plotdistcens(d, NPMLE = TRUE, lwd = 5, main = "")
addbounds(d)
```

# New CDF, Q-Q and P-P plots implemented for censored data

## Use of cdfcompcens() to assess the fit of 3 distributions on data smokedfish
```{r, echo = FALSE}
data(smokedfish)
d <-  log10(smokedfish)
fitsfn <- fitdistcens(d,"norm")
fitsfl <- fitdistcens(d,"logis")
dgumbel <- function(x,a,b) 1/b*exp((a-x)/b)*exp(-exp((a-x)/b))
pgumbel <- function(q,a,b) exp(-exp((a-q)/b))
qgumbel <- function(p,a,b) a-b*log(-log(p))
fitsfg<-fitdistcens(d,"gumbel",start=list(a=-3,b=3))
```
```{r, echo = FALSE, fig.height = 3.5, fig.width = 5}
# par(mfrow = c(2, 1))
layout(matrix(c(1,2,2), nrow =3, byrow = TRUE))
par(mar = c(3, 4, 2, 0.1))
plotdistcens(d, NPMLE = FALSE, lwd = 0.2, col = "blue", main = "")
# addbounds(d)
cdfcompcens(list(fitsfn,fitsfl,fitsfg), fitlty = 1)
# addbounds(d)
```

## Use of qqcompcens() for one distribution
```{r, echo = FALSE, fig.height = 4.5, fig.width = 8}
par(mfrow = c(1, 2))
par(mar = c(4, 4, 2, 1))
cdfcompcens(fitsfn, fitlty = 1)
qqcompcens(fitsfn)
```

## Use of ppcompcens() for one distribution
```{r, echo = FALSE, fig.height = 4.5, fig.width = 8}
par(mfrow = c(1, 2))
par(mar = c(4, 4, 2, 1))
cdfcompcens(fitsfn, fitlty = 1)
ppcompcens(fitsfn)
```

## Q-Q plots and P-P plot for the 3 distributions
```{r, echo = FALSE, fig.height = 4.5, fig.width = 8}
par(mfrow = c(1, 2))
par(mar = c(4, 4, 2, 1))
qqcompcens(list(fitsfn,fitsfl,fitsfg))
ppcompcens(list(fitsfn,fitsfl,fitsfg))
```

## An alternative presentation of the Q-Q plots for the 3 dist. 
```{r, echo = FALSE, fig.height = 5.5, fig.width = 6}
par(mfrow = c(2,2))
par(mar = c(4, 4, 2, 1))
qqcompcens(fitsfn, fitcol = "red", fillrect = "red")
qqcompcens(fitsfl, fitcol = "green", fillrect = "green")
qqcompcens(fitsfg, fitcol = "blueviolet", fillrect = "blueviolet")
```

Will be soon implemented in the plotstyle ggplot.

## Another example with data salinity
```{r, echo = FALSE, fig.height = 5.5, fig.width = 6}
data(salinity)
d <-  log10(salinity)
f <- fitdistcens(d,"norm")
par(mfrow = c(2,2))
par(mar = c(4, 4, 2, 1))
plotdistcens(d, NPMLE = FALSE, main = "")
cdfcompcens(f)
qqcompcens(f)
ppcompcens(f)
```

## How to use of these new goodness-of-fit plots ?

Example of code:
```{r, eval = FALSE}
data(smokedfish)
d <-  log10(smokedfish)
# Plot of the NPMLE CDF on censored data
plotdistcens(d)
# Two MLE fits
fitsfn <- fitdistcens(d,"norm")
fitsfl <- fitdistcens(d,"logis")
# Three goodness-of-fit plots for one fit
plot(fitsfn)
# Goodness-of-fit plots for one or more fits
cdfcompcens(list(fitsfn,fitsfl))
qqcompcens(list(fitsfn,fitsfl))
ppcompcens(list(fitsfn,fitsfl))
```

## Other recent improvements of fitdistrplus

**Version 1.0-8**

* add of an optional use of ggplot2 in cdfcomp(), denscomp(), qqcomp() and ppcomp().

**Version 1.0-10**

* Improvement of goodness-of-fit plots for discrete distributions in denscomp().
* Add of new default starting values for distributions in actuar.

**Version 1.0-11**

* add of an optional use of ggplot2 in cdfcompcens(), denscompcens() and ppcompcens().

## References
* Turnbull BW (1974). Nonparametric estimation of a survivorship function with doubly censored data. Journal of American Statistical Association, 69, 169-173.
* Gentleman, R., & Geyer, C. J. (1994). Maximum likelihood for interval censored data: Consistency and computation. Biometrika, 81(3), 618-623.
* Wang, Y. (2008). Dimension-reduced nonparametric maximum likelihood computation for interval-censored data. Computational Statistics & Data Analysis, 52(5), 2388-2402.
* Wang, Y., & Taylor, S. M. (2013). Efficient computation of nonparametric survival functions via a hierarchical mixture formulation. Statistics and Computing, 23(6), 713-725.
* Wang, Y., & Fani, S. (2018). Nonparametric maximum likelihood computation of a U-shaped hazard function. Statistics and Computing, 28(1), 187-200.

## 
\begin{center}
\begin{Large}
Thank you for your attention !

We are waiting for your feedback on these new tools.
\end{Large}
\end{center}

### 2020-06-25 R/ppcompcens.R
Add a 'fitlwd' argument in the ppcompcens, qqcompcens and CIcdfplot functions

@@ -27,7 +27,7 @@
 
 
 ppcompcens <- function(ft, xlim, ylim, xlogscale = FALSE, ylogscale = FALSE, main, xlab, ylab, fillrect,
-                       fitcol, addlegend = TRUE, legendtext, xlegend = "bottomright", ylegend = NULL, 
+                       fitcol, fitlwd, addlegend = TRUE, legendtext, xlegend = "bottomright", ylegend = NULL, 
                        line01 = TRUE, line01col = "black", line01lty = 1, ynoise = TRUE, NPMLE.method = "Wang", 
                        plotstyle = "graphics", ...)
 {
@@ -81,7 +81,9 @@ ppcompcens <- function(ft, xlim, ylim, xlogscale = FALSE, ylogscale = FALSE, mai
   # manage default parameters
   nft <- length(ft)
   if (missing(fitcol)) fitcol <- 2:(nft+1)
+  if (missing(fitlwd)) fitlwd <- 1
   fitcol <- rep(fitcol, length.out=nft)
+  fitlwd <- rep(fitlwd, length.out=nft)
   if (missing(fillrect)) 
     if ((nft == 1) | plotstyle == "ggplot") fillrect <- "lightgrey" else fillrect <- NA
 
@@ -179,19 +181,19 @@ ppcompcens <- function(ft, xlim, ylim, xlogscale = FALSE, ylogscale = FALSE, mai
           noise2mult <- runif(nQi, 0.99, 1.01)
           rect(xleft = Pitheo.low, ybottom = Pi.low * noise2mult, 
                xright = Pitheo.up, ytop = Pi.up * noise2mult, 
-               border = fitcol[i], col = fillrect[i])
+               border = fitcol[i], col = fillrect[i], lwd = fitlwd[i])
         }
         else
         {
           noise2add <- runif(nQi, -0.01, 0.01)
           rect(xleft = Pitheo.low, ybottom = Pi.low + noise2add, 
                xright = Pitheo.up, ytop = Pi.up + noise2add, 
-               border = fitcol[i], col = fillrect[i])
+               border = fitcol[i], col = fillrect[i], lwd = fitlwd[i])
         }
       } else # ! ynoise
       {
         rect(xleft = Pitheo.low, ybottom = Pi.low, xright = Pitheo.up, ytop = Pi.up, 
[…]

### 2020-06-25 R/qqcompcens.R

@@ -27,7 +27,7 @@
 
 
 qqcompcens <- function(ft, xlim, ylim, xlogscale = FALSE, ylogscale = FALSE, main, xlab, ylab, fillrect,
-                       fitcol, addlegend = TRUE, legendtext, xlegend = "bottomright", ylegend = NULL, 
+                       fitcol, fitlwd, addlegend = TRUE, legendtext, xlegend = "bottomright", ylegend = NULL, 
                        line01 = TRUE, line01col = "black", line01lty = 1, ynoise = TRUE, 
                        NPMLE.method = "Wang", plotstyle = "graphics", ...)
 {
@@ -81,7 +81,9 @@ qqcompcens <- function(ft, xlim, ylim, xlogscale = FALSE, ylogscale = FALSE, mai
   # manage default parameters
   nft <- length(ft)
   if (missing(fitcol)) fitcol <- 2:(nft+1)
+  if (missing(fitlwd)) fitlwd <- 1
   fitcol <- rep(fitcol, length.out=nft)
+  fitlwd <- rep(fitlwd, length.out=nft)
   if (missing(fillrect)) 
     if ((nft == 1) | plotstyle == "ggplot") fillrect <- "lightgrey" else fillrect <- NA
   
@@ -231,7 +233,7 @@ qqcompcens <- function(ft, xlim, ylim, xlogscale = FALSE, ylogscale = FALSE, mai
           rect(xleft = Qitheo.left4plot, ybottom = Qi.left4plot * noise2mult, 
                xright = Qitheo.right4plot, 
                ytop = Qi.right4plot * noise2mult, 
-               border = fitcol[i], col = fillrect[i])
+               border = fitcol[i], col = fillrect[i], lwd = fitlwd[i])
         }
         else
         {
@@ -239,13 +241,13 @@ qqcompcens <- function(ft, xlim, ylim, xlogscale = FALSE, ylogscale = FALSE, mai
           rect(xleft = Qitheo.left4plot, ybottom = Qi.left4plot + noise2add, 
                xright = Qitheo.right4plot, 
                ytop = Qi.right4plot + noise2add, 
-               border = fitcol[i], col = fillrect[i])
+               border = fitcol[i], col = fillrect[i], lwd = fitlwd[i])
         }
       } else # ! ynoise
       {
[…]

### 2020-06-25 tests/t-ppcompcens.R

@@ -44,6 +44,7 @@ log10LC50 <-log10(salinity)
 fn <- fitdistcens(log10LC50,"norm")
 fl <- fitdistcens(log10LC50,"logis")
 ppcompcens(list(fn, fl))
+ppcompcens(list(fn, fl), fitlwd = c(5, 2))
 
 if (visualize)
 {
@@ -59,6 +60,7 @@ if (visualize)
 
 if (requireNamespace ("ggplot2", quietly = TRUE) ) {
   ppcompcens(list(fn, fl), plotstyle = "ggplot", fitcol = "red")
+  ppcompcens(list(fn, fl), plotstyle = "ggplot", fitlwd = c(5, 2))
 }
 
 

### 2020-06-25 tests/t-qqcompcens.R

@@ -48,14 +48,17 @@ plot(fl)
 qqcompcens(fn)
 qqcompcens(fl)
 qqcompcens(list(fn, fl))
+qqcompcens(list(fn, fl), fitlwd = c(5, 2))
 
 if (requireNamespace("ggplot2", quietly = TRUE)) {
   qqcompcens(list(fn, fl), plotstyle = "ggplot")
+  qqcompcens(list(fn, fl), plotstyle = "ggplot", fitlwd = c(5, 2))
 }
 
 if (requireNamespace("ggplot2", quietly = TRUE) & visualize) {
   qqcompcens(fl, plotstyle = "ggplot")
   qqcompcens(fn, plotstyle = "ggplot")
+  
 }
 
 

### 2022-02-23 R/detectbound.R
Replace 'if(class())' by 'inherits()'

@@ -82,7 +82,7 @@ detectbound <- function(distname, vstart, obs, fix.arg=NULL, echo=FALSE)
         print(dx2)
       }
       
-      if(class(dx1) == "try-error" && class(dx2) != "try-error")
+      if(inherits(dx1, "try-error") && !inherits(dx2, "try-error"))
       {
         lowb[a] <- aval[i]
       }
@@ -90,7 +90,7 @@ detectbound <- function(distname, vstart, obs, fix.arg=NULL, echo=FALSE)
       {
         lowb[a] <- aval[i]
       }
-      if(class(dx1) != "try-error" && class(dx2) == "try-error")
+      if(!inherits(dx1, "try-error") && inherits(dx2, "try-error"))
       {
         uppb[a] <- aval[i]
       }

### 2022-02-23 R/util-testdensity.R

@@ -68,7 +68,7 @@ test1fun <- function(fn, start.arg, fix.arg, dpqr)
   res0 <- try(do.call(fn, c(list(numeric(0)), start.arg, fix.arg)), silent=TRUE)
   t0 <- paste("The", fn, "function should return a zero-length vector when input has length zero and not raise an error")
   t1 <- paste("The", fn, "function should return a zero-length vector when input has length zero")
-  if(class(res0) == "try-error")
+  if(inherits(res0, "try-error"))
   {
     res$txt <- t0
     return(res)
@@ -83,7 +83,7 @@ test1fun <- function(fn, start.arg, fix.arg, dpqr)
   x <- c(0, 1, Inf, NaN, -1)
   res1 <- try(do.call(fn, c(list(x), start.arg, fix.arg)), silent=TRUE)
   t2 <- paste("The", fn, "function should return a vector of with NaN values when input has inconsistent values and not raise an error")
-  if(class(res1) == "try-error")
+  if(inherits(res1, "try-error"))
   {
     res$txt <- t2
     return(res)
@@ -94,7 +94,7 @@ test1fun <- function(fn, start.arg, fix.arg, dpqr)
   res2 <- try(do.call(fn, c(list(x), start.arg, fix.arg)), silent=TRUE)
   t4 <- paste("The", fn, "function should return a vector of with NA values when input has missing values and not raise an error")
   t5 <- paste("The", fn, "function should return a vector of with NA values when input has missing values and not remove missing values")
-  if(class(res2) == "try-error")
+  if(inherits(res2, "try-error"))
   {
     res$txt <- t4
     return(res)
@@ -110,7 +110,7 @@ test1fun <- function(fn, start.arg, fix.arg, dpqr)
   start.arg <- lapply(start.arg, function(x) -x)
   res3 <- try(do.call(fn, c(list(x), start.arg, fix.arg)), silent=TRUE)
   t6 <- paste("The", fn, "function should return a vector of with NaN values when input has inconsistent parameters and not raise an error")
-  if(class(res3) == "try-error")
+  if(inherits(res3, "try-error"))
   {
     res$txt <- t6
     return(res)
@@ -121,7 +121,7 @@ test1fun <- function(fn, start.arg, fix.arg, dpqr)
   names(start.arg) <- paste0(names(start.arg), "_")
   res4 <- try(do.call(fn, c(list(x), start.arg, fix.arg)), silent=TRUE)
[…]

### 2022-02-23 tests/t-fitdist.R

@@ -137,7 +137,7 @@ if (visualize) # LONG TO RUN ON CRAN
   {
     cat("\nscaling", 10^i, "\n")
     res <- try(mledist(x2*10^i, "cauchy"), silent=TRUE)
-    if(class(res) == "try-error")
+    if(inherits(res, "try-error"))
       print(res)
     else
     {
@@ -166,7 +166,7 @@ if (visualize) # LONG TO RUN ON CRAN
     cat("\nscaling", 10^(-2*i), "\n")
     res <- mledist(x1*10^(-2*i), "norm")
     Hm1 <- try(solve(res$hessian), silent=TRUE)
-    if(class(Hm1) == "try-error")
+    if(inherits(Hm1, "try-error"))
       print(Hm1)
     else
     {

### 2022-02-23 tests/t-init-actuar.R

@@ -24,7 +24,7 @@ if(FALSE)
     f <- function(eta)
       (qi[1]^eta-1)^y1*(qi[2]^eta-1)^y2*(qi[3]^eta-1) - 1
     eta <- try(uniroot(f, c(-10, -1e-6))$root, silent=TRUE)
-    if(class(eta) == "try-error")
+    if(inherits(eta, "try-error"))
       eta <- -1
     alpha <- -1/eta
     lambda <- (qi[1]^eta-1)^y*(qi[2]^eta-1)^(-y-1)

### 2023-04-25 tests/t-init-actuar.R
Remove dots in the 'start.arg.default' function

@@ -6,7 +6,7 @@ if(FALSE)
   
   #test actuar initialization
   
-  start.arg.default <- fitdistrplus:::start.arg.default
+  startargdefault <- fitdistrplus:::startargdefault
   
   #burr
   library(actuar)
@@ -40,7 +40,7 @@ if(FALSE)
   #transformed gamma
   x <- rtrgamma(1000, 3, 2, .1)
   
-  start1 <- start.arg.default(x, "trgamma")
+  start1 <- startargdefault(x, "trgamma")
   
   plot(ecdf(x))
   curve(ptrgamma(x, start1$shape1, start1$shape2, start1$rate), add=TRUE, lty=2)
@@ -50,7 +50,7 @@ if(FALSE)
   #inverse transformed gamma
   y <- rinvtrgamma(1000, 3, 2, .1)
   
-  start1 <- start.arg.default(y, "invtrgamma")
+  start1 <- startargdefault(y, "invtrgamma")
   
   plot(ecdf(y))
   curve(pinvtrgamma(x, start1$shape1, start1$shape2, start1$rate), add=TRUE, lty=2)

### 2024-02-22 share/startingvalues/starting-value-fellerpareto.R
new starting value for all actuar distribution

@@ -4,7 +4,7 @@ library(fitdistrplus)
 n <- 1e2
 U <- rgamma(n, 1/2)
 V <- rgamma(n, 3)
-X <- rfpareto(n, 0, 1/2, 3/2, 5/2, 1)
+X <- rfpareto(n, 0, 1/2, 3/2, 5/2, scale=2)
 
 hist(log(U/V))
 hist(-log(1-U/V))
@@ -34,15 +34,34 @@ thetarel <- function(x, alpha, gamma)
   ( ((3/4)^(-1/alpha) - 1)^(1/gamma) - ((1/4)^(-1/alpha) - 1)^(1/gamma) )/(q1 - q3)
 }  
 
-thetahat <- thetarel(X, 2, 2)
+thetarel2 <- function(x, mu)
+{
+  y <- 2*x*(x-mu)/(1+(x-mu)^2)
+  z <- x/(x-mu)
+  sqrt(mean(y)/(mean(z)-1))
+}
+
+
+thetahat <- thetarel2(X, min(X)*.95)
 
 Y <- (X-min(X)*.95)/thetahat
-Z <- Y/(1+Y)
-Y <- (X-min(X)*.95)/sd(X)
 Z <- Y/(1+Y)
 Y <- (X-3)/10
 Z <- Y/(1+Y)
 
+cdfcomp(fitdist(Z^2, "beta"))
 cdfcomp(fitdist(Z, "beta"))
 coef(fitdist(Z, "beta"))
+coef(fitdist(Z^2, "beta"))
 
+tauhat <- coef(fitdist(Z, "beta"))[1]
+alphahat <- coef(fitdist(Z, "beta"))[2]
+
+gammarel <- function(x, mu, theta, alpha, tau)
+{
+  y <- log(1+(x-mu)/theta)
+  z <- log((x-mu)/theta)
+  1+(digamma(tau) - digamma(alpha+tau) + mean(y))/mean(z)
+}
+
+gammarel(X, min(X)*.9, thetahat, alphahat, tauhat)

### 2024-02-22 share/startingvalues/starting-value-zerotrunc.R

@@ -1,0 +1,9 @@
+library(actuar)
+
+n <- 1e3
+x <- rztpois(n, 2)
+
+hist(x)
+hist(rpois(n, 2)+1)
+
+cdfcomp(fitdist(x-1, "pois"))

### 2024-02-22 tests/t-startingvalues-zeromod-family.R

@@ -1,0 +1,31 @@
+library(fitdistrplus)
+library(actuar)
+
+set.seed(1234)
+
+n <- 1e3
+x <- rzmpois(n, 2, 1/2)
+
+fitdistrplus:::startarg_discrete_actuar_family(x, "zmpois")
+fitdist(x, "zmpois")
+
+x <- rzmnbinom(n, 2, 1/2, 1/3)
+fitdistrplus:::startarg_discrete_actuar_family(x, "zmnbinom")
+fitdist(x, "zmnbinom")
+
+x <- rzmgeom(n, 1/2, 1/3)
+fitdistrplus:::startarg_discrete_actuar_family(x, "zmgeom")
+fitdist(x, "zmgeom")
+
+
+x <- rzmbinom(n, 30, 1/2, 1/3)
+
+fitdistrplus:::startarg_discrete_actuar_family(x, "zmbinom")
+fitdist(x, "zmbinom", lower=0)
+
+
+x <- rpoisinvgauss(n, mean = 10, dispersion = 2)
+
+fitdistrplus:::startarg_discrete_actuar_family(x, "poisinvgauss")
+fitdist(x, "poisinvgauss", lower=0)
+

### 2024-02-22 tests/t-startingvalues-zerotrunc-family.R

@@ -1,0 +1,25 @@
+library(fitdistrplus)
+library(actuar)
+
+set.seed(1234)
+
+n <- 1e3
+x <- rztpois(n, 2)
+
+fitdistrplus:::startarg_discrete_actuar_family(x, "ztpois")
+fitdist(x, "ztpois")
+
+
+fitdistrplus:::startarg_discrete_actuar_family(x, "ztnbinom")
+fitdist(x, "ztnbinom")
+
+
+fitdistrplus:::startarg_discrete_actuar_family(x, "ztgeom")
+fitdist(x, "ztgeom")
+
+
+x <- rztbinom(n, 30, 1/2)
+
+fitdistrplus:::startarg_discrete_actuar_family(x, "ztbinom")
+fitdist(x, "ztbinom", lower=0)
+

### 2024-06-10 R/bootdist.R
layout and remove a bug in optimalgo.Rmd

@@ -47,14 +47,15 @@ bootdist <- function (f, bootmethod="param", niter=1001, silent=TRUE,
     stop("Bootstrap is not yet available when using weights")
     
     #simulate bootstrap data
-    if (bootmethod == "param") { # parametric bootstrap
+    if (bootmethod == "param") 
+    { # parametric bootstrap
         rdistname <- paste("r", f$distname, sep="")
         if (!exists(rdistname, mode="function"))
             stop(paste("The ", rdistname, " function must be defined"))
         rdata <- do.call(rdistname, c(list(niter*f$n), as.list(f$estimate), f$fix.arg))
         dim(rdata) <- c(f$n, niter)
-    }
-    else { # non parametric bootstrap
+    }else 
+    { # non parametric bootstrap
         rdata <- sample(f$data, size=niter*f$n, replace=TRUE)
         dim(rdata) <- c(f$n, niter)
     }
@@ -139,14 +140,15 @@ bootdist <- function (f, bootmethod="param", niter=1001, silent=TRUE,
     options(warn=owarn, show.error.messages=oerr)
     
     rownames(resboot) <- c(names(start), "convergence")
-    if (length(resboot[, 1])>2) {
+    if (length(resboot[, 1]) > 2)
+    {
         estim <- data.frame(t(resboot)[, -length(resboot[, 1])])
         bootCI <- cbind(apply(resboot[-length(resboot[, 1]), ], 1, median, na.rm=TRUE), 
         apply(resboot[-length(resboot[, 1]), ], 1, quantile, 0.025, na.rm=TRUE), 
         apply(resboot[-length(resboot[, 1]), ], 1, quantile, 0.975, na.rm=TRUE))
         colnames(bootCI) <- c("Median", "2.5%", "97.5%")
-    }
-    else {
+    }else 
+    {
         estim <- as.data.frame(t(resboot)[, -length(resboot[, 1])])
         names(estim) <- names(f$estimate)
         bootCI <- c(median(resboot[-length(resboot[, 1]), ], na.rm=TRUE), 
@@ -164,7 +166,8 @@ bootdist <- function (f, bootmethod="param", niter=1001, silent=TRUE,
     res    
 }
 
-print.bootdist <- function(x, ...){
+print.bootdist <- function(x, ...)
+{
     if (!inherits(x, "bootdist"))
         stop("Use only with 'bootdist' objects")
     if (x$method=="param") 
[…]

### 2024-06-12 R/bootdist.R
add a density function for bootdist* objects

@@ -26,170 +26,170 @@
 bootdist <- function (f, bootmethod="param", niter=1001, silent=TRUE, 
                       parallel = c("no", "snow", "multicore"), ncpus)
 { 
-    if (niter<10) 
-        stop("niter must be an integer above 10")
-    bootmethod <- match.arg(bootmethod, c("param", "nonparam"))
-
-    parallel <- match.arg(parallel, c("no", "snow", "multicore"))
-    if (parallel == "multicore" & .Platform$OS.type == "windows")
-    {
-        parallel <- "snow"
-        warning("As the multicore option is not supported on Windows it was replaced by snow")
-    }
-    if ((parallel == "snow" | parallel == "multicore") & missing(ncpus)) 
-        stop("You have to specify the number of available processors to parallelize 
+  if (niter<10) 
+    stop("niter must be an integer above 10")
+  bootmethod <- match.arg(bootmethod, c("param", "nonparam"))
+  
+  parallel <- match.arg(parallel, c("no", "snow", "multicore"))
+  if (parallel == "multicore" & .Platform$OS.type == "windows")
+  {
+    parallel <- "snow"
+    warning("As the multicore option is not supported on Windows it was replaced by snow")
+  }
+  if ((parallel == "snow" | parallel == "multicore") & missing(ncpus)) 
+    stop("You have to specify the number of available processors to parallelize 
              the bootstrap")
-            
-    if (!inherits(f, "fitdist"))
-        stop("Use only with 'fitdist' objects")
-     
-    if(!is.null(f$weights))
+  
+  if (!inherits(f, "fitdist"))
+    stop("Use only with 'fitdist' objects")
+  
+  if(!is.null(f$weights))
     stop("Bootstrap is not yet available when using weights")
-    
-    #simulate bootstrap data
-    if (bootmethod == "param") 
-    { # parametric bootstrap
-        rdistname <- paste("r", f$distname, sep="")
-        if (!exists(rdistname, mode="function"))
-            stop(paste("The ", rdistname, " function must be defined"))
-        rdata <- do.call(rdistname, c(list(niter*f$n), as.list(f$estimate), f$fix.arg))
-        dim(rdata) <- c(f$n, niter)
-    }else 
-    { # non parametric bootstrap
-        rdata <- sample(f$data, size=niter*f$n, replace=TRUE)
[…]

### 2024-06-12 tests/t-bootdist.R

@@ -24,7 +24,7 @@ print(lapply(b1, head))
 plot(b1)
 summary(b1)
 
-# (1) bis test new plot arguments
+# (2) new plot arguments
 #for new graph functions
 f1 <- fitdist(rgamma(nsample, 2, 3), "gamma")
 b1 <- bootdist(f1, niter=nbboot, silent=TRUE)
@@ -299,4 +299,20 @@ xval <- sort(unique(x))
 summary(bootdist(f1, niter = nbboot))
 try(summary(bootdist(f2, niter = nbboot))) # not yet developed
 
+# (18) density of bootdist()
+#
+x <- rlnorm(1e3)
+b0 <- bootdist(fitdist(x, "lnorm"), niter = 50)
+b1 <- bootdist(fitdist(x, "lnorm"), niter = 100)
+b2 <- bootdist(fitdist(x, "lnorm"), niter = 200)
 
+#d1 <- fitdistrplus:::density.bootdist(b0, b1, b2)
+d1 <- density(b0, b1, b2)
+str(d1)
+plot(d1)
+print(d1)
+
+d1 <- density(b1)
+str(d1)
+plot(d1)
+print(d1)

### 2024-06-26 tests/t-startingvalues-trgamma-family.R
add a new argument

@@ -20,3 +20,24 @@ fitdist(x, "weibull")
 fitdist(x, "exp")
 
 
+#weird examples
+x <- rep(1, n)
+
+fitdistrplus:::startarg_transgamma_family(x, "gamma")
+
+#previous code
+n <- length(x)
+m <- mean(x)
+v <- (n - 1)/n*var(x)
+list(shape=m^2/v, rate=m/v)
+
+
+#normal -> weibull
+x <- abs(rnorm(n = 20, mean = 150, sd = 10))
+m <- mean(log(x))
+v <- var(log(x))
+shape <- 1.2/sqrt(v)
+scale <- exp(m + 0.572/shape)
+s <- list(shape=shape, scale=scale)
+quantile(x)
+fitdistrplus:::startarg_transgamma_family(x, "weibull")

### 2024-06-26 R/util-startarg.R
get back to original estimator for Weibull

@@ -473,20 +473,20 @@ startarg_transgamma_family <- function(x, distr)
   {
     if (any(x < 0)) 
       stop("values must be positive to fit a Weibull distribution")
-    q25 <- as.numeric(quantile(x, 0.25, na.rm = TRUE))
-    q75 <- as.numeric(quantile(x, 0.75, na.rm = TRUE))
-    if(q25 < q75) #check to avoid division by zero
-    {
-      q50 <- median(x, na.rm = TRUE)
-      tauhat <- (log(-log(1-1/4)) - log(-log(1-3/4))) / (log(q25) - log(q75))
-      thetahat <- q50/(-log(1-1/2))^(tauhat)
-    }else
-    {
+    #q25 <- as.numeric(quantile(x, 0.25, na.rm = TRUE))
+    #q75 <- as.numeric(quantile(x, 0.75, na.rm = TRUE))
+    #if(q25 < q75) #check to avoid division by zero
+    #{
+    #  q50 <- median(x, na.rm = TRUE)
+    #  tauhat <- (log(-log(1-1/4)) - log(-log(1-3/4))) / (log(q25) - log(q75))
+    #  thetahat <- q50/(-log(1-1/2))^(tauhat)
+    #}else
+    #{
       m <- mean(log(x), na.rm = TRUE)
       v <- var(log(x), na.rm = TRUE)
       tauhat <- 1.2/sqrt(v)
       thetahat <- exp(m + 0.572/tauhat)
-    }
+    #}
     
     start <- list(shape = tauhat, scale = thetahat)
   }else if (distr == "exp") 

### 2024-06-26 tests/t-startingvalues-trgamma-family.R

@@ -38,6 +38,6 @@ m <- mean(log(x))
 v <- var(log(x))
 shape <- 1.2/sqrt(v)
 scale <- exp(m + 0.572/shape)
-s <- list(shape=shape, scale=scale)
-quantile(x)
-fitdistrplus:::startarg_transgamma_family(x, "weibull")
+s1 <- list(shape=shape, scale=scale)
+cdfcomp(fitdist(x, "weibull", start=s1))
+

### 2024-07-29 R/util-startarg.R
bug fix in startarg gamma

@@ -481,7 +481,7 @@ startarg_transgamma_family <- function(x, distr)
     if(v > 0)
     {
       alphahat <- m^2/v
-      thetahat <- m/v
+      thetahat <- v/m
     }else #exponential case
     {
       alphahat <- 1 

### 2024-10-01 tests/t-CIcdfplot.R
Replace 'library()' with 'require()' which is (about 5 times) faster

@@ -1,4 +1,4 @@
-library(fitdistrplus)
+require("fitdistrplus")
 
 nbboot <- 201
 nbboot <- 10

### 2024-10-01 tests/t-Surv2fitdistcens.R

@@ -1,4 +1,4 @@
-require(fitdistrplus)
+require("fitdistrplus")
 data("fremale")
 
 # fremale test

### 2024-10-01 tests/t-bootdistcens.R

@@ -1,4 +1,4 @@
-library(fitdistrplus)
+require("fitdistrplus")
 nbboot <- 101
 nbboot <- 11
 nsample <- 10

### 2024-10-01 tests/t-cdfcompcens.R

@@ -1,4 +1,4 @@
-library(fitdistrplus)
+require("fitdistrplus")
 
 visualize <- FALSE # TRUE for manual tests with visualization of results
 
@@ -43,7 +43,7 @@ if (visualize)
 {
   if(any(installed.packages()[,"Package"] == "actuar"))
   {
-    require(actuar)
+    require("actuar")
     data(smokedfish)
     fln <- fitdistcens(smokedfish,"lnorm")
     fll <- fitdistcens(smokedfish,"llogis")

### 2024-10-01 tests/t-cvg-algo.R

@@ -1,9 +1,7 @@
 
 if(FALSE)
 {  
-  library(fitdistrplus)
-  
-  
+    require("fitdistrplus")
   
   #(1) beta distribution
   #

### 2024-10-01 tests/t-descdist.R

@@ -1,4 +1,4 @@
-library(fitdistrplus)
+require("fitdistrplus")
 nbboot <- 100
 nbboot <- 10
 nsample <- 10

### 2024-10-01 tests/t-fitdist-hessianpb.R

@@ -1,4 +1,4 @@
-library(fitdistrplus)
+require("fitdistrplus")
 
 
 
@@ -27,7 +27,7 @@ vcov(fitdist(serving, "gamma", start=as.list(f1$estimate)))
 #an example with lower bounds for ztbinom
 if(FALSE)
 {
-  library(actuar)
+    require("actuar")
   n <- 1e3
   x <- rztbinom(n, 30, 1/2)
   lnl(c(35, 1/2), obs=x, dfun=dztbinom)

### 2024-10-01 tests/t-fitdist-test-arguments.R

@@ -1,4 +1,4 @@
-library(fitdistrplus)
+require("fitdistrplus")
 nbboot <- 100
 nbboot <- 10
 

### 2024-10-01 tests/t-fitdistcens.R

@@ -1,4 +1,4 @@
-library(fitdistrplus)
+require("fitdistrplus")
 
 nsample <- 500
 nsample <- 10
@@ -35,7 +35,7 @@ log10EC50 <-log10(fluazinam)
   #wrap genoud function rgenoud package
   mygenoud  <-  function(fn, par, ...) 
   {
-    require(rgenoud)
+    require("rgenoud")
     res  <-  genoud(fn, starting.values=par, ...)        
     standardres  <-  c(res, convergence=0)
     

### 2024-10-01 tests/t-lnL-surf.R

@@ -1,4 +1,4 @@
-library(fitdistrplus)
+require("fitdistrplus")
 nsample <- 1000
 nsample <- 10
 

### 2024-10-01 tests/t-manageparam.R

@@ -1,5 +1,4 @@
-library(fitdistrplus)
-
+require("fitdistrplus")
 
 manageparam <- fitdistrplus:::manageparam
 

### 2024-10-01 tests/t-mledist-asymptotic-vcov.R

@@ -1,4 +1,4 @@
-require(fitdistrplus)
+require("fitdistrplus")
 nsample <- 1e6
 nsample <- 10
 
@@ -14,7 +14,7 @@ f1$vcov
 
 if(any(installed.packages()[, "Package"] == "actuar"))
 {
-  require(actuar)
+  require("actuar")
   #simulate a sample
   x4 <- rpareto(nsample, 6, 2)
   

### 2024-10-01 tests/t-msedist.R

@@ -1,4 +1,4 @@
-library(fitdistrplus)
+require("fitdistrplus")
 nboot <- 1000
 nboot <- 10
 nsample <- 10
@@ -47,7 +47,7 @@ abline(v=msedist(x1, "exp")$estimate, lty=2, col="blue")
 legend("bottomright", lty=1:2, col=c("green", "blue"), leg=c("theoretical value", "fitted value"))
 
 
-# library(BMT)
+# require("BMT")
 # x <- rBMT(1e3, 1/4, 3/4)
 # 
 # BMTfit.mpse(x)
@@ -101,7 +101,7 @@ plot(mse_lnorm_boot, enhance = TRUE, trueval=c(0,1))
 #--------------------------------------------------------
 # Pareto sample
 
-library(actuar)
+require("actuar")
 
 x1 <- rburr(nsample, 2,2,2)
 
@@ -178,7 +178,7 @@ legend("bottomright", lty=1, col=c("red", "green", "blue"), leg=c("MLE", "MSE no
 
 #--------------------------------------------------------
 # real dataset
-# library(CASdatasets)
+# require("CASdatasets")
 # data("ushustormloss")
 # x <- ushustormloss$Normalized.CL05
 # 

### 2024-10-01 tests/t-ppcomp.R

@@ -1,4 +1,4 @@
-library(fitdistrplus)
+require("fitdistrplus")
 
 # ?ppcomp
 

### 2024-10-01 tests/t-qqcomp.R

@@ -1,4 +1,4 @@
-library(fitdistrplus)
+require("fitdistrplus")
 
 # ?qqcomp
 visualize <- FALSE # TRUE for manual tests with visualization of results

### 2024-10-01 tests/t-startfixarg-overall.R

@@ -1,4 +1,4 @@
-library(fitdistrplus)
+require("fitdistrplus")
 
 # (1) non-censored data (continuous)
 #

### 2024-10-01 tests/t-startingvalues-fellerpareto-family.R

@@ -1,5 +1,5 @@
-library(fitdistrplus)
-library(actuar)
+require("fitdistrplus")
+require("actuar")
 
 set.seed(1234)
 

### 2024-10-01 tests/t-startingvalues.R

@@ -1,6 +1,5 @@
-library(fitdistrplus)
-library(actuar)
-
+require("fitdistrplus")
+require("actuar")
 
 x <- c(3.1334614, 1.0300544, 0.8839272, 0.7503320, 1.9378476, 1.7571313, 0.5369516, 
        3.1181863, 1.7412457, 1.1514215, 1.6252670, 1.3333488, 3.0594471, 1.0479867, 
@@ -34,5 +33,3 @@ x <- c(1.083905, 1.266773, 1.044362, 1.557070, 1.227676, 1.195054, 1.241101,
 #equivalent to Pareto type 1                  
 ft_par1 <-fitdist(x, "pareto1", upper=c(Inf, min(x)), fix.arg=list(min = 1),
                   control=list(trace=1, REPORT=1), silent=FALSE) 
-
-

### 2024-10-01 tests/t-util-npsurv-mainfunction.R

@@ -1,4 +1,4 @@
-library(fitdistrplus)
+require("fitdistrplus")
 vizualise <- FALSE
 mytrace <- 1
 mytrace <- 0
@@ -45,7 +45,7 @@ if(vizualise)
   
   # if(FALSE)
   # {
-  #   require(npsurv)
+  #   require("npsurv")
   #   rescheck <- npsurv::npsurv(ap, w=1, maxit=100, tol=1e-6, verb=3) 
   #   c(resap$ll, rescheck$ll)
   #   
@@ -129,7 +129,7 @@ if(vizualise)
   
   # if(FALSE)
   # {
-  #   require(npsurv)
+  #   require("npsurv")
   #   rescheck <- npsurv::npsurv(fakedata, w=1, maxit=100, tol=1e-6, verb=3)
   #   c(resfk$ll, rescheck$ll)
   # 
@@ -171,7 +171,7 @@ if(vizualise)
   
   # if(FALSE)
   # {
-  #   require(npsurv)
+  #   require("npsurv")
   #   d8bis <- d8
   #   d8bis$left[is.na(d8bis$left)] <- -Inf
   #   d8bis$right[is.na(d8bis$right)] <- Inf

### 2024-10-25 tests/t-cdfcomp.R
finish scaling of objective function

@@ -235,12 +235,12 @@ set.seed(1234)
 x2 <- c(rnorm(nsample, 5),  rnorm(nsample, 10))
 # MLE fit
 fit1 <- fitdist(x2, "norm2", "mle", start=list(poid=1/3, m1=4, s1=2, m2=8, s2=2), 
-                lower=c(0, 0, 0, 0, 0))
+                optim="L-BFGS-B", lower=c(0, 0, 0, 0, 0))
 fit2 <- fitdist(x2, "norm2", "qme", probs=c(1/6, 1/4, 1/3, 1/2, 2/3), 
-                start=list(poid=1/3, m1=4, s1=2, m2=8, s2=2), 
-                lower=c(0, 0, 0, 0, 0), upper=c(1/2, Inf, Inf, Inf, Inf))
+                start=list(poid=1/3, m1=4, s1=2, m2=8, s2=2), optim="L-BFGS-B",  
+                lower=c(0, 0, 0, 0, 0), upper=c(1, Inf, Inf, Inf, Inf))
 fit3 <- fitdist(x2, "norm2", "mge", gof="AD", 
-                start=list(poid=1/3, m1=4, s1=2, m2=8, s2=2), 
+                start=list(poid=1/3, m1=4, s1=2, m2=8, s2=2), optim="L-BFGS-B", 
                 lower=c(0, 0, 0, 0, 0), upper=c(1/2, Inf, Inf, Inf, Inf))
 
 cdfcomp(list(fit1, fit2, fit3), datapch=".")

### 2024-10-25 tests/t-fitdistcens.R

@@ -12,6 +12,7 @@ data(smokedfish)
 fitsf  <-  fitdistcens(smokedfish, "lnorm")
 summary(fitsf)
 
+#fitted coef is -3.63    3.54, fitted loglik is -90.7
 
 smokedfish2 <- tail(smokedfish, 20)
 
@@ -30,25 +31,29 @@ fitsf  <-  fitdistcens(smokedfish, "lnorm", calcvcov = FALSE)
 # (6) custom optimisation function - example with the genetic algorithm
 #
 data(fluazinam)
-log10EC50 <-log10(fluazinam)
+log10EC50 <- log10(fluazinam)
 
-  #wrap genoud function rgenoud package
-  mygenoud  <-  function(fn, par, ...) 
-  {
-    require("rgenoud")
-    res  <-  genoud(fn, starting.values=par, ...)        
-    standardres  <-  c(res, convergence=0)
-    
-    return(standardres)
-  }
+summary(fitdistcens(log10EC50, "logis"))
+
+#fitted coef is 2.152 0.691 , fitted loglik is -20.6
+
+#wrap genoud function rgenoud package
+mygenoud  <-  function(fn, par, ...) 
+{
+  require("rgenoud")
+  res  <-  genoud(fn, starting.values=par, ...)        
+  standardres  <-  c(res, convergence=0)
   
-  # call fitdistcens with a 'custom' optimization function
-  fit.with.genoud <- fitdistcens(log10EC50, "logis", custom.optim=mygenoud, nvars=2, 
-                                 start=list(location=1, scale=1),    
-                                 Domains=cbind(c(0,0), c(5, 5)), boundary.enforcement=1, 
-                                 print.level=1, hessian=TRUE)
-  
-  summary(fit.with.genoud)
+  return(standardres)
+}
+
+# call fitdistcens with a 'custom' optimization function
+fit.with.genoud <- fitdistcens(log10EC50, "logis", custom.optim=mygenoud, nvars=2, 
+                               start=list(location=1, scale=1),    
+                               Domains=cbind(c(0,0), c(5, 5)), boundary.enforcement=1, 
+                               print.level=1, hessian=TRUE)
+
+summary(fit.with.genoud)
 
 
 # (9) check keepdata
@@ -66,6 +71,7 @@ if (visualize) # LONG TO RUN ON CRAN AND NEEDS VISALIZATION OF RESULTS
   plot(f1)
[…]

### 2024-10-25 tests/t-mledist-cens.R

@@ -13,6 +13,7 @@ qgumbel  <-  function(p,a,b) a-b*log(-log(p))
 mledist(log10EC50, "gumbel", start=list(a=0,b=2), optim.method="Nelder-Mead")
 mledist(log10EC50, "gumbel", start=list(a=0,b=2)) #default NM
 
+#fitted coef is 1.63 1.14 , fitted loglik is -20.3
 
 try(mledist(rbind(log10EC50, c("a", "b")), "gamma"))
 try(mledist(rbind(log10EC50, c(NA, NA)), "gamma"))
@@ -25,6 +26,8 @@ try(mledist(rbind(log10EC50, c(3, NaN)), "gamma"))
 
 mledist(log10EC50, "lnorm", optim.method="BFGS")
 mledist(log10EC50, "lnorm", optim.method="Nelder")
+
+#fitted coef is 0.623   0.928, fitted loglik is -21.1
 
 #optim() is used
 mledist(log10EC50, "lnorm", optim.method="L-BFGS-B", lower=0)
@@ -40,6 +43,8 @@ xleft <- c(-1.8, -0.6, -0.1, 0.07, 0.14, 1, 1.2, 1.2, 1.2)
 xright <- c(-1.8, -0.6, -0.1, 0.07, 0.14, 1, NA, NA, NA)
 d <- data.frame(left = xleft, right = xright)
 f1 <- mledist(d, "norm")
+
+#fitted coef is 0.545 1.342, fitted loglik is -12.9
 
 dbis <- data.frame(left = c(-1.8, -0.6, -0.1, 0.07, 0.14, 1, 1.2), 
                    right = c(-1.8, -0.6, -0.1, 0.07, 0.14, 1, NA))

### 2024-10-25 tests/t-mledist-nocens.R

@@ -15,9 +15,16 @@ mledist(x1,"norm")
 # (2) defining your own distribution functions, here for the Gumbel distribution
 # for other distributions, see the CRAN task view dedicated to probability distributions
 
-dgumbel <- function(x,a,b) 1/b*exp((a-x)/b)*exp(-exp((a-x)/b))
+dgumbel <- function(x,a,b)
+{
+  cat(a, b, "\n")
+  res <- 1/b*exp((a-x)/b)*exp(-exp((a-x)/b))
+  if(any(is.infinite(log(res))))
+    stop(log(res))
+  res
+}
 mledist(x1,"gumbel",start=list(a=10,b=5), silent=TRUE)
-mledist(x1,"gumbel",start=list(a=10,b=5), silent=FALSE)
+mledist(x1,"gumbel",start=list(a=10,b=5), silent=FALSE, control=list(trace=1), lower=0)
 
 #fitted coef around -0.6267919  0.8564855, fitted loglik -139.3812
 
@@ -405,3 +412,26 @@ if(FALSE)
   # 6 2.999036 1.999887 
   # 7 3.000222 1.999981 
 }
+
+# (20) threshold parameter leading to infinite log-density
+
+set.seed(123)
+dsexp <- function(x, rate, shift)
+  dexp(x-shift, rate=rate)
+psexp <- function(x, rate, shift)
+  pexp(x-shift, rate=rate)
+rsexp <- function(n, rate, shift)
+  rexp(n, rate=rate)+shift
+x <- rsexp(1000, 1/4, 1)
+sum(log(dsexp(x, 1, 1.995)))
+mledist(x, "sexp", start=list(rate=1, shift=0), upper= c(Inf, min(x)), control=list(trace=1, REPORT=1), optim="L-BFGS-B")
+mledist(x, "sexp", start=list(rate=1, shift=0), upper= c(Inf, min(x)), control=list(trace=1, REPORT=1))
+
+#fitted coef is 0.243 1.003 , fitted loglik -2415, fitted hessian is        
+#       rate     shift
+# rate   1000 -1.00e+03
+# shift -1000  2.14e-05
+
+
+
+

### 2024-12-04 share/checkrevdep/2024-06_v1-2-0/2024-06_v1-2-0_problems_AS.qmd
Add a 'checkrevdep' folder in 'share' to store reports of reverse dependencies checks

```{r}
#| eval: false
# version stable CRAN
install.packages("fitdistrplus")

# version dev github
install.packages("remotes")
remotes::install_github("lbbe-software/fitdistrplus")

```
```{r}
library(fitdistrplus)
```

# baytrends
```{r}
#| label: baytrends
#| warning: true
#| error: true
# dans le fichier test-impute.R
# https://github.com/cran/baytrends/blob/master/tests/testthat/test-impute.R
library(baytrends, quietly = TRUE)
x  <- dataCensored[1:20,"tdp"]
calc_impute <- impute(x,'norm')
calc_impute <- impute(x,'lnorm')
```

# BMT
```{r}
#| label: BMT
#| warning: true
#| error: true
# ligne 108 du fichier d'aide de BMTfit
# https://github.com/cran/BMT/blob/master/man/BMTfit.Rd
library(BMT, quietly = TRUE)
data("groundbeef")
serving <- groundbeef$serving
fit.mme <- BMTfit(serving, method="mme")
```

# criticality

```{r}
#| label: criticality
#| warning: true
#| error: true
# https://github.com/cran/criticality/blob/master/man/BN.Rd
library(criticality, quietly = TRUE)
ext.dir <- paste0(tempdir(), "/criticality/extdata")
dir.create(ext.dir, recursive = TRUE, showWarnings = FALSE)
extdata <- paste0(.libPaths()[1], "/criticality/extdata")
file.copy(paste0(extdata, "/facility.csv"), ext.dir, recursive = TRUE)
file.copy(paste0(extdata, "/mcnp-dataset.RData"), ext.dir, recursive = TRUE)
bb <- BN(
  facility.data = "facility.csv",
  ext.dir = ext.dir
)
```

# phenesse
```{r}
#| label: phenesse1
#| warning: true
#| error: true
# dans le fichier test-weib_percentile.R
# https://github.com/cran/phenesse/blob/master/tests/testthat/test-weib_percentile.R
# pour reproduire l'erreur, augmenter le nombre d'itérations
# très très gourmand en ressources, fait planter R, voire l'ordi !!
# ATTENTION !!
library(phenesse, quietly = TRUE)
x <- rnorm(n = 10, mean = 150, sd = 10)
weib_percentile(x, percentile = 0.5, iterations = 1)

# phenesse (local) create_predict_df(observations)
# fitdistrplus::fitdist(observations, distr = "weibull", method = "mle")
```

```{r}
#| label: phenesse2
#| warning: true
#| error: true
# dans la vignette
# ligne 54 https://github.com/cran/phenesse/blob/master/vignettes/phenesse_vignette.Rmd
# pour reproduire l'erreur, changer le nombre d'itérations
# très très gourmand en ressources, fait planter R, voire l'ordi !!
# ATTENTION !!
data(inat_examples)
s_cybele <- subset(inat_examples, scientific_name == "Speyeria cybele")
weib_percentile(observations = s_cybele$doy, percentile = 0.01, iterations = 1)
```

# stratifyR

```{r}
#| label: stratifyR
#| warning: true
#| error: true
# dans la vignette
library(stratifyR, quietly = TRUE)
Folate <- anaemia$Folate
res <- strata.data(Folate, h = 2, n = 500)
```

# traitstrap
```{r}
#| label: traitstrap1
#| warning: true
#| error: true
# trait_fit_distributions
# https://github.com/cran/traitstrap/blob/main/man/trait_fit_distributions.Rd
library(traitstrap, quietly = TRUE)
library(dplyr, quietly = TRUE)
data(community)
data(trait)

filled_traits <- trait_fill(
  comm = community |>
 filter(PlotID %in% c("A", "B")),
  traits = trait,
  scale_hierarchy = c("Site", "PlotID"),
  taxon_col = "Taxon", value_col = "Value",
  trait_col = "Trait", abundance_col = "Cover"
)

fitted_distributions <- trait_fit_distributions(
  filled_traits = filled_traits,
  distribution_type = "normal"
)
```

```{r}
#| label: traitstrap2
#| warning: true
#| error: true
test_vector <- rbeta(n = 100, shape1 = .5, shape2 = .5)
aa <- traitstrap:::get_dist_parms(data = test_vector, distribution_type = "beta")
```

### 2024-12-04 share/checkrevdep/2024-06_v1-2-0/2024-06_v1-2-0_problems_CD.qmd
Add a 'checkrevdep' folder in 'share' to store reports of reverse dependencies checks

```{r}
#| eval: false
# version stable CRAN
install.packages("fitdistrplus")

# version dev github
install.packages("remotes")
remotes::install_github("lbbe-software/fitdistrplus")

```
```{r}
library(fitdistrplus)
```

# baytrends

Le format des données n'est pas le bon. voici un exemple du contournement.

```{r}
#| label: baytrends
#| warning: true
#| error: true
# dans le fichier test-impute.R
# https://github.com/cran/baytrends/blob/master/tests/testthat/test-impute.R
library(baytrends, quietly = TRUE)
x  <- dataCensored[1:20,"tdp"]
calc_impute <- impute(x,'norm')
calc_impute <- impute(x,'lnorm')

str(x)

xnew <- data.frame(time1=x[, 1], time2=x[,2], status=x[,3])

xcorrect <- Surv2fitdistcens(xnew$time1, xnew$time2, xnew$status, type="interval")

```

# BMT

la fonction `mBMT()` définissant les moments s'arrête au moment d'ordre 4... il faut aller jusqu'à 8. un exemplde code ci-dessous.

```{r}
#| label: BMT
#| warning: true
#| error: true
# ligne 108 du fichier d'aide de BMTfit
# https://github.com/cran/BMT/blob/master/man/BMTfit.Rd
library(BMT, quietly = TRUE)
data("groundbeef")
serving <- groundbeef$serving
fit.mme <- BMTfit(serving, method="mme")
`BMTfit.mme`
mBMT2 <- function(order, p3, p4, type.p.3.4, p1, p2, type.p.1.2) 
{
  if(order == 1)
    return( BMTmean(p3, p4, type.p.3.4, p1, p2, type.p.1.2) )
  else if(order == 2)
    return( BMTsd(p3, p4, type.p.3.4, p1, p2, type.p.1.2) )
  else if(order == 3)
    return( BMTskew(p3, p4, type.p.3.4, p1, p2, type.p.1.2) )
  else if(order == 4)
    return( BMTkurt(p3, p4, type.p.3.4, p1, p2, type.p.1.2) )
  else
  {
    #equation (13)
    k1 <- 0:order
    k2 <- rep(order-k1, each=length(k1))
    k3 <- order-k1-k2
    allk <- cbind(k1, k2, k3)
    allk <- allk[allk[,3] > 0, ]
    kappal <- p3
    kappar <- p4
    Tk <- function(i)
    {
      k <- allk[i,]
      res <- 3^(k[1]+k[2]) / prod(factorial(k))
      res <- res * kappal^k1 * (1-kappar)^k2 / choose(3*order+2, 1+sum(1:3*k))
      res
    }
    sum(sapply(1:NROW(allk), Tk)) * 2 * factorial(order) / (order+1)
  }
}
dBMT2 <- dBMT
pBMT2 <- pBMT

memp <- function(x, order)
  {
        n <- length(x)
        s.mean <- mean(x)
        s.sd <- sd(x)
        res <- switch(order, s.mean, s.sd, n/((n - 1) * (n - 
            2)) * sum((x - s.mean)^3)/s.sd^3, (n - 1)/((n - 2) * 
            (n - 3)) * ((n + 1) * n/(n - 1)^2 * sum((x - s.mean)^4)/s.sd^4 - 
            3 * (n - 1)) + 3)
        return(res)
    }

type.p.3.4 <- "t w"
type.p.1.2 <- "c-d"
int.type.p.3.4 <- pmatch(type.p.3.4, c("t w", "a-s"))
int.type.p.1.2 <- pmatch(type.p.1.2, c("c-d", "l-s"))
myfix <- list(type.p.3.4=type.p.3.4, type.p.1.2 = type.p.1.2)

mmedist(serving, "BMT2", order=1:4, memp,
        start=list(p3 = 0.5, p4 = 0.5, p1 = min(serving) - 
    0.1, p2 = max(serving) + 0.1),
    fix.arg=myfix )

```

# criticality

certains jeux de données sont constants... donc une variance nulle est estimée. ça doit bloquer à un autre endroit. l'exemple est mal choisi.

```{r}
#| label: criticality
#| warning: true
#| error: true
# https://github.com/cran/criticality/blob/master/man/BN.Rd
library(criticality, quietly = TRUE)
ext.dir <- paste0(tempdir(), "/criticality/extdata")
dir.create(ext.dir, recursive = TRUE, showWarnings = FALSE)
extdata <- paste0(.libPaths()[1], "/criticality/extdata")
file.copy(paste0(extdata, "/facility.csv"), ext.dir, recursive = TRUE)
file.copy(paste0(extdata, "/mcnp-dataset.RData"), ext.dir, recursive = TRUE)
bb <- BN(
  facility.data = "facility.csv",
  ext.dir = ext.dir
)

facilitydata <- read.csv(paste0(extdata, "/facility.csv"))

for(myctrl in levels(factor(facilitydata$ctrl)))
{
  x <- subset(facilitydata, ctrl == myctrl)$thk
  if(inherits(try(fitdist(x, "gamma")), "try-error"))
  {
    cat("ctrl", myctrl, "thk values\n")
    print(x)

  }
}

```

# phenesse

il y a un vrai pb de calcul intensif. il n'est pas raisonnable d'utiliser autant de données.

```{r}
#| label: phenesse1
#| warning: true
#| error: true
# dans le fichier test-weib_percentile.R
# https://github.com/cran/phenesse/blob/master/tests/testthat/test-weib_percentile.R
# pour reproduire l'erreur, augmenter le nombre d'itérations
# très très gourmand en ressources, fait planter R, voire l'ordi !!
# ATTENTION !!
library(phenesse, quietly = TRUE)
x <- rnorm(n = 10, mean = 150, sd = 10)
#x <- rweibull(n=10, 3)
weib_percentile(x, percentile = 0.5, iterations = 1)

fitdistrplus::fitdist(x, distr = "weibull", method = "mle")

# phenesse (local) create_predict_df(observations)
# fitdistrplus::fitdist(observations, distr = "weibull", method = "mle")
```

```{r}
#| label: phenesse2
#| warning: true
#| error: true
# dans la vignette
# ligne 54 https://github.com/cran/phenesse/blob/master/vignettes/phenesse_vignette.Rmd
# pour reproduire l'erreur, changer le nombre d'itérations
# très très gourmand en ressources, fait planter R, voire l'ordi !!
# ATTENTION !!
data(inat_examples)
s_cybele <- subset(inat_examples, scientific_name == "Speyeria cybele")
weib_percentile(observations = s_cybele$doy, percentile = 0.01, iterations = 1)
weib_percentile(observations = s_cybele$doy, percentile = 0.01, iterations = 10)
```

# stratifyR

je n'ai pas compris le pb. le bug survient avant l'utilisation de `fitdist()`.

```{r}
#| label: stratifyR
#| warning: true
#| error: true
# dans la vignette
library(stratifyR, quietly = TRUE)
Folate <- anaemia$Folate
res <- strata.data(Folate, h = 2, n = 500)

#bug avant l'utilisation de fitdist()...
```

# traitstrap

certains jeux de données sont constants... donc une variance nulle est estimée. ça doit bloquer à un autre endroit. l'exemple est là encore mal choisi.

```{r}
#| label: traitstrap1
#| warning: true
#| error: true
# trait_fit_distributions
# https://github.com/cran/traitstrap/blob/main/man/trait_fit_distributions.Rd
library(traitstrap, quietly = TRUE)
library(dplyr, quietly = TRUE)
data(community)
data(trait)

filled_traits <- trait_fill(
  comm = community |>
 filter(PlotID %in% c("A", "B")),
  traits = trait,
  scale_hierarchy = c("Site", "PlotID"),
  taxon_col = "Taxon", value_col = "Value",
  trait_col = "Trait", abundance_col = "Cover"
)

fitted_distributions <- trait_fit_distributions(
  filled_traits = filled_traits,
  distribution_type = "normal"
)

for(mytaxon in levels(factor(filled_traits$Taxon)))
{
  x <- subset(filled_traits, Taxon == mytaxon)$sum_abun
  if(inherits(try(fitdist(x, "norm")), "try-error"))
  {
    cat("Taxon", mytaxon, "thk values\n")
    print(x)
  }
}

```

```{r}
#| label: traitstrap2
#| warning: true
#| error: true
test_vector <- rbeta(n = 100, shape1 = .5, shape2 = .5)
fitdist(test_vector, "beta", method="mme")
aa <- traitstrap:::get_dist_parms(data = test_vector, distribution_type = "beta")
```

### 2025-01-10 R/logLik-surface.R
update FAQ 3.2 and add defensive programing

@@ -41,10 +41,18 @@ llplot <- function(mlefit, loglik = TRUE, expansion = 1, lseq = 50,
   
   distr <- mlefit$distname
   np <- length(mlefit$estimate)
+  if(any(!is.finite(mlefit$estimate)))
+    stop("Some fitted parameters are either Inf or NaN. Use directly llsurface() or llcurve().")
+  
   if (np == 1)
   {
     estim.value <- mlefit$estimate
     estim.sd <- mlefit$sd
+    idInfNaN <- !is.finite(estim.sd)
+    if(any(idInfNaN))
+    {
+      estim.sd[idInfNaN] <- abs(estim.value[idInfNaN]) * 0.05
+    }
     plot.arg <- names(mlefit$estimate)
     fix.arg <- mlefit$fix.arg
     llcurve(data, distr, plot.arg = plot.arg, 
@@ -53,48 +61,57 @@ llplot <- function(mlefit, loglik = TRUE, expansion = 1, lseq = 50,
             lseq = lseq, fix.arg = fix.arg, loglik = loglik, weights = mlefit$weights, ...)
     if (fit.show) points(estim.value, ifelse(loglik, mlefit$loglik, exp(mlefit$loglik)), 
                          pch = fit.pch)
-  } else # so if np > 1
-    if (np == 2)
+  }else if (np == 2) # so if np > 1
+  {
+    estim.value <- mlefit$estimate
+    estim.sd <- mlefit$sd
+    idInfNaN <- !is.finite(estim.sd)
+    if(any(idInfNaN))
     {
-      estim.value <- mlefit$estimate
-      estim.sd <- mlefit$sd
-      plot.arg <- names(mlefit$estimate)
-      fix.arg <- mlefit$fix.arg
-      llsurface(data, distr, plot.arg = plot.arg, 
-                min.arg = estim.value - estim.sd * 2 *expansion, 
-                max.arg = estim.value + estim.sd * 2 *expansion, 
-                lseq = lseq, fix.arg = fix.arg, loglik = loglik,
-                back.col = back.col, nlev = nlev, pal.col = pal.col,
-                weights = mlefit$weights, ...)
-      if (fit.show) 
-        points(estim.value[1], estim.value[2], pch = fit.pch)
-    } else # so if np > 2
-    {
-      def.par <- par(no.readonly = TRUE)
-      ncombi <- choose(np, 2)
[…]

### 2025-01-10 tests/t-llplot.R

@@ -65,3 +65,21 @@ if(visualize)
   llplot(fb, fit.show = TRUE, loglik = FALSE) # should give the same
   
 }
+
+# (4) An example with NaN stderror
+# 
+
+if(visualize)
+{
+  claims <- read.csv("~/Documents/recherche-enseignement/code/R/riskassessment/bug/20250107/Claims.csv")
+  x <- claims$UltimateCost/1000
+  
+  fit_B_mle <- fitdist(x, "burr", method="mle", lower=0)
+  llplot(fit_B_mle, expansion = 10, fit.show = TRUE)
+  
+  fit_IB_mle <- fitdist(x, "invburr", method="mle", lower=0) # converges to a wrong solution
+  llplot(fit_IB_mle, expansion = 10, fit.show = TRUE)
+  
+  fit_IB_mle$estimate[1] <- NaN
+  llplot(fit_IB_mle, expansion = 10, fit.show = TRUE)
+}

### 2025-01-10 tests/t-startingvalues-invBurr-family.R

@@ -8,11 +8,11 @@ trueshape1 <- pi
 
 if(FALSE)
 {
-plot(ecdf(rburr(n, trueshape1, 2, scale=truescale)))
-lines(ecdf(1/rinvburr(n, trueshape1, 2, scale=1/truescale)), col="red")
-
-curve(pburr(x, trueshape1, 2, scale=truescale), from=0, to=5*truescale)
-curve(pinvburr(1/x, trueshape1, 2, scale=1/truescale, lower=FALSE), add=TRUE, col="red")
+  plot(ecdf(rburr(n, trueshape1, 2, scale=truescale)))
+  lines(ecdf(1/rinvburr(n, trueshape1, 2, scale=1/truescale)), col="red")
+  
+  curve(pburr(x, trueshape1, 2, scale=truescale), from=0, to=5*truescale)
+  curve(pinvburr(1/x, trueshape1, 2, scale=1/truescale, lower=FALSE), add=TRUE, col="red")
 }
 
 z <- rburr(n, trueshape1, 2, scale=truescale)
@@ -33,81 +33,6 @@ myleg <- paste("theo. Inv. Burr", paste(signif(c(trueshape1, 2, 1/truescale), 3)
 cdfcomp(fit_y_IB, xlogscale = TRUE, do.points = FALSE, addlegend = FALSE, main=mytitle)
 curve(pinvburr(x, trueshape1, 2, scale=1/truescale), add=TRUE, col="green")
 legend("topleft", lty=1, col=c("black", "red", "green"), 
-       c("empirical", "fitted", myleg))
-
-
-
-if(FALSE)
-{
-z <- rinvburr(n, trueshape1, 2, scale=truescale)
-y <- 1/z
-fit_y_B <- fitdist(y, "burr", lower=0)
-mytitle <- paste("fitted Burr", paste(signif(coef(fit_y_B), 3), collapse= ", "))
-myleg <- paste("theo. Burr", paste(signif(c(trueshape1, 2, 1/truescale), 3), collapse= ", "))
-
-cdfcomp(fit_y_B, xlogscale = TRUE, do.points = FALSE, addlegend = FALSE, main=mytitle)
-curve(pburr(x, trueshape1, 2, scale=1/truescale), add=TRUE, col="green")
-legend("topleft", lty=1, col=c("black", "red", "green"),
-       c("empirical", "fitted", myleg))
-
-
-y <- rburr(n, trueshape1, 2, scale=1/truescale)
[…]

### 2025-04-07 R/mledist.R
bug fix in mledist()

@@ -169,16 +169,16 @@ mledist <- function (data, distr, start=NULL, fix.arg=NULL, optim.method="defaul
       fnobj <- function(par, fix.arg, obs, ddistnam)
       {
         T1 <- do.call(ddistnam, c(list(obs), as.list(par), as.list(fix.arg), log=TRUE) )
-        idx <- is.finite(T1) 
-        ifelse(sum(idx) > 0, -sum( T1[idx] )/data.size, .Machine$integer.max/data.size)
+        idx <- is.infinite(T1) 
+        ifelse(sum(idx) > 0, .Machine$integer.max, -mean( T1 ))
       }
     }else
     {
       fnobj <- function(par, fix.arg, obs, ddistnam) 
       {
         T1 <- log( do.call(ddistnam, c(list(obs), as.list(par), as.list(fix.arg)) ) )
-        idx <- is.finite(T1) 
-        ifelse(sum(idx) > 0, -sum( T1[idx] )/data.size, .Machine$integer.max/data.size)
+        idx <- is.infinite(T1) 
+        ifelse(sum(idx) > 0, .Machine$integer.max, -mean( T1 ))
       }
     }
   }else if(cens && is.null(weights)) #censored data
@@ -193,8 +193,8 @@ mledist <- function (data, distr, start=NULL, fix.arg=NULL, optim.method="defaul
         p4 <- do.call(pdistnam, c(list(icens$right), as.list(par), as.list(fix.arg))) # without log=TRUE here
         p5 <- do.call(pdistnam, c(list(icens$left), as.list(par), as.list(fix.arg))) # without log=TRUE here
         T4 <- -sum(log(p4 - p5))
-        idx <- is.finite(T1) & is.finite(T2) & is.finite(T3) & is.finite(T4)
-        ifelse(sum(idx) > 0, (T1[idx]+T2[idx]+T3[idx]+T4[idx])/data.size, .Machine$integer.max/data.size)
+        idx <- is.infinite(T1) | is.infinite(T2) | is.infinite(T3) | is.infinite(T4)
+        ifelse(sum(idx) > 0, .Machine$integer.max, (T1+T2+T3+T4)/data.size)
       }
     else
       fnobjcens <- function(par, fix.arg, rcens, lcens, icens, ncens, ddistnam, pdistnam)
[…]

### 2025-04-07 tests/t-startingvalues-fellerpareto-family.R

@@ -25,7 +25,7 @@ fitdist(x, "pareto3", lower=eps, optim="L-BFGS-B")
 #fitdist(x, "pareto1", lower=eps, optim="L-BFGS-B")
 #fitdist(x, "pareto", lower=eps, optim="L-BFGS-B")
 
-fitdist(x, "llogis", lower=0, optim="L-BFGS-B")
+fitdist(x, "llogis", lower=eps, optim="L-BFGS-B")
 
 
 fitdistrplus:::startarg_fellerpareto_family(x, "fpareto")
